### 2024-11-18 paper/paper.qmd
update model

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Lexun Yu
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(here)
library(arrow)

analysis_data <- read_parquet(here("data/02-analysis_data/cleaned_data.parquet"))

```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-11-18 scripts/05-exploratory_data_analysis.R

@@ -51,6 +51,9 @@ hist(log(gust_speed_km_h), breaks=100, main="Log(Max Gust Speed)")
 
 pairs(analysis_data[, 2:8], main = "Pairwise Scatterplots (Columns 1-8)")
 
+selected_data <- analysis_data[, c("wind_speed", "pressure_station", "snow", "mean_temp", "total_rain", "gust_speed_km_h")]
+pairs(as.data.frame(selected_data), main = "Pairwise Scatterplots (Excluding Snow, Max Temp, and Min Temp)")
+
 selected_data <- analysis_data[, c("wind_speed", "pressure_station", "mean_temp", "total_rain", "gust_speed_km_h")]
 pairs(as.data.frame(selected_data), main = "Pairwise Scatterplots (Excluding Snow, Max Temp, and Min Temp)")
 

### 2024-11-18 scripts/06-model_data.R

@@ -12,11 +12,31 @@
 library(tidyverse)
 library(rstanarm)
 library(brms)
+library(car)
+library(modelsummary)
 
 #### Read data ####
 analysis_data <- read_parquet("data/02-analysis_data/cleaned_data.parquet")
 
 ### Model data ####
+
+# MLR
+
+
+m1 <- lm(mean_temp ~ wind_speed + total_precipitation + snow + 
+  pressure_station + max_temp + min_temp + total_rain + gust_speed_km_h, data=analysis_data)
+summary(m1)
+
+m2 <- lm(mean_temp ~ wind_speed + total_precipitation + snow + 
+  pressure_station + total_rain + gust_speed_km_h, data=analysis_data)
+summary(m2)
+
+m3 <- lm(mean_temp ~ wind_speed + total_precipitation +
+  pressure_station + total_rain + gust_speed_km_h, data=analysis_data)
+summary(m3)
+AIC(m3)
+
+# Bayesian Model
 formula <- bf(mean_temp ~ wind_speed + pressure_station + total_rain + gust_speed_km_h)
 
 priors <- c(
@@ -38,6 +58,26 @@ model <- brm(
 summary(model)
 
 pp_check(model)
+bayes_R2(model)
+summary(m1)$r.squared
+summary(m1)$adj.r.squared
+
+# GLM
+glm_model <- glm(mean_temp ~ wind_speed + total_precipitation + pressure_station + 
+  total_rain + gust_speed_km_h,
+  data = analysis_data,
+  family = gaussian())
+
+summary(glm_model)
+modelsummary(glm_model)
+
+glm_model_log <- glm(log(mean_temp) ~ wind_speed + total_precipitation + pressure_station +
+                 total_rain + gust_speed_km_h,
+                 data = analysis_data,
+                 family = gaussian())
+
+summary(glm_model_log)
+modelsummary(glm_model_log)
 
 #### Save model ####
 saveRDS(

### 2024-11-18 scripts/06-model_data.R
update model

@@ -14,6 +14,8 @@ library(rstanarm)
 library(brms)
 library(car)
 library(modelsummary)
+library(boot)
+
 
 #### Read data ####
 analysis_data <- read_parquet("data/02-analysis_data/cleaned_data.parquet")
@@ -37,30 +39,40 @@ summary(m3)
 AIC(m3)
 
 # Bayesian Model
-formula <- bf(mean_temp ~ wind_speed + pressure_station + total_rain + gust_speed_km_h)
+bayesian_model <- brm(mean_temp ~ wind_speed + total_precipitation + pressure_station +
+  total_rain + gust_speed_km_h,
+  data = analysis_data,
+  family = gaussian(),
+  prior = c(prior(normal(0, 10), class = "b"),
+            prior(normal(0, 10), class = "Intercept")),
+  chains = 4, iter = 2000, cores = 4)
 
-priors <- c(
-  prior(normal(0, 5), class = "b"),
-  prior(normal(0, 10), class = "Intercept")
+# Summary of results
+summary(bayesian_model)
+
+# Posterior predictive checks
+pp_check(bayesian_model)
+
+analysis_data$log_mean_temp <- log(analysis_data$mean_temp + 3)
+
+
+bayesian_model_log <- brm(
+  formula = log_mean_temp ~ wind_speed + total_precipitation + pressure_station + 
+             total_rain + gust_speed_km_h,
+  data = analysis_data,
+  family = gaussian(),
+  prior = c(
+    prior(normal(0, 10), class = "b"),       # Priors for coefficients
+    prior(normal(0, 10), class = "Intercept") # Prior for the intercept
+  ),
+  chains = 4, iter = 2000, cores = 4
 )
 
-model <- brm(
-    formula = formula,
-    data = analysis_data,
-    family = gaussian(),
-    prior = priors,
-    chains = 4, 
-    iter = 2000, 
-    warmup = 500, 
-    cores = 4
-)
+# Summary of the model
+summary(bayesian_model_log)
 
-summary(model)
-
-pp_check(model)
-bayes_R2(model)
-summary(m1)$r.squared
-summary(m1)$adj.r.squared
+# Posterior predictive checks
+pp_check(bayesian_model_log)
 
 # GLM
 glm_model <- glm(mean_temp ~ wind_speed + total_precipitation + pressure_station + 
[…]

### 2024-11-18 scripts/05-exploratory_data_analysis.R
update style

@@ -19,24 +19,24 @@ analysis_data <- read_parquet("data/02-analysis_data/cleaned_data.parquet")
 #### Plot ####
 attach(analysis_data)
 summary(analysis_data)
-hist(mean_temp, breaks=100, main="Mean Temperature", col="red")
-hist(log(mean_temp), breaks=100, main="Mean Temperature", col="red")
+hist(mean_temp, breaks = 100, main = "Mean Temperature", col = "red")
+hist(log(mean_temp), breaks = 100, main = "Mean Temperature", col = "red")
 
 par(mfrow = c(3, 3), mar = c(4, 4, 2, 1))
-hist(wind_speed, breaks=100, main="Wind Speed")
-hist(total_precipitation, breaks=100, main="Total Precipitation")
-hist(snow, breaks=100, main="total snow")
-hist(pressure_station, breaks=100, main="Pressure")
-hist(max_temp, breaks=100, main="Max Temperature")
-hist(min_temp, breaks=100, main="Min Temperature")
-hist(mean_temp, breaks=100, main="Mean Temperature")
-hist(total_rain, breaks=100, main="Rain")
-hist(gust_speed_km_h, breaks=100, main="Max Gust Speed")
+hist(wind_speed, breaks = 100, main = "Wind Speed")
+hist(total_precipitation, breaks = 100, main = "Total Precipitation")
+hist(snow, breaks = 100, main = "total snow")
+hist(pressure_station, breaks = 100, main = "Pressure")
+hist(max_temp, breaks = 100, main = "Max Temperature")
+hist(min_temp, breaks = 100, main = "Min Temperature")
+hist(mean_temp, breaks = 100, main = "Mean Temperature")
+hist(total_rain, breaks = 100, main = "Rain")
+hist(gust_speed_km_h, breaks = 100, main = "Max Gust Speed")
 
 
 
 # Set layout for multiple plots
-par(mfrow = c(3, 3), mar = c(4, 4, 2, 1))  # Adjust margins for better fit
+par(mfrow = c(3, 3), mar = c(4, 4, 2, 1)) # Adjust margins for better fit
 
 # Log-transformed histograms
 hist(log(analysis_data$wind_speed), breaks = 100, main = "Log(Wind Speed)", xlab = "Log(Wind Speed)", col = "blue")
[…]

### 2024-11-18 scripts/06-model_data.R

@@ -15,7 +15,7 @@ library(brms)
 library(car)
 library(modelsummary)
 library(boot)
-
+library(arrow)
 
 #### Read data ####
 analysis_data <- read_parquet("data/02-analysis_data/cleaned_data.parquet")
@@ -25,27 +25,31 @@ analysis_data <- read_parquet("data/02-analysis_data/cleaned_data.parquet")
 # MLR
 
 
-m1 <- lm(mean_temp ~ wind_speed + total_precipitation + snow + 
-  pressure_station + max_temp + min_temp + total_rain + gust_speed_km_h, data=analysis_data)
+m1 <- lm(mean_temp ~ wind_speed + total_precipitation + snow +
+  pressure_station + max_temp + min_temp + total_rain + gust_speed_km_h, data = analysis_data)
 summary(m1)
 
-m2 <- lm(mean_temp ~ wind_speed + total_precipitation + snow + 
-  pressure_station + total_rain + gust_speed_km_h, data=analysis_data)
+m2 <- lm(mean_temp ~ wind_speed + total_precipitation + snow +
+  pressure_station + total_rain + gust_speed_km_h, data = analysis_data)
 summary(m2)
 
 m3 <- lm(mean_temp ~ wind_speed + total_precipitation +
-  pressure_station + total_rain + gust_speed_km_h, data=analysis_data)
+  pressure_station + total_rain + gust_speed_km_h, data = analysis_data)
 summary(m3)
 AIC(m3)
 
 # Bayesian Model
-bayesian_model <- brm(mean_temp ~ wind_speed + total_precipitation + pressure_station +
-  total_rain + gust_speed_km_h,
+bayesian_model <- brm(
+  mean_temp ~ wind_speed + total_precipitation + pressure_station +
+    total_rain + gust_speed_km_h,
   data = analysis_data,
   family = gaussian(),
-  prior = c(prior(normal(0, 10), class = "b"),
-            prior(normal(0, 10), class = "Intercept")),
-  chains = 4, iter = 2000, cores = 4)
+  prior = c(
+    prior(normal(0, 10), class = "b"),
+    prior(normal(0, 10), class = "Intercept")
+  ),
+  chains = 4, iter = 2000, cores = 4
+)
 
 # Summary of results
 summary(bayesian_model)
@@ -57,12 +61,12 @@ analysis_data$log_mean_temp <- log(analysis_data$mean_temp + 3)
 
 
 bayesian_model_log <- brm(
[…]

### 2024-11-18 scripts/05-exploratory_data_analysis.R
update model

@@ -1,11 +1,9 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Graph and check the cleaned raw data
+# Author: Lexun Yu
+# Date: 18 November 2024
+# Contact: lx.yu@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
@@ -57,5 +55,3 @@ pairs(as.data.frame(selected_data), main = "Pairwise Scatterplots (Excluding Sno
 selected_data <- analysis_data[, c("wind_speed", "pressure_station", "mean_temp", "total_rain", "gust_speed_km_h")]
 pairs(as.data.frame(selected_data), main = "Pairwise Scatterplots (Excluding Snow, Max Temp, and Min Temp)")
 
-
-#### Save model ####

### 2024-11-18 scripts/06-model_data.R

@@ -1,11 +1,9 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Building GLM model and BRM model
+# Author: Lexun Yu
+# Date: 18 November 2024
+# Contact: lx.yu@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
@@ -57,8 +55,10 @@ summary(bayesian_model)
 # Posterior predictive checks
 pp_check(bayesian_model)
 
-analysis_data$log_mean_temp <- log(analysis_data$mean_temp + 3)
+analysis_data$mean_temp_F <- (analysis_data$mean_temp * 1.8) + 32
+analysis_data$log_mean_temp <- log(analysis_data$mean_temp_F)
 
+hist(analysis_data$log_mean_temp)
 
 bayesian_model_log <- brm(
   formula = log_mean_temp ~ wind_speed + total_precipitation + pressure_station +
@@ -77,6 +77,7 @@ summary(bayesian_model_log)
 
 # Posterior predictive checks
 pp_check(bayesian_model_log)
+plot(residuals(bayesian_model_log))
 
 # GLM
 glm_model <- glm(
@@ -90,7 +91,7 @@ summary(glm_model)
 modelsummary(glm_model)
 
 glm_model_log <- glm(
-  log(mean_temp) ~ wind_speed + total_precipitation + pressure_station +
+  log_mean_temp ~ wind_speed + total_precipitation + pressure_station +
     total_rain + gust_speed_km_h,
   data = analysis_data,
   family = gaussian()
@@ -99,16 +100,45 @@ glm_model_log <- glm(
 summary(glm_model_log)
 modelsummary(glm_model_log)
 
-# Model Comparison
-cv_glm <- cv.glm(data = analysis_data, glmfit = glm_model_log, K = 10)
 
-# RMSE for GLM
-rmse_glm <- sqrt(cv_glm$delta[1]) # Extract cross-validation error
-print(rmse_glm)
+#### Validation Testing ####
+set.seed(123)
[…]

### 2024-11-18 scripts/03-clean_data.R
update model and data

@@ -79,4 +79,7 @@ cleaned_data <- cleaned_data %>%
 cleaned_data <-
   na.omit(cleaned_data)
 
+cleaned_data$mean_temp_F <- (cleaned_data$mean_temp * 1.8) + 32
+cleaned_data$log_mean_temp <- log(cleaned_data$mean_temp_F)
+
 write_parquet(cleaned_data, "data/02-analysis_data/cleaned_data.parquet")

### 2024-11-18 scripts/05-exploratory_data_analysis.R

@@ -18,7 +18,8 @@ analysis_data <- read_parquet("data/02-analysis_data/cleaned_data.parquet")
 attach(analysis_data)
 summary(analysis_data)
 hist(mean_temp, breaks = 100, main = "Mean Temperature", col = "red")
-hist(log(mean_temp), breaks = 100, main = "Mean Temperature", col = "red")
+hist(mean_temp_F, breaks = 100, main = "Mean Temperature", col = "red")
+hist(log_mean_temp, breaks = 100, main = "Mean Temperature", col = "red")
 
 par(mfrow = c(3, 3), mar = c(4, 4, 2, 1))
 hist(wind_speed, breaks = 100, main = "Wind Speed")

### 2024-11-18 scripts/06-model_data.R

@@ -14,6 +14,8 @@ library(car)
 library(modelsummary)
 library(boot)
 library(arrow)
+library(MASS)
+
 
 #### Read data ####
 analysis_data <- read_parquet("data/02-analysis_data/cleaned_data.parquet")
@@ -54,9 +56,6 @@ summary(bayesian_model)
 
 # Posterior predictive checks
 pp_check(bayesian_model)
-
-analysis_data$mean_temp_F <- (analysis_data$mean_temp * 1.8) + 32
-analysis_data$log_mean_temp <- log(analysis_data$mean_temp_F)
 
 hist(analysis_data$log_mean_temp)
 
@@ -131,6 +130,7 @@ mae_brm <- mean(abs(test_data$mean_temp - brm_predictions))
 print(c(GLM_RMSE = rmse_glm, BRM_RMSE = rmse_brm))
 print(c(GLM_MAE = mae_glm, BRM_MAE = mae_brm))
 
+plot(glm_model)
 
 #### Save model ####
 saveRDS(

### 2024-11-18 paper/paper.qmd
finish part of the introduction

@@
   - Lexun Yu
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
+thanks: "Code and data are available at: [https://github.com/yulexun/ClimateChangeYVR](https://github.com/yulexun/ClimateChangeYVR)."
 date: today
@@
 
-Overview paragraph
+Climate change is a global challenges today.
+Patterns such as rising temperatures, shifting weather systems, and increased frequency of severe weather events.
+In 2021, floods swept through streets in Japanese cities, displacing millions, while extreme heat fueled wildfires in Siberia [@greenpeace_east_asia_5_2021].
+Climate change impacts human health, ecosystems, food security, water supplies, and economic stability.
+Understanding the factors driving temperature changes is necessary for designing effective mitigation strategies.
+This requires examining the various contributors to temperature variations.
+
+Some scholars have examined the changing climate.
+@xu_melting_2009 analyze the effects of rising temperatures in the Himalayas, highlighting increased frequency and duration of extreme events and shifts in ecosystems.
+These changes pose challenges to water supply, agriculture, and human populations.
+@visser_eliminating_2021 investigates the relationship between precipitation and temperature using data from the Australian Bureau of Meteorology.
+Visser's regression model indicates that average precipitation intensities increase with temperature, suggesting more intense rainfall in a warmer climate.
+The role of sea level pressure is also significant.
+@wills_systematic_2022 note that observed trends in sea level pressure have intensified warming in the Indo-Pacific Warm Pool and caused slight cooling in the eastern equatorial Pacific.
+However, as @zhang_economic_2017 argue, much of the research has focused on temperature and precipitation.
+@zhang_economic_2017 expands on this by incorporating additional predictors—relative humidity and wind speed—and concludes, using data from the Ministry of Agriculture of China, that these variables are important in understanding climate dynamics.
+
+With the data obtained from @canadian_centre_for_climate_services_adjusted_2022 and @meteorological_service_of_canada_past_2023, this research paper aims to identify the factors influencing temperature at Vancouver International Airport.
+Located on the west coast of Richmond, the airport sits on Sea Island, surrounded by water.
+As a transportation hub for passengers and freight, assessing the location's safety in a warming climate is important.
 
@@
 
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
-
 Talk more about it.
@@
 (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
 
@@
 
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
-
 # Discussion
@@
 
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
-
 \newpage

### 2024-11-18 scripts/05-exploratory_data_analysis.R

@@ -55,4 +55,3 @@ pairs(as.data.frame(selected_data), main = "Pairwise Scatterplots (Excluding Sno
 
 selected_data <- analysis_data[, c("wind_speed", "pressure_station", "mean_temp", "total_rain", "gust_speed_km_h")]
 pairs(as.data.frame(selected_data), main = "Pairwise Scatterplots (Excluding Snow, Max Temp, and Min Temp)")
-

### 2024-11-18 scripts/06-model_data.R

@@ -73,6 +73,7 @@ bayesian_model_log <- brm(
 
 # Summary of the model
 summary(bayesian_model_log)
+modelsummary(bayesian_model_log)
 
 # Posterior predictive checks
 pp_check(bayesian_model_log)

### 2024-11-19 scripts/06-model_data.R
update saved model

@@ -24,7 +24,6 @@ analysis_data <- read_parquet("data/02-analysis_data/cleaned_data.parquet")
 
 # MLR
 
-
 m1 <- lm(mean_temp ~ wind_speed + total_precipitation + snow +
   pressure_station + max_temp + min_temp + total_rain + gust_speed_km_h, data = analysis_data)
 summary(m1)
@@ -142,4 +141,18 @@ saveRDS(
 saveRDS(
   glm_model,
   file = "models/glm_model.rds"
+)
+
+saveRDS(
+  m1,
+  file = "models/mlr_1.rds"
+)
+
+saveRDS(
+  m2,
+  file = "models/mlr_2.rds"
+)
+saveRDS(
+  glm_model,
+  file = "models/glm_nolog.rds"
 )

### 2024-11-19 paper/paper.qmd
update paper intro
update model and data

@@
 library(arrow)
+library(corrplot)
 
 analysis_data <- read_parquet(here("data/02-analysis_data/cleaned_data.parquet"))
-
+m1 <- readRDS(here("models/mlr_1.rds"))
+m2 <- readRDS(here("models/mlr_2.rds"))
+glm_nolog <- readRDS(here("models/glm_nolog.rds"))
+glm <- readRDS(here("models/glm_model.rds"))
+brm <- readRDS(here("models/brm_model.rds"))
 ```
@@
 Located on the west coast of Richmond, the airport sits on Sea Island, surrounded by water.
-As a transportation hub for passengers and freight, assessing the location's safety in a warming climate is important.
+As a transportation hub for passengers and freight, it is important to assess the location's safety in a warming climate.
 
@@
 
-Why it matters paragraph
-
 Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
@@
 # Data {#sec-data}
+```{r}
+
+selected_data_1 <- analysis_data[, c("wind_speed", "pressure_station", "snow", "log_mean_temp", "total_rain", "gust_speed_km_h")]
+pairs(as.data.frame(selected_data_1), main = "Pairwise Scatterplots (Excluding Max Temp, and Min Temp)")
+cor_matrix <- cor(selected_data_1)
+print(cor_matrix)
+corrplot(cor_matrix, method = "color", type = "upper", 
+         tl.col = "black", tl.srt = 45, 
+         col = colorRampPalette(c("blue", "white", "red"))(200))
+
+selected_data_3 <- analysis_data[, c("wind_speed", "pressure_station", "log_mean_temp", "total_rain", "gust_speed_km_h")]
+pairs(as.data.frame(selected_data_3), main = "Pairwise Scatterplots of log transformed temperature (Excluding Snow, Max Temp, and Min Temp)")
+cor_matrix <- cor(selected_data_3)
+print(cor_matrix)
+corrplot(cor_matrix, method = "color", type = "upper", 
+         tl.col = "black", tl.srt = 45, 
+         col = colorRampPalette(c("blue", "white", "red"))(200))
+
+```
 
@@
 
+## MLR Model
+
+```{r}
+fitted_values <- fitted(m1)
+ggplot(data = analysis_data, aes(x = fitted_values, y = mean_temp_F)) +
+  geom_point(color = "blue", alpha = 0.7) +
+  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed", size = 1) +
+  labs(x = "Predicted Temperature", y = "Actual Temperature", title = "Actual Temperature vs. Predicted Values")
+
+```
+
+```{r}
+fitted_values <- fitted(m2)
+ggplot(data = analysis_data, aes(x = fitted_values, y = mean_temp_F)) +
+  geom_point(color = "blue", alpha = 0.7) +
+  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed", size = 1) +
+  labs(x = "Predicted Temperature", y = "Actual Temperature", title = "Actual Temperature vs. Predicted Values")
+```
+
 ### Model justification

### 2024-11-19 scripts/06-model_data.R

@@ -24,38 +24,25 @@ analysis_data <- read_parquet("data/02-analysis_data/cleaned_data.parquet")
 
 # MLR
 
-m1 <- lm(mean_temp ~ wind_speed + total_precipitation + snow +
+m1 <- lm(mean_temp_F ~ wind_speed + total_precipitation + snow +
   pressure_station + max_temp + min_temp + total_rain + gust_speed_km_h, data = analysis_data)
 summary(m1)
 
-m2 <- lm(mean_temp ~ wind_speed + total_precipitation + snow +
+m2 <- lm(mean_temp_F ~ wind_speed + total_precipitation + snow +
   pressure_station + total_rain + gust_speed_km_h, data = analysis_data)
 summary(m2)
 
-m3 <- lm(mean_temp ~ wind_speed + total_precipitation +
+m3 <- lm(mean_temp_F ~ wind_speed + total_precipitation +
   pressure_station + total_rain + gust_speed_km_h, data = analysis_data)
 summary(m3)
+
+m4 <- lm(mean_temp_F ~ wind_speed +
+  pressure_station + total_rain + gust_speed_km_h, data = analysis_data)
+summary(m4)
+
 AIC(m3)
-
+AIC(m4)
 # Bayesian Model
-bayesian_model <- brm(
-  mean_temp ~ wind_speed + total_precipitation + pressure_station +
-    total_rain + gust_speed_km_h,
-  data = analysis_data,
-  family = gaussian(),
-  prior = c(
-    prior(normal(0, 10), class = "b"),
-    prior(normal(0, 10), class = "Intercept")
-  ),
-  chains = 4, iter = 2000, cores = 4
-)
-
-# Summary of results
-summary(bayesian_model)
-
-# Posterior predictive checks
-pp_check(bayesian_model)
-
 hist(analysis_data$log_mean_temp)
 
 bayesian_model_log <- brm(
@@ -80,7 +67,7 @@ plot(residuals(bayesian_model_log))
 
 # GLM
 glm_model <- glm(
-  mean_temp ~ wind_speed + total_precipitation + pressure_station +
+  mean_temp_F ~ wind_speed + total_precipitation + pressure_station +
     total_rain + gust_speed_km_h,
   data = analysis_data,
   family = gaussian()
@@ -152,6 +139,7 @@ saveRDS(
   m2,
   file = "models/mlr_2.rds"
 )
+saveRDS(m4, file = "models/mlr_4.rds")
 saveRDS(
   glm_model,
   file = "models/glm_nolog.rds"

### 2024-11-20 paper/paper.qmd
update methodology

@@
 
-# Appendix {-}
+# Appendix {#sec-appendix}
 
-# Additional data details
+# License
+Contains information licensed under the [Open Government Licence – Canada](https://open.canada.ca/en/open-government-licence-canada)
 
-# Model details {#sec-model-details}
+# Methodology of ECCC
 
-## Posterior predictive check
+The Adjusted and Homogenized Canadian Climate Data (AHCCD) is a collection of climate datasets developed by @environment_and_climate_change_canada_adjusted_2021.
+These datasets provide long-term, quality-controlled data that have been adjusted to correct for non-climatic influences.
+
+## Population, Frame, and Sample
+
+The population of interest in the AHCCD is the entirety of Canada’s climate data, representing diverse geographical regions and climate conditions.
+The frame of the dataset are the climatological stations maintained by the ECCC that span across the countries in important locations such as airports, and banks of lakes or rivers.
+These stations record data on climate elements such as temperature, precipitation, surface pressure, and wind speed over extended periods.
+The sample is the selected stations across Canada, with adjustments applied to address inconsistencies.
+The datasets cover periods extending back to 1895 for precipitation, while other variables like wind speed and surface pressure start from 1953 or later.
+The recorded sample consists of monthly, seasonal, and annual data about surface air temperature, precipitation, pressure, and wind speed, according to @environment_and_climate_change_canada_adjusted_2021.
+
+## Sample Corrections and Adjustments
+
+The original data for AHCCD are extracted from the National Climate Data Archive of Environment Canada.
+These data include daily observations, such as maximum and minimum temperatures, precipitation, surface pressure, and wind speed.
+Observations are quality-controlled and adjusted to correct for biases due to changes in instruments, observation procedures, and other factors.
+
+Precipitation data adjustments account for wind undercatch, evaporation, and gauge-specific losses.
+According to @environment_and_climate_change_canada_climate_2021-1, corrections to account for wind undercatch, evaporation, and gauge specific wetting losses were implemented, especially in snowy conditions where snowfall is not fully captured by standard gauges.
+Corrections are made with the study by Devine and Mekis.
+
+Surface air temperature adjustments apply Quantile-Matching techniques to remove inhomogeneities.
+According to @environment_and_climate_change_canada_climate_2021, With Vincent and Wang's third generation homogenized temperature, Quantile-Matching ensures that the temperature data remain consistent across different periods, even when observation practices change.
+
+Surface pressure and wind speed data undergo adjustments based on metadata and statistical tests for systematic shifts.
+According to @environment_and_climate_change_canada_climate_2021-2, wind speed is first adjusted with a logarithmic wind profile, then tested for homogeneity using a technique based on regression models.
+It involves the identification of variation due to changes in anemometer and location change.
+The pressure data is corrected due to systematic shifts of non-updated station elevation and relocation, as stated by @environment_and_climate_change_canada_climate_2021-3.
+
+## Sampling Approach and Trade-offs
+
+According to the published methodology and the webpage by @dunbar_adjusted_2020, they employ a systematic sampling approach by selecting specific climatological stations with long-term, consistent data records.
+In some cases, observations from neighboring or overlapping stations are merged to extend time series.
+The AHCCD dataset may also contain missing values, which can vary depending on the variable, station, and time.
+Additionally, the AHCCD dataset is site-specific, meaning it provides data specific to individual observation stations.
+
+## Missing Data Handling
+
+Non-response, such as gaps in the data due to missing records, is managed by employing statistical and physical methods to homogenize the data.
+For instance, the AHCCD adjusts for shifts detected through historical evidence and metadata analysis.
+For large amount of missing data, ECCC mark the data as NA in the dataset [@canadian_centre_for_climate_services_adjusted_2022].
+
+## Strengths and Weaknesses
+
+The AHCCD by @dunbar_adjusted_2020 provides long-term, high-quality climate records adjusted for non-climatic factors such as changes in instrumentation, observation procedures, and station relocations, ensuring consistency and reliability for trend analysis in climate change.
+
+The documentation acknowledges the possibility of missing values, which naturally arise in long-term observational datasets due to factors such as station interruptions, relocation, or equipment malfunctions [@environment_and_climate_change_canada_adjusted_2021].
+Moreover, the dataset’s coverage in Arctic regions is limited to the restricted to the mid-1940s to present, as this limitation reflects the historical absence of earlier systematic observations in these remote regions.
+# Posterior predictive check
 

### 2024-11-21 paper/paper.qmd
update paper

@@
 library(corrplot)
+library(knitr)
 
+raw_data_climate <- read_csv(here("data/01-raw_data/climateyvr.csv"))
+raw_data_ahccd <- read_csv(here("data/01-raw_data/ahccdyvr.csv"))
 analysis_data <- read_parquet(here("data/02-analysis_data/cleaned_data.parquet"))
@@
 
-## Overview
-
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
-
-Overview text
-
 ## Measurement
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+The measurement of Canadian weather data involves a network of weather stations and data collection methods managed by Environment and Climate Change Canada (ECCC).
+These stations continuously measure meteorological parameters such as temperature, precipitation, wind speed, and pressure [@meteorological_service_of_canada_past_2023].
+
+According to the glossary published by @meteorological_service_of_canada_past_2023, Each day, measurement of temperature, rain, snow, precipitation, and gust speed are recorded.
+The wind and gust speed is measured in km/h with anemometer dials at a standard height of 10 meters above the ground.
+Rain and precipitation are measured in millimeter using the standard Canadian rain guage, a cylindrical container 40 cm high and 11.3 cm in diameter.
+Snow is measured in centimeters at several points that appear representative of the immediate area and then averaged.
+These raw data
+are combined to one entry and added to the historical climate database with a generated climate id and the station's location and id. Each row also have month and year of the data measured.
+
+For climate research, including climate change studies, @environment_and_climate_change_canada_adjusted_2021 has developed the Adjusted and Homogenized Canadian Climate Data (AHCCD) dataset.
+This dataset undergoes rigorous quality control and homogenization processes to address non-climatic factors that can affect long-term data consistency, such as station relocations or changes in instrumentation.
+The AHCCD ensures that observed trends reflect actual climate changes rather than artificial shifts in the data.
+In the AHCCD dataset, the precipitation, rain, pressure, snow and wind speed are adjusted with models to account for missing data and other non-climate factors.
+The detailed adjustments and corrections are documented in @sec-methodology.
+For example, precipitation measurements, which are often underestimated, are adjusted to ensure accuracy, especially in regions like the Arctic [@environment_and_climate_change_canada_climate_2021-1].
+In the AHCCD dataset, parameters measured are recorded with the units, date, station ids, location and unique identifiers.
+The AHCCD data maintains a one-to-one correspondence with the historical weather dataset by a matching station id system, ensuring that each entry in the AHCCD aligns directly with a specific observation in the historical dataset.
+
+## Raw Data
+
+In this project, we focus on weather data from YVR Airport, extracting only the datasets containing measurements taken at this specific location from the database.
+In both datasets, each row corresponds to a single averaged observation for a specific month and year.
+Each entry includes climate information such as temperature and wind speed, with their respective units recorded alongside the values.
+Additionally, a unique station ID and geographic coordinates (x, y) are included at the beginning of each entry for reference.
+The column headers of the raw historical weather dataset is displayed in @tbl-raw-climate.
+The column headers of the AHCCD dataset is displayed in @tbl-raw-ahccd.
+
+```{r}
+#| label: tbl-raw-climate
+#| tbl-cap: Column Headers of Raw Climate Data 
+#| echo: false
+#| warning: false
+raw_data <- raw_data_climate
+
+# Get the column names of your data
+column_names <- colnames(raw_data)
+
+# Reshape the column names into a matrix with, for example, 4 columns
+num_cols <- 3
+column_matrix <- matrix(column_names, ncol = num_cols, byrow = TRUE)
+
+# Convert the matrix to a data frame for kable
+column_df <- as.data.frame(column_matrix)
+
+# Display in multiple columns with kable
+kable(column_df, format = "latex", booktabs = TRUE, col.names = NULL)
+```
+
+```{r}
+#| label: tbl-raw-ahccd
+#| tbl-cap: Column Headers of Raw AHCCD Data 
+#| echo: false
+#| warning: false
+raw_data <- raw_data_ahccd
+
+# Get the column names of your data
+column_names <- colnames(raw_data)
+
+# Reshape the column names into a matrix with, for example, 4 columns
+num_cols <- 2
+column_matrix <- matrix(column_names, ncol = num_cols, byrow = TRUE)
+
+# Convert the matrix to a data frame for kable
+column_df <- as.data.frame(column_matrix)
+
+# Display in multiple columns with kable
+kable(column_df, format = "latex", booktabs = TRUE, col.names = NULL)
+```
 
@@
 
-# Methodology of ECCC
+# Methodology of ECCC  {#sec-methodology}
 
@@
 Moreover, the dataset’s coverage in Arctic regions is limited to the restricted to the mid-1940s to present, as this limitation reflects the historical absence of earlier systematic observations in these remote regions.
+
 # Posterior predictive check

### 2024-11-21 paper/paper.qmd
update model

@@
 
-## Outcome variables
+The variables in the two datasets contains the following:
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+- Geographical Information: Longitude (x) and Latitude (y), with corresponding identifiers for location (Station Name in @tbl-raw-climate, station_id and province in @tbl-raw-ahccd).
+- Temperature Metrics: Mean, maximum, and minimum temperatures (Mean Temp, Mean Max Temp, Mean Min Temp, Extr Max Temp, Extr Min Temp) and associated flags for data validity in @tbl-raw-climate.
+Similar metrics (temp_mean, temp_max, temp_min) in @tbl-raw-ahccd, with additional units included.
+- Precipitation and Snowfall: Total precipitation (Total Precip) and total snow (Total Snow), with flags for data quality in @tbl-raw-climate.
+Equivalent precipitation and snow variables (total_precip, snow) in @tbl-raw-ahccd, with units explicitly defined.
+- Wind and Gust Metrics: Direction and speed of maximum gusts (Dir of Max Gust, Spd of Max Gust) in @tbl-raw-climate, with units and flags.
+Wind speed (wind_speed) and related metrics in @tbl-raw-ahccd, with units included.
+- Pressure Information: Sea level and station pressure variables in @tbl-raw-ahccd (pressure_sea_level, pressure_station) with units.
+- Temporal Information: Date and time variables (Date/Time in @tbl-raw-climate, date, period_value in @tbl-raw-ahccd) to track observations across time periods.
+- Flags and Identifiers: Flags for data validity in both tables, such as precipitation flags, temperature flags, and identifiers like Climate ID or identifier.
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+## Data Cleaning
 
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-Talk way more about it.
-
-## Predictor variables
-
-Add graphs, tables and text.
-
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+###
 

### 2024-11-21 scripts/03-clean_data.R

@@ -40,6 +40,16 @@ data_climate <- raw_data_climate[
     raw_data_climate$date_time <= as.Date("2010-08-01"),
 ]
 
+data_ahccd_validate <- raw_data_ahccd[
+  raw_data_ahccd$date_time >= as.Date("2010-09-01") &
+    raw_data_ahccd$date_time <= as.Date("2013-04-01"),
+]
+
+data_climate_validate <- raw_data_climate[
+  raw_data_climate$date_time >= as.Date("2010-09-01") &
+    raw_data_climate$date_time <= as.Date("2013-04-01"),
+]
+
 # only keep variables related to this research
 # Define the variables to keep for both datasets
 variables_to_keep <- c(
@@ -51,6 +61,9 @@ variables_to_keep <- c(
 cleaned_data_ahccd <- data_ahccd %>%
   select(all_of(variables_to_keep))
 
+cleaned_data_ahccd_validate <- data_ahccd_validate %>%
+  select(all_of(variables_to_keep))
+
 # For the second dataset (data_climate)
 variables_to_keep_climate <- c(
   "date_time", "mean_max_temp_c", "mean_min_temp_c",
@@ -60,7 +73,11 @@ variables_to_keep_climate <- c(
 cleaned_data_climate <- data_climate %>%
   select(all_of(variables_to_keep_climate))
 
+cleaned_data_climate_validate <- data_climate_validate %>%
+  select(all_of(variables_to_keep_climate))
+
 cleaned_data <- full_join(cleaned_data_ahccd, cleaned_data_climate, by = "date_time")
+cleaned_data_validate <- full_join(cleaned_data_ahccd_validate, cleaned_data_climate_validate, by = "date_time")
 
 cleaned_data <- cleaned_data %>%
   rename(
@@ -76,10 +93,31 @@ cleaned_data <- cleaned_data %>%
     gust_speed_km_h = spd_of_max_gust_km_h
   )
 
+  cleaned_data_validate <- cleaned_data_validate %>%
+    rename(
+      date = date_time,
+      wind_speed = wind_speed_vitesse_vent,
+      total_precipitation = total_precip_precip_totale,
+      snow = snow_neige,
+      pressure_station = pressure_station_pression_station,
+      max_temp = mean_max_temp_c,
+      min_temp = mean_min_temp_c,
+      mean_temp = mean_temp_c,
[…]

### 2024-11-21 scripts/06-model_data.R

@@ -22,33 +22,39 @@ analysis_data <- read_parquet("data/02-analysis_data/cleaned_data.parquet")
 
 ### Model data ####
 
+set.seed(123)
+train_indices <- sample(1:nrow(analysis_data), size = 0.7 * nrow(analysis_data))
+train_data <- analysis_data[train_indices, ]
+test_data <- analysis_data[-train_indices, ]
+
 # MLR
 
 m1 <- lm(mean_temp_F ~ wind_speed + total_precipitation + snow +
-  pressure_station + max_temp + min_temp + total_rain + gust_speed_km_h, data = analysis_data)
+  pressure_station + max_temp + min_temp + total_rain + gust_speed_km_h, data = train_data)
 summary(m1)
 
 m2 <- lm(mean_temp_F ~ wind_speed + total_precipitation + snow +
-  pressure_station + total_rain + gust_speed_km_h, data = analysis_data)
+  pressure_station + total_rain + gust_speed_km_h, data = train_data)
 summary(m2)
 
 m3 <- lm(mean_temp_F ~ wind_speed + total_precipitation +
-  pressure_station + total_rain + gust_speed_km_h, data = analysis_data)
+  pressure_station + total_rain + gust_speed_km_h, data = train_data)
 summary(m3)
 
 m4 <- lm(mean_temp_F ~ wind_speed +
-  pressure_station + total_rain + gust_speed_km_h, data = analysis_data)
+  pressure_station + total_rain + gust_speed_km_h, data = train_data)
 summary(m4)
 
 AIC(m3)
 AIC(m4)
 # Bayesian Model
-hist(analysis_data$log_mean_temp)
+hist(train_data$log_mean_temp)
+
 
 bayesian_model_log <- brm(
   formula = log_mean_temp ~ wind_speed + total_precipitation + pressure_station +
     total_rain + gust_speed_km_h,
-  data = analysis_data,
+  data = train_data,
   family = gaussian(),
   prior = c(
     prior(normal(0, 10), class = "b"), # Priors for coefficients
@@ -69,7 +75,7 @@ plot(residuals(bayesian_model_log))
 glm_model <- glm(
   mean_temp_F ~ wind_speed + total_precipitation + pressure_station +
     total_rain + gust_speed_km_h,
-  data = analysis_data,
+  data = train_data,
   family = gaussian()
 )
 
[…]

### 2024-11-21 scripts/07-replications.R

@@ -1,16 +1,0 @@
-#### Preamble ####
-# Purpose: Replicated graphs from... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
-#### Workspace setup ####
-library(tidyverse)
-# [...UPDATE THIS...]
-
-#### Load data ####
-# [...ADD CODE HERE...]

### 2024-11-21 scripts/07-validate.R

@@ -1,0 +1,117 @@
+#### Preamble ####
+# Purpose: Replicated graphs from... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(arrow)
+library(rstantools)
+library(brms)
+
+validate_data <- read_parquet("data/02-analysis_data/cleaned_data_validate.parquet")
+bayesian_model_log <- readRDS("models/brm_model.rds")
+glm_model_log <- readRDS("models/glm_model.rds")
+
+#### BRM ####
+# Extract relevant predictor variables from `validate_data`
+predictors <- validate_data %>%
+  select(wind_speed, total_precipitation, pressure_station, total_rain, gust_speed_km_h)
+
+# Predict log-transformed mean temperature (log_mean_temp in Fahrenheit)
+predicted_log_mean_temp <- posterior_epred(
+  bayesian_model_log,
+  newdata = predictors
+)
+
+# Convert log-transformed predictions to the original scale (mean temperature in Fahrenheit)
+predicted_mean_temp_F <- exp(predicted_log_mean_temp)
+
+# Convert predicted mean temperature from Fahrenheit to Celsius
+predicted_mean_temp_C <- (predicted_mean_temp_F - 32) * 5 / 9
+
+# Calculate the mean prediction (in Celsius) for each observation
+predicted_summary <- apply(predicted_mean_temp_C, 2, function(x) {
+  mean(x)
+})
+
+# Add the real response (mean_temp in Celsius) and predictions to the validation dataset
+comparison <- validate_data %>%
+  mutate(
+    predicted_mean_temp = predicted_summary
+  ) %>%
+  select(date, mean_temp, predicted_mean_temp)
+
+# Compare real vs predicted responses
+print(comparison)
+
+comparison_brm <- comparison
+
+# Visualize the comparison between actual and predicted mean temperatures (without confidence interval)
+library(ggplot2)
+
+ggplot(comparison, aes(x = date)) +
+  geom_line(aes(y = mean_temp, color = "Actual")) +
[…]

### 2024-11-22 paper/paper.qmd
update paper data section

@@
 library(knitr)
+library(kableExtra)
 
@@
 analysis_data <- read_parquet(here("data/02-analysis_data/cleaned_data.parquet"))
+train_data <- read_parquet(here("data/02-analysis_data/train_data.parquet"))
+test_data <- read_parquet(here("data/02-analysis_data/test_data.parquet"))
 m1 <- readRDS(here("models/mlr_1.rds"))
@@
 
-With the data obtained from @canadian_centre_for_climate_services_adjusted_2022 and @meteorological_service_of_canada_past_2023, this research paper aims to identify the factors influencing temperature at Vancouver International Airport.
+Temperatures significantly impact airport operations.
+Rising temperatures significantly affect aircraft performance, potentially leading to take-off weight restrictions and the need for longer runways.
+This directly impacts airport capacity and operations [@coffel_climate_2015].
+Temperature forecast models vary in different locations, and different regions have unique climate characteristics that models may not fully capture [@american_meteorological_society_weather_nodate].
+
+This research paper aims to identify the factors influencing temperature at Vancouver International Airport and build a model for temperature prediction with the data obtained from @canadian_centre_for_climate_services_adjusted_2022 and @meteorological_service_of_canada_past_2023.
 Located on the west coast of Richmond, the airport sits on Sea Island, surrounded by water.
@@
 
+The data gathering and analysis is done in R [@citeR] with the following packages: knitr [@knitr], tidyverse [@tidyverse], ggplot2 [@ggplot2], dplyr [@dplyr], arrow [@arrow], here [@here], and lubridate [@lubridate].
+
 # Data {#sec-data}
-```{r}
 
-selected_data_1 <- analysis_data[, c("wind_speed", "pressure_station", "snow", "log_mean_temp", "total_rain", "gust_speed_km_h")]
+<!-- ```{r}
+pairs(as.data.frame(selected_data_0), main = "Pairwise Scatterplots (Including Max Temp, and Min Temp)"),
 pairs(as.data.frame(selected_data_1), main = "Pairwise Scatterplots (Excluding Max Temp, and Min Temp)")
-cor_matrix <- cor(selected_data_1)
-print(cor_matrix)
-corrplot(cor_matrix, method = "color", type = "upper", 
-         tl.col = "black", tl.srt = 45, 
-         col = colorRampPalette(c("blue", "white", "red"))(200))
-
-selected_data_3 <- analysis_data[, c("wind_speed", "pressure_station", "log_mean_temp", "total_rain", "gust_speed_km_h")]
-pairs(as.data.frame(selected_data_3), main = "Pairwise Scatterplots of log transformed temperature (Excluding Snow, Max Temp, and Min Temp)")
-cor_matrix <- cor(selected_data_3)
-print(cor_matrix)
-corrplot(cor_matrix, method = "color", type = "upper", 
-         tl.col = "black", tl.srt = 45, 
-         col = colorRampPalette(c("blue", "white", "red"))(200))
-
-```
+``` -->
 
@@
 Snow is measured in centimeters at several points that appear representative of the immediate area and then averaged.
-These raw data
-are combined to one entry and added to the historical climate database with a generated climate id and the station's location and id. Each row also have month and year of the data measured.
+These raw data are combined to one entry and added to the historical climate database with a generated climate id and the station's location and id.
+Each row also have month and year of the data measured.
 
@@
 The AHCCD data maintains a one-to-one correspondence with the historical weather dataset by a matching station id system, ensuring that each entry in the AHCCD aligns directly with a specific observation in the historical dataset.
+
+The limitations are documented in @sec-methodology.
 
@@
 
-## Data Cleaning
+## Data Cleaning {#sec-datacleaning}
 
-###
+The data cleaning process consists of two steps.
+First, we standardize and clean the column headers.
+Second, we merge the two datasets into a single combined dataset.
+The dataset used in this analysis combines information from two distinct sources: climate data (raw_data_climate) and historical weather data (raw_data_ahccd).
+The analysis spans data collected between 1959 and 2010 for training and testing purposes.
+
+The cleaned dataset contains a range of weather variables providing detailed monthly observations.
+The `date` variable represents the observation month, standardized to the first day of each month.
+**`wind_speed`** (km/h) captures average monthly wind speeds, while **`total_precipitation`** (mm) measures the total monthly precipitation, including rain and snow.
+**`snow`** (mm) records total snowfall, and **`pressure_station`** (kPa) indicates atmospheric pressure at the observation station.
+**`max_temp`** (°C), **`min_temp`** (°C), and **`mean_temp`** (°C) represent the monthly averages of maximum, minimum, and overall temperatures, respectively.
+**`total_rain`** (mm) focuses solely on rainfall amounts, distinct from snowfall.
+**`gust_speed_km_h`** (km/h) records the monthly average of maximum gust speeds.
+Additionally, constructed variables include **`mean_temp_F`**, the mean temperature was converted to Fahrenheit using
+$$
+(\text{mean\_temp} \times 1.8) + 32
+$$
+, and **`log_mean_temp`**, the log-transformed Fahrenheit temperature, was calculated as 
+$$
+\log(\text{mean\_temp\_F})
+$$
+.
+All column names were cleaned and standardized using `janitor` in @tidyverse to ensure consistency and readability.
+Dates were parsed into a unified format (`yyyy-mm-dd`) and aligned with monthly observations using the @lubridate package.
+The datasets were merged into a single combined dataset using the `date_time` variable as the common key.
+Finally, constructed variables, including `mean_temp_F` and `log_mean_temp`, were added to facilitate advanced modeling techniques.
+
+## Cleaned Data
+
+The top 6 rows of the cleaned data is displayed in @tbl-cleaned-data.
+
+```{r}
+#| label: tbl-cleaned-data
+#| tbl-cap: Sample of Cleaned Weather Data
+#| echo: false
+#| warning: false
+analysis_data |>
+  head(6) |>
+  mutate(log_mean_temp = round(log_mean_temp, 2)) |>
+  kable(
+    col.names = c("Date", "Wind Speed", "Total Precipitation", "Snow", "Pres.", "Max Temp", "Min Temp", "Mean Temp", "Rain", "Max Gust Speed", "Mean Temp in F", "Log of Mean Temp"),
+    booktabs = TRUE
+    ) 
+```
+
+The summary statistics of the combined dataset is displayed in @tbl-summary.
+
+## Characteristics in Cleaned Data
+### Skewness in Mean Temperature
+
+```{r}
+#| label: fig-hist-temp
+#| fig-cap: Mean Temp Shows More Normality and Less Skewness After Adjustment
+#| fig-subcap: ['Original Mean Temp has Skewness and Negative Numbers', 'Mean Temp in F Transformed the Value to All Positive', 'Log-Transformed Data Shows a More Symmetric and Less Skewed Distribution']
+#| echo: false
+#| warning: false
+#| layout-ncol: 2
+attach(analysis_data)
+
+hist(mean_temp, 
+     breaks = 100, 
+     main = "Distribution of Mean Temperature (°C)", 
+     xlab = "Mean Temperature (°C)", 
+     col = "red")
+
+# Histogram for Mean Temperature in Fahrenheit
+hist(mean_temp_F, 
+     breaks = 100, 
+     main = "Distribution of Mean Temperature (°F)", 
+     xlab = "Mean Temperature (°F)", 
+     col = "green")
+
+# Histogram for Log of Mean Temperature in Fahrenheit
+hist(log_mean_temp, 
+     breaks = 100, 
+     main = "Log-Transformed Mean Temperature (°F)", 
+     xlab = "Log of Mean Temperature (°F)", 
+     col = "blue")
+```
+
+@fig-hist-temp displays the histogram of the response variable Mean Temperature.
+@fig-hist-temp-1 shows the original mean temperature, which is skewed and includes negative values, making it unsuitable for direct modeling.
+To address this, We first transformed the data to Fahrenheit in @fig-hist-temp-2, shifting all values to be positive.
+However, to further normalize the distribution and reduce skewness, we applied a logarithmic transformation in @fig-hist-temp-3.
+The log transformation stabilizes variance, improves symmetry, and addresses non-linearity in the data, making it more appropriate for modelling.
+
+### Total Snow Is Zero-Inflated
+
+@fig-boxplot-snow clearly shows significant zero inflation, with a large number of observations concentrated at zero and a few extreme outliers far above the majority of the data.
+This distribution suggests that the variable snow contains excessive structural zeros, likely representing instances where no snowfall occurred.
+
+<!-- Poor Model Fit: Traditional regression models assume a continuous or discrete distribution of values.
+The extreme concentration of zeros violates this assumption, leading to biased or inaccurate parameter estimates.
+Misleading Interpretability: The effect of snow on the response variable might be confounded by the disproportionate impact of the zeros, making the interpretation of results unreliable.
+Violated Assumptions: Many models assume homoscedasticity (constant variance) and normality of residuals. The zero inflation in snow increases heteroscedasticity and skews the residuals.
+My bayesian model depends on normal likelihood, so normality and constant variance matters. Gaussian likelihood in your model assumes continuous, symmetric data -->
+
+```{r}
+#| label: fig-boxplot-snow
+#| fig-cap: Total Snow shows Zero Inflation
+#| echo: false
+#| warning: false
+boxplot(analysis_data$snow,
+        main = "Boxplot of Total Snow",
+        ylab = "Snow (in cm)", 
+        col = "lightblue",
+        notch = TRUE)
+```
+
+### Maximum Temperature, Minimum Temperature and Mean Temperature Has Strong Linear Relationships
+
+```{r}
+#| label: fig-corr-temp
+#| fig-cap: Temperature Values Have High Correlations
+#| fig-subcap: ['High Correlations Between Max, Min and Mean Temperature', 'Other Predictors Does Not Have High Correlations', 'Linear Relationship Between Max and Mean Temperature', 'Linear Relationship Between Min and Mean Temperature']
+#| echo: false
+#| warning: false
+#| layout-ncol: 2
+selected_data_0 <- analysis_data[, c("wind_speed", "pressure_station", "snow",  "min_temp", "log_mean_temp", 'max_temp', "total_rain", "gust_speed_km_h")]
+colnames(selected_data_0) <- c("Wind Speed (km/h)", "Station Pressure (hPa)", "Snow (cm)", 
+                               "Min Temperature (°C)", "Log Mean Temperature (°F)", 
+                               "Max Temperature (°C)", "Total Rain (mm)", 
+                               "Gust Speed (km/h)")
+
+cor_matrix <- cor(selected_data_0)
+corrplot(cor_matrix, method = "color", type = "upper", 
+         tl.col = "black", tl.srt = 45, 
+         col = colorRampPalette(c("blue", "white", "red"))(200))
+
+selected_data_1 <- analysis_data[, c("wind_speed", "pressure_station", "snow", "log_mean_temp", "total_rain", "gust_speed_km_h")]
+colnames(selected_data_1) <- c("Wind Speed (km/h)", "Station Pressure (hPa)", "Snow (cm)", 
+                               "Log Mean Temperature (°F)", "Total Rain (mm)", 
+                               "Gust Speed (km/h)")
+
+cor_matrix <- cor(selected_data_1)
+corrplot(cor_matrix, method = "color", type = "upper", 
+         tl.col = "black", tl.srt = 45, 
+         col = colorRampPalette(c("blue", "white", "red"))(200))
+
+# Scatter plot with linear trendline for Max Temperature vs. Mean Temperature
+ggplot(analysis_data, aes(x = `max_temp`, y = `mean_temp`)) +
+  geom_point(color = "blue", alpha = 0.6) +   # Scatter plot
+  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add linear trendline
+  labs(title = "Correlation: Max Temperature vs. Mean Temperature",
+       x = "Max Temperature (°C)",
+       y = "Mean Temperature (°C)") +
+  theme_minimal()
+
+# Scatter plot with linear trendline for Min Temperature vs. Mean Temperature
+ggplot(analysis_data, aes(x = `min_temp`, y = `mean_temp`)) +
+  geom_point(color = "green", alpha = 0.6) +   # Scatter plot
+  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add linear trendline
+  labs(title = "Correlation: Min Temperature vs. Mean Temperature",
+       x = "Min Temperature (°C)",
+       y = "Mean Temperature (°C)") +
+  theme_minimal()
+```
 
@@
 fitted_values <- fitted(m1)
-ggplot(data = analysis_data, aes(x = fitted_values, y = mean_temp_F)) +
+ggplot(data = train_data, aes(x = fitted_values, y = mean_temp_F)) +
   geom_point(color = "blue", alpha = 0.7) +
@@
 fitted_values <- fitted(m2)
-ggplot(data = analysis_data, aes(x = fitted_values, y = mean_temp_F)) +
+ggplot(data = train_data, aes(x = fitted_values, y = mean_temp_F)) +
   geom_point(color = "blue", alpha = 0.7) +
@@
 \appendix
-
-# Appendix {#sec-appendix}
+# Appendix {-}
 
@@
 Contains information licensed under the [Open Government Licence – Canada](https://open.canada.ca/en/open-government-licence-canada)
+
+# Summary Statistic of cleaned dataset
+\newpage
+```{r}
+#| label: tbl-summary
+#| tbl-cap: Summary Statistics of Raw Climate Data 
+#| echo: false
+#| warning: false
+# Generate the table
+analysis_data |>
+  select(wind_speed, total_precipitation, snow, pressure_station,
+  max_temp, min_temp) |>
+  summary() |>
+  kable(
+    col.names = c(
+      "Wind Speed", "Total Precipitation", "Snow", "Pres.", 
+      "Max Temp", "Min Temp"
+    ),
+    booktabs = TRUE,
+    align = "c",  # Center-align all columns
+    format = "latex"  # Ensure LaTeX output
+  )
+analysis_data |>
+  select(mean_temp, total_rain, gust_speed_km_h,
+  mean_temp_F, log_mean_temp) |>
+  summary() |>
+  kable(
+    col.names = c(
+      "Mean Temp", "Rain", 
+      "Max Gust Speed", "Mean Temp in F", "Log of Mean Temp"
+    ),
+    booktabs = TRUE,
+    align = "c",  # Center-align all columns
+    format = "latex"  # Ensure LaTeX output
+  )
+```
+\newpage
 

### 2024-11-22 scripts/06-model_data.R

@@ -26,6 +26,8 @@ set.seed(123)
 train_indices <- sample(1:nrow(analysis_data), size = 0.7 * nrow(analysis_data))
 train_data <- analysis_data[train_indices, ]
 test_data <- analysis_data[-train_indices, ]
+write_parquet(train_data, "data/02-analysis_data/train_data.parquet")
+write_parquet(test_data, "data/02-analysis_data/test_data.parquet")
 
 # MLR
 

### 2024-11-22 paper/paper.qmd
finish data section in paper

@@
 
-## Characteristics in Cleaned Data
+## Characteristics of Cleaned Data
 ### Skewness in Mean Temperature
@@
 boxplot(analysis_data$snow,
+        horizontal = TRUE,
         main = "Boxplot of Total Snow",
-        ylab = "Snow (in cm)", 
+        ylab = "Total Snow",
+        xlab = "Snow (in cm)", 
         col = "lightblue",
@@
 ### Maximum Temperature, Minimum Temperature and Mean Temperature Has Strong Linear Relationships
+
+@fig-corr-temp-1 highlights strong relationships between temperature variables, showing strong positive correlations between Max Temperature (°C), Min Temperature (°C), and Mean Temperature (°C).
+Scatter plots in @fig-corr-temp-3 and @fig-corr-temp-4 show near-perfect linear relationships, indicating that Max Temperature (°C) and Min Temperature (°C) are highly collinear with Mean Temperature (°C).
+In contrast, other predictors shown in @fig-corr-temp-2, such as Wind Speed (km/h), Station Pressure (hPa), and Total Rain (mm), show weaker correlations with the temperature variables and with each other, suggesting they contribute unique and independent information to the model.
 

### 2024-11-22 paper/paper.qmd
update paper

@@
 # Data {#sec-data}
-
-<!-- ```{r}
-pairs(as.data.frame(selected_data_0), main = "Pairwise Scatterplots (Including Max Temp, and Min Temp)"),
-pairs(as.data.frame(selected_data_1), main = "Pairwise Scatterplots (Excluding Max Temp, and Min Temp)")
-``` -->
 
@@
 . 
-All column names were cleaned and standardized using `janitor` in @tidyverse to ensure consistency and readability.
-Dates were parsed into a unified format (`yyyy-mm-dd`) and aligned with monthly observations using the @lubridate package.
+Additionally, a Box-Cox transformation was applied to the `total_precipitation` variable to address skewness and stabilize variance, resulting in the new variable `total_precipitation_boxcox`.
+For `gust_speed_km_h`, a log transformation was used to stabilize variance and reduce right-skewness in its distribution, creating the new variable `log_gust_speed`.
+
+All column names were cleaned and standardized using `janitor` in tidyverse [@tidyverse] to ensure consistency and readability.
+Dates were parsed into a unified format (`yyyy-mm-dd`) and aligned with monthly observations using the lubridate package [@lubridate].
 The datasets were merged into a single combined dataset using the `date_time` variable as the common key.
-Finally, constructed variables, including `mean_temp_F` and `log_mean_temp`, were added to facilitate advanced modeling techniques.
+Finally, constructed variables, including `mean_temp_F`, `total_precipitation_boxcox`, `log_gust_speed` and `log_mean_temp`, were added to the cleaned data.
 
@@
   head(6) |>
-  mutate(log_mean_temp = round(log_mean_temp, 2)) |>
+  select(
+    wind_speed, total_precipitation, snow, pressure_station, 
+    max_temp, min_temp, mean_temp, total_rain
+  ) |>
+  mutate(across(everything(), ~ round(., 2))) |>
   kable(
-    col.names = c("Date", "Wind Speed", "Total Precipitation", "Snow", "Pres.", "Max Temp", "Min Temp", "Mean Temp", "Rain", "Max Gust Speed", "Mean Temp in F", "Log of Mean Temp"),
-    booktabs = TRUE
+    col.names = c("Wind Speed", "Total Precip.", "Snow", "Pressure", 
+                  "Max Temp", "Min Temp", "Mean Temp", "Rain"),
+    align = "c",  # Center-align all columns
+    format = "latex"  # Ensure LaTeX output
+  )
+analysis_data |>
+  head(6) |>
+  select(
+    gust_speed_km_h, log_mean_temp, total_precipitation_boxcox, log_gust_speed
+  ) |>
+  mutate(across(everything(), ~ round(., 2))) |>
+  kable(
+    col.names = c("Gust Speed", "Log of Mean Temp", "Box-Cox Total Precip.", "Log of Gust Speed"),
+    align = "c",  # Center-align all columns
+    format = "latex"  # Ensure LaTeX output
   )
@@
 ## Characteristics of Cleaned Data
-### Skewness in Mean Temperature
+### Skewness in Mean Temperature, Total Precipitation and Gust Speed
 
@@
 
+In @fig-hist-all, Transformations are also applied to Total Precipitation and Gust Speed.
+Total Precipitation has a strong right skew, with most values low and a few extreme high values.
+A logarithmic transformation is used to compress the large values, reducing skewness and making the distribution more symmetric.
+For Gust Speed, a Box-Cox transformation is applied to adjust its moderate skewness.
+This transformation reshaped the data to better approximate a normal distribution.
+These adjustments improve the suitability of these variables for statistical analyses that assume normality.
+
 ### Total Snow Is Zero-Inflated
@@
 
-### Maximum Temperature, Minimum Temperature and Mean Temperature Has Strong Linear Relationships
+### Some Variables Has Strong Linear Relationships
+
+#### Maximum Temperature, Minimum Temperature and Mean Temperature
 
@@
        y = "Mean Temperature (°C)") +
+  theme_minimal()
+```
+
+#### Total Precipitation and Total Rain
+
+Similar to temperature, precipitation and total rain also have a relatively strong linear relationship as illustrated in @fig-corr-rain.
+
+```{r}
+#| label: fig-corr-rain
+#| fig-cap: Precipitation and Rain Have High Correlations
+#| echo: false
+#| warning: false
+
+ggplot(analysis_data, aes(x = `total_precipitation`, y = `total_rain`)) +
+  geom_point(color = "green", alpha = 0.6) +   # Scatter plot
+  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add linear trendline
+  labs(title = "Correlation: Total Precipitation vs. Total Rain",
+       x = "Total Precipitation (mm)",
+       y = "Total Rain (mm)") +
   theme_minimal()
@@
   )
+analysis_data |>
+  select(total_precipitation_boxcox, log_gust_speed) |>
+  summary() |>
+  kable(
+    col.names = c(
+      "Box Cox of Precipitation", "Log of Gust Speed"
+    ),
+    booktabs = TRUE,
+    align = "c",  # Center-align all columns
+    format = "latex"  # Ensure LaTeX output
+  )
 ```
 \newpage
+
+```{r}
+#| label: fig-hist-all
+#| fig-cap: Other Variables Show Normal Distribution
+#| fig-subcap: ['Wind Speed Shows Symmetric Distribution', 'Total Precipitation Shows Normal Distribution but Skewed Towards Low Values', 'Pressure Station Has a Nearly Normal Distribution', 'Gust Speed Shows Normal Distribution and Moderate Skewness', 'Box-Cox Transformed Gust Speed shows Reduces Skewness', 'Log Transformed Gust Speed shows Reduces Skewness']
+#| echo: false
+#| warning: false
+#| layout-ncol: 2
+
+df <- analysis_data
+
+# Plot histogram for wind_speed
+ggplot(df, aes(x = wind_speed)) +
+  geom_histogram(fill = "blue", color = "black", alpha = 0.7) +
+  labs(title = "Histogram of Wind Speed", x = "Wind Speed (km/h)", y = "Frequency") +
+  theme_minimal()
+
+# Plot histogram for total_precipitation
+ggplot(df, aes(x = total_precipitation)) +
+  geom_histogram(fill = "green", color = "black", alpha = 0.7) +
+  labs(title = "Histogram of Total Precipitation", x = "Total Precipitation (mm)", y = "Frequency") +
+  theme_minimal()
+
+# Plot histogram for pressure_station
+ggplot(df, aes(x = pressure_station)) +
+  geom_histogram(fill = "black", color = "black", alpha = 0.7) +
+  labs(title = "Histogram of Pressure Station", x = "Pressure (hPa)", y = "Frequency") +
+  theme_minimal()
+
+ggplot(df, aes(x = gust_speed_km_h)) +
+  geom_histogram(fill = "red", color = "black", alpha = 0.7) +
+  labs(title = "Histogram of Pressure Station", x = "Pressure (hPa)", y = "Frequency") +
+  theme_minimal()
+
+# Plot histogram for gust_speed_km_h
+ggplot(df, aes(x = total_precipitation_boxcox)) +
+  geom_histogram(fill = "green", color = "black", alpha = 0.7) +
+  labs(title = "Histogram of Box-Cox of Gust Speed", x = "Gust Speed (km/h)", y = "Frequency") +
+  theme_minimal()
+
+ggplot(df, aes(x = log_gust_speed)) +
+  geom_histogram(fill = "red", color = "black", alpha = 0.7) +
+  labs(title = "Histogram of Log of Gust Speed", x = "Gust Speed (km/h)", y = "Frequency") +
+  theme_minimal()
+
+```
+
+```{r}
+# Fit a model
+
+```
 

### 2024-11-22 scripts/03-clean_data.R

@@ -119,5 +119,77 @@ cleaned_data$log_mean_temp <- log(cleaned_data$mean_temp_F)
 cleaned_data_validate$mean_temp_F <- (cleaned_data_validate$mean_temp * 1.8) + 32
 cleaned_data_validate$log_mean_temp <- log(cleaned_data_validate$mean_temp_F)
 
+cleaned_data$log_total_precipitation <- log(cleaned_data$total_precipitation)
+cleaned_data$log_gust_speed <- log(cleaned_data$gust_speed_km_h)
+
+# Add a constant to ensure all values are positive
+cleaned_data$mean_temp_f_adj <- cleaned_data$mean_temp_F
+cleaned_data$total_precipitation_adj <- cleaned_data$total_precipitation
+cleaned_data$gust_speed_adj <- cleaned_data$gust_speed_km_h
+
+library(MASS)
+
+# Fit a simple intercept-only model
+mean_temp_model <- lm(mean_temp_f_adj ~ 1, data = cleaned_data)
+
+# Find the optimal lambda for Box-Cox
+boxcox_mean_temp <- boxcox(mean_temp_model, lambda = seq(-2, 2, by = 0.1))
+lambda_mean_temp <- boxcox_mean_temp$x[which.max(boxcox_mean_temp$y)]
+cat("Optimal lambda for mean_temp_f:", lambda_mean_temp, "\n")
+
+# Apply the transformation
+if (lambda_mean_temp == 0) {
+  cleaned_data$mean_temp_f_boxcox <- log(cleaned_data$mean_temp_f_adj)
+} else {
+  cleaned_data$mean_temp_f_boxcox <- (cleaned_data$mean_temp_f_adj^lambda_mean_temp - 1) / lambda_mean_temp
+}
+
+# Fit a simple intercept-only model
+precipitation_model <- lm(total_precipitation_adj ~ 1, data = cleaned_data)
+
+# Find the optimal lambda for Box-Cox
+boxcox_precipitation <- boxcox(precipitation_model, lambda = seq(-2, 2, by = 0.1))
+lambda_precipitation <- boxcox_precipitation$x[which.max(boxcox_precipitation$y)]
+cat("Optimal lambda for total_precipitation:", lambda_precipitation, "\n")
+
+# Apply the transformation
+if (lambda_precipitation == 0) {
+  cleaned_data$total_precipitation_boxcox <- log(cleaned_data$total_precipitation_adj)
+} else {
[…]

### 2024-11-22 scripts/06-model_data.R

@@ -53,11 +53,23 @@ AIC(m4)
 hist(train_data$log_mean_temp)
 
 
-bayesian_model_log <- brm(
-  formula = log_mean_temp ~ wind_speed + total_precipitation + pressure_station +
-    total_rain + gust_speed_km_h,
+bayesian_model_log_1 <- brm(
+  formula = log_mean_temp ~ wind_speed + pressure_station +
+    total_precipitation + gust_speed_km_h,
   data = train_data,
-  family = gaussian(),
+  family = lognormal(),
+  prior = c(
+    prior(normal(0, 10), class = "b"), # Priors for coefficients
+    prior(normal(0, 10), class = "Intercept") # Prior for the intercept
+  ),
+  chains = 4, iter = 2000, cores = 4
+)
+
+bayesian_model_log <- brm(
+  formula = log_mean_temp ~ wind_speed + pressure_station +
+    total_precipitation_boxcox + log_gust_speed,
+  data = train_data,
+  family = lognormal(),
   prior = c(
     prior(normal(0, 10), class = "b"), # Priors for coefficients
     prior(normal(0, 10), class = "Intercept") # Prior for the intercept
@@ -75,8 +87,8 @@ plot(residuals(bayesian_model_log))
 
 # GLM
 glm_model <- glm(
-  mean_temp_F ~ wind_speed + total_precipitation + pressure_station +
-    total_rain + gust_speed_km_h,
+  mean_temp_F ~ wind_speed + pressure_station +
+    total_precipitation + gust_speed_km_h,
   data = train_data,
   family = gaussian()
 )
@@ -84,9 +96,16 @@ glm_model <- glm(
 summary(glm_model)
 modelsummary(glm_model)
 
+glm_model_log_1 <- glm(
+  log_mean_temp ~ wind_speed + pressure_station +
+    total_precipitation + gust_speed_km_h,
+  data = train_data,
+  family = gaussian()
+)
+
 glm_model_log <- glm(
-  log_mean_temp ~ wind_speed + total_precipitation + pressure_station +
-    total_rain + gust_speed_km_h,
+  log_mean_temp ~ wind_speed + pressure_station +
+    total_precipitation_boxcox + log_gust_speed,
   data = train_data,
   family = gaussian()
 )
@@ -95,7 +114,6 @@ summary(glm_model_log)
 modelsummary(glm_model_log)
 
 
-#### Validation Testing ####
 #### Validation Testing ####
 
 # GLM
[…]

### 2024-11-22 scripts/07-validate.R

@@ -18,59 +18,55 @@ validate_data <- read_parquet("data/02-analysis_data/cleaned_data_validate.parqu
 bayesian_model_log <- readRDS("models/brm_model.rds")
 glm_model_log <- readRDS("models/glm_model.rds")
 
-#### BRM ####
-# Extract relevant predictor variables from `validate_data`
-predictors <- validate_data %>%
-  select(wind_speed, total_precipitation, pressure_station, total_rain, gust_speed_km_h)
+# #### BRM ####
+# # Extract relevant predictor variables from `validate_data`
+# predictors <- validate_data %>%
+#   select(wind_speed, total_precipitation, pressure_station, total_rain, gust_speed_km_h)
 
-# Predict log-transformed mean temperature (log_mean_temp in Fahrenheit)
-predicted_log_mean_temp <- posterior_epred(
-  bayesian_model_log,
-  newdata = predictors
-)
+# # Predict log-transformed mean temperature (log_mean_temp in Fahrenheit)
+# predicted_log_mean_temp <- posterior_epred(
+#   bayesian_model_log,
+#   newdata = predictors
+# )
 
-# Convert log-transformed predictions to the original scale (mean temperature in Fahrenheit)
-predicted_mean_temp_F <- exp(predicted_log_mean_temp)
+# # Convert log-transformed predictions to the original scale (mean temperature in Fahrenheit)
+# predicted_mean_temp_F <- exp(predicted_log_mean_temp)
 
-# Convert predicted mean temperature from Fahrenheit to Celsius
-predicted_mean_temp_C <- (predicted_mean_temp_F - 32) * 5 / 9
+# # Convert predicted mean temperature from Fahrenheit to Celsius
+# predicted_mean_temp_C <- (predicted_mean_temp_F - 32) * 5 / 9
 
-# Calculate the mean prediction (in Celsius) for each observation
-predicted_summary <- apply(predicted_mean_temp_C, 2, function(x) {
-  mean(x)
-})
+# # Calculate the mean prediction (in Celsius) for each observation
+# predicted_summary <- apply(predicted_mean_temp_C, 2, function(x) {
+#   mean(x)
+# })
 
-# Add the real response (mean_temp in Celsius) and predictions to the validation dataset
-comparison <- validate_data %>%
-  mutate(
-    predicted_mean_temp = predicted_summary
-  ) %>%
-  select(date, mean_temp, predicted_mean_temp)
[…]

### 2024-11-22 paper/paper.qmd
update paper

@@
 brm <- readRDS(here("models/brm_model.rds"))
+```
+
+```{r}
+# Calculate diagnostic measures
+glm_model <- glm
+# Assuming `glm_model` is your fitted model
+
+# 1. Linearity: Residuals vs Fitted Plot
+plot(glm_model, which = 1, main = "Linearity Check: Residuals vs Fitted")
+# Look for a random scatter of points. Patterns indicate non-linearity.
+
+# 2. Homoscedasticity: Scale-Location Plot
+plot(glm_model, which = 3, main = "Homoscedasticity Check: Scale-Location")
+# Points should be evenly spread. A funnel shape indicates heteroscedasticity.
+
+# 3. Normality of Residuals: Q-Q Plot
+plot(glm_model, which = 2, main = "Normality Check: Q-Q Plot")
+# Points should lie approximately along the diagonal line.
+
+# 4. Residual Histogram: Another Normality Check
+residuals <- residuals(glm_model)
+hist(residuals, main = "Histogram of Residuals", xlab = "Residuals")
+# Check for symmetry. A skewed histogram suggests non-normal residuals.
+
+# 5. Multicollinearity: Variance Inflation Factor (VIF)
+library(car)
+vif_values <- vif(glm_model)
+print(vif_values)
+# VIF > 5 indicates high multicollinearity. Consider removing highly correlated predictors.
+
+# 6. Cook's Distance: Influence of Observations
+cooksd <- cooks.distance(glm_model)
+plot(cooksd, main = "Cook's Distance", type = "h")
+abline(h = 4 / nrow(model.frame(glm_model)), col = "red", lty = 2)
+# Points above the red line are highly influential.
+
+# 7. Leverage: Identify High Leverage Points
+leverage <- hatvalues(glm_model)
+plot(leverage, main = "Leverage Check", type = "h")
+abline(h = 2 * mean(leverage), col = "red", lty = 2)
+# Points above the red line have high leverage.
+
+# 8. Correlation Between Predictors
+cor_matrix <- cor(model.frame(glm_model)[, -1]) # Exclude response variable
+print(round(cor_matrix, 2))
+# Check for highly correlated predictors (e.g., correlation > 0.8 or < -0.8).
+
 ```
@@
 #| fig-cap: Temperature Values Have High Correlations
-#| fig-subcap: ['High Correlations Between Max, Min and Mean Temperature', 'Other Predictors Does Not Have High Correlations', 'Linear Relationship Between Max and Mean Temperature', 'Linear Relationship Between Min and Mean Temperature']
+#| fig-subcap: ['High Correlations Between Max, Min and Mean Temperature, Total Precip and Rain', 'Other Predictors Does Not Have High Correlations', 'Linear Relationship Between Max and Mean Temperature', 'Linear Relationship Between Min and Mean Temperature']
 #| echo: false
@@
 #| layout-ncol: 2
-selected_data_0 <- analysis_data[, c("wind_speed", "pressure_station", "snow",  "min_temp", "log_mean_temp", 'max_temp', "total_rain", "gust_speed_km_h")]
+selected_data_0 <- analysis_data[, c("wind_speed", "pressure_station", "snow",  "min_temp", "log_mean_temp", 'max_temp', "total_rain", "gust_speed_km_h", 'total_precipitation')]
 colnames(selected_data_0) <- c("Wind Speed (km/h)", "Station Pressure (hPa)", "Snow (cm)", 
@@
                                "Max Temperature (°C)", "Total Rain (mm)", 
-                               "Gust Speed (km/h)")
+                               "Gust Speed (km/h)", "Total Precipitation (mm)")
 
@@
 
-selected_data_1 <- analysis_data[, c("wind_speed", "pressure_station", "snow", "log_mean_temp", "total_rain", "gust_speed_km_h")]
+selected_data_1 <- analysis_data[, c("wind_speed", "pressure_station", "snow", "log_mean_temp", "total_precipitation", "gust_speed_km_h")]
 colnames(selected_data_1) <- c("Wind Speed (km/h)", "Station Pressure (hPa)", "Snow (cm)", 
-                               "Log Mean Temperature (°F)", "Total Rain (mm)", 
+                               "Log Mean Temperature (°F)", "Total Precip (mm)", 
                                "Gust Speed (km/h)")

### 2024-11-22 scripts/06-model_data.R

@@ -109,7 +109,20 @@ glm_model_log <- glm(
   data = train_data,
   family = gaussian()
 )
+# Adding polynomial transformations to the model
+glm_model_poly <- glm(
+  log_mean_temp ~ 
+    poly(wind_speed, 2) +                   # Quadratic polynomial for wind_speed
+    poly(pressure_station, 2) +            # Quadratic polynomial for pressure_station
+    poly(total_precipitation_boxcox, 2) +  # Quadratic polynomial for total_precipitation_boxcox
+    poly(log_gust_speed, 2),               # Quadratic polynomial for log_gust_speed
+  data = train_data,
+  family = gaussian()
+)
 
+# Summary of the model
+summary(glm_model_poly)
+modelsummary(glm_model_poly)
 summary(glm_model_log)
 modelsummary(glm_model_log)
 
@@ -201,7 +214,7 @@ model_comparison <- tibble(
 
 print(model_comparison)
 
-model <- glm_model_log
+model <- glm_model_poly
 
 # Extract residuals and fitted values
 residuals <- resid(model)

### 2024-11-23 paper/paper.qmd
update model, data and paper

@@
 library(kableExtra)
+library(car)
 
@@
 test_data <- read_parquet(here("data/02-analysis_data/test_data.parquet"))
-m1 <- readRDS(here("models/mlr_1.rds"))
-m2 <- readRDS(here("models/mlr_2.rds"))
-glm_nolog <- readRDS(here("models/glm_nolog.rds"))
-glm <- readRDS(here("models/glm_model.rds"))
-brm <- readRDS(here("models/brm_model.rds"))
+m1 <- readRDS(here("models/m1.rds"))
+m4 <- readRDS(here("models/m4.rds"))
+glm <- readRDS(here("models/glm_poly.rds"))
+brm <- readRDS(here("models/brm.rds"))
+glm_log <- readRDS(here("models/glm_log.rds"))
 ```
@@
 plot(glm_model, which = 1, main = "Linearity Check: Residuals vs Fitted")
+plot(glm_log, which = 1, main = "Linearity Check: Residuals vs Fitted")
+
 # Look for a random scatter of points. Patterns indicate non-linearity.
@@
 plot(glm_model, which = 3, main = "Homoscedasticity Check: Scale-Location")
+plot(glm_log, which = 3, main = "Homoscedasticity Check: Scale-Location")
 # Points should be evenly spread. A funnel shape indicates heteroscedasticity.
@@
 plot(glm_model, which = 2, main = "Normality Check: Q-Q Plot")
+plot(glm_log, which = 2, main = "Normality Check: Q-Q Plot")
+
 # Points should lie approximately along the diagonal line.
@@
 hist(residuals, main = "Histogram of Residuals", xlab = "Residuals")
+residuals <- residuals(glm_log)
+hist(residuals, main = "Histogram of Residuals", xlab = "Residuals")
+
 # Check for symmetry. A skewed histogram suggests non-normal residuals.
@@
 # 5. Multicollinearity: Variance Inflation Factor (VIF)
-library(car)
 vif_values <- vif(glm_model)
+print(vif_values)
+vif_values <- vif(glm_log)
 print(vif_values)
@@
 abline(h = 4 / nrow(model.frame(glm_model)), col = "red", lty = 2)
+
+cooksd <- cooks.distance(glm_log)
+plot(cooksd, main = "Cook's Distance", type = "h")
+abline(h = 4 / nrow(model.frame(glm_log)), col = "red", lty = 2)
 # Points above the red line are highly influential.
@@
 abline(h = 2 * mean(leverage), col = "red", lty = 2)
+
+leverage <- hatvalues(glm_log)
+plot(leverage, main = "Leverage Check", type = "h")
+abline(h = 2 * mean(leverage), col = "red", lty = 2)
 # Points above the red line have high leverage.
-
-# 8. Correlation Between Predictors
-cor_matrix <- cor(model.frame(glm_model)[, -1]) # Exclude response variable
-print(round(cor_matrix, 2))
-# Check for highly correlated predictors (e.g., correlation > 0.8 or < -0.8).
 
@@
 Additionally, a Box-Cox transformation was applied to the `total_precipitation` variable to address skewness and stabilize variance, resulting in the new variable `total_precipitation_boxcox`.
-For `gust_speed_km_h`, a log transformation was used to stabilize variance and reduce right-skewness in its distribution, creating the new variable `log_gust_speed`.
+For `gust_speed_km_h`, `wind_speed` and `pressure_station`, a log transformation was used to stabilize variance and reduce right-skewness in their distribution, creating the new variable `log_gust_speed`, `log_wind_speed` and `log_pressure`.
 
@@
 The datasets were merged into a single combined dataset using the `date_time` variable as the common key.
-Finally, constructed variables, including `mean_temp_F`, `total_precipitation_boxcox`, `log_gust_speed` and `log_mean_temp`, were added to the cleaned data.
+Finally, constructed variables, including `mean_temp_F`, `total_precipitation_boxcox`, `log_gust_speed`, `log_mean_temp`, `log_wind_speed` and `log_pressure` were added to the cleaned data.
 
@@
 #| tbl-cap: Sample of Cleaned Weather Data
+#| tbl-subcap: ['Sample of Cleaned Weather Data', 'Gust Speed and Transformed Data']
 #| echo: false
@@
                   "Max Temp", "Min Temp", "Mean Temp", "Rain"),
-    align = "c",  # Center-align all columns
-    format = "latex"  # Ensure LaTeX output
+    format = "latex", booktabs = TRUE,
+  )
+
+analysis_data |>
+  head(6) |>
+  select(log_gust_speed, log_wind_speed, log_pressure
+  ) |>
+  mutate(across(everything(), ~ round(., 2))) |>
+  kable(
+    col.names = c("Log of Gust Speed", "Log of Wind Speed", "Log of Pressire"),
+    format = "latex", booktabs = TRUE,
   )
@@
   select(
-    gust_speed_km_h, log_mean_temp, total_precipitation_boxcox, log_gust_speed
+    gust_speed_km_h, log_mean_temp, total_precipitation_boxcox
   ) |>
@@
   kable(
-    col.names = c("Gust Speed", "Log of Mean Temp", "Box-Cox Total Precip.", "Log of Gust Speed"),
-    align = "c",  # Center-align all columns
-    format = "latex"  # Ensure LaTeX output
+    col.names = c("Gust Speed", "Log of Mean Temp", "Box-Cox Total Precip."),
+    format = "latex", booktabs = TRUE,
   )
@@
 ## Characteristics of Cleaned Data
+
+All of the variables in the dataset are numeric, the histograms are plotted in @fig-hist-temp and @fig-hist-all. The following section explains the characteristics of these variables.
+
 ### Skewness in Mean Temperature, Total Precipitation and Gust Speed
@@
 
-In @fig-hist-all, Transformations are also applied to Total Precipitation and Gust Speed.
+In @fig-hist-all, Transformations are also applied to Wind Speed, Pressure, Total Precipitation and Gust Speed.
 Total Precipitation has a strong right skew, with most values low and a few extreme high values.
-A logarithmic transformation is used to compress the large values, reducing skewness and making the distribution more symmetric.
+We apply log transformation to predictors including wind speed, pressure and precipitation.
 For Gust Speed, a Box-Cox transformation is applied to adjust its moderate skewness.
@@
 #| fig-cap: Total Snow shows Zero Inflation
+#| fig-subcap: ['A Large Proportion of Observations With No Snowfall', 'Prevalence of Zero Values and a Skewed Pattern in the Non-zero Snowfall Measurements']
 #| echo: false
 #| warning: false
+#| layout-ncol: 2
 boxplot(analysis_data$snow,
@@
         notch = TRUE)
+
+ggplot(analysis_data, aes(x = snow)) +
+  geom_histogram(fill = "red", color = "black", alpha = 0.7) +
+  labs(title = "Histogram of Snow", x = "Snow (cm)", y = "Frequency") +
+  theme_minimal()
+
 ```
 
-### Some Variables Has Strong Linear Relationships
+### Variables with Strong Linear Relationships
 
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+The goal of our modelling strategy is to find a model that can predict temperature changes with other weather data.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+We build a bayesian model, a general linear model and a general linear model with a 2 degrees of polynomial transformation.
+We determine the best model is the linear model with polynomial transformation.
+The detailed steps are recorded in @sec-model-select.
 
@@
 ```{r}
-fitted_values <- fitted(m2)
+fitted_values <- fitted(m4)
 ggplot(data = train_data, aes(x = fitted_values, y = mean_temp_F)) +
@@
 analysis_data |>
-  select(total_precipitation_boxcox, log_gust_speed) |>
+  select(total_precipitation_boxcox, log_gust_speed, log_wind_speed, log_pressure) |>
   summary() |>
@@
     col.names = c(
-      "Box Cox of Precipitation", "Log of Gust Speed"
+      "Box Cox of Precipitation", "Log of Gust Speed", "Log of Wind Speed",
+      "Log of Pressure"
     ),
@@
 #| fig-cap: Other Variables Show Normal Distribution
-#| fig-subcap: ['Wind Speed Shows Symmetric Distribution', 'Total Precipitation Shows Normal Distribution but Skewed Towards Low Values', 'Pressure Station Has a Nearly Normal Distribution', 'Gust Speed Shows Normal Distribution and Moderate Skewness', 'Box-Cox Transformed Gust Speed shows Reduces Skewness', 'Log Transformed Gust Speed shows Reduces Skewness']
+#| fig-subcap: ['Wind Speed Shows Symmetric Distribution', 'Total Precipitation Shows Normal Distribution but Skewed Towards Low Values', 'Pressure Station Has a Nearly Normal Distribution', 'Gust Speed Shows Normal Distribution and Moderate Skewness', 'Box-Cox Transformed Precipitation shows Reduces Skewness', 'Log Transformed Gust Speed shows Reduces Skewness', 'Log Transformed Pressure shows Reduces Skewness', 'Log Transformed Wind Speed shows Reduces Skewness']
 #| echo: false
 #| warning: false
-#| layout-ncol: 2
+#| layout-ncol: 3
 
@@
 
+# Plot histogram for log_pressure
+ggplot(df, aes(x = log_pressure)) +
+  geom_histogram(fill = "black", color = "black", alpha = 0.7) +
+  labs(title = "Histogram of Log of Pressure Station", x = "Pressure (hPa)", y = "Frequency") +
+  theme_minimal()
+
+# Plot histogram for wind_speed
+ggplot(df, aes(x = log_wind_speed)) +
+  geom_histogram(fill = "blue", color = "black", alpha = 0.7) +
+  labs(title = "Histogram of Log of Wind Speed", x = "Wind Speed (km/h)", y = "Frequency") +
+  theme_minimal()
+
 ```
 
-```{r}
-# Fit a model
-
-```
+# Additional Model Detail {#sec-model-select}
 

### 2024-11-23 scripts/03-clean_data.R

@@ -127,6 +127,10 @@ cleaned_data$mean_temp_f_adj <- cleaned_data$mean_temp_F
 cleaned_data$total_precipitation_adj <- cleaned_data$total_precipitation
 cleaned_data$gust_speed_adj <- cleaned_data$gust_speed_km_h
 
+cleaned_data$log_wind_speed <- log(cleaned_data$wind_speed)
+cleaned_data$log_pressure <- log(cleaned_data$pressure_station)
+
+
 library(MASS)
 
 # Fit a simple intercept-only model

### 2024-11-23 scripts/06-model_data.R

@@ -35,41 +35,19 @@ m1 <- lm(mean_temp_F ~ wind_speed + total_precipitation + snow +
   pressure_station + max_temp + min_temp + total_rain + gust_speed_km_h, data = train_data)
 summary(m1)
 
-m2 <- lm(mean_temp_F ~ wind_speed + total_precipitation + snow +
-  pressure_station + total_rain + gust_speed_km_h, data = train_data)
-summary(m2)
-
-m3 <- lm(mean_temp_F ~ wind_speed + total_precipitation +
-  pressure_station + total_rain + gust_speed_km_h, data = train_data)
-summary(m3)
-
 m4 <- lm(mean_temp_F ~ wind_speed +
-  pressure_station + total_rain + gust_speed_km_h, data = train_data)
+  pressure_station + total_precipitation + gust_speed_km_h, data = train_data)
 summary(m4)
 
 AIC(m3)
 AIC(m4)
 # Bayesian Model
-hist(train_data$log_mean_temp)
-
-
-bayesian_model_log_1 <- brm(
-  formula = log_mean_temp ~ wind_speed + pressure_station +
-    total_precipitation + gust_speed_km_h,
-  data = train_data,
-  family = lognormal(),
-  prior = c(
-    prior(normal(0, 10), class = "b"), # Priors for coefficients
-    prior(normal(0, 10), class = "Intercept") # Prior for the intercept
-  ),
-  chains = 4, iter = 2000, cores = 4
-)
 
 bayesian_model_log <- brm(
   formula = log_mean_temp ~ wind_speed + pressure_station +
     total_precipitation_boxcox + log_gust_speed,
   data = train_data,
-  family = lognormal(),
+  family = gaussian(),
   prior = c(
     prior(normal(0, 10), class = "b"), # Priors for coefficients
     prior(normal(0, 10), class = "Intercept") # Prior for the intercept
@@ -86,15 +64,6 @@ pp_check(bayesian_model_log)
 plot(residuals(bayesian_model_log))
 
 # GLM
-glm_model <- glm(
-  mean_temp_F ~ wind_speed + pressure_station +
-    total_precipitation + gust_speed_km_h,
-  data = train_data,
-  family = gaussian()
-)
-
-summary(glm_model)
-modelsummary(glm_model)
 
 glm_model_log_1 <- glm(
   log_mean_temp ~ wind_speed + pressure_station +
@@ -104,7 +73,7 @@ glm_model_log_1 <- glm(
[…]

### 2024-11-23 paper/paper.qmd
update the model detail

@@
 library(car)
+library(modelsummary)
 
@@
 # Points above the red line have high leverage.
+
+AIC(glm_model)
+AIC(glm_log)
+BIC(glm_model)
+BIC(glm_log)
 
@@
 
-## Cleaned Data
+### Cleaned Data
 
@@
 
-### Variables with Strong Linear Relationships
+### Variables with Strong Linear Relationships {#sec-multic-all}
 
-#### Maximum Temperature, Minimum Temperature and Mean Temperature
+#### Maximum Temperature, Minimum Temperature and Mean Temperature {#sec-multic-temp}
 
@@
 We determine the best model is the linear model with polynomial transformation.
-The detailed steps are recorded in @sec-model-select.
+The detailed steps are recorded in @sec-model-detail.
 
@@
 ## MLR Model
-
-```{r}
-fitted_values <- fitted(m1)
-ggplot(data = train_data, aes(x = fitted_values, y = mean_temp_F)) +
-  geom_point(color = "blue", alpha = 0.7) +
-  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed", size = 1) +
-  labs(x = "Predicted Temperature", y = "Actual Temperature", title = "Actual Temperature vs. Predicted Values")
-
-```
 
@@
 Contains information licensed under the [Open Government Licence – Canada](https://open.canada.ca/en/open-government-licence-canada)
+
+# Model Details {#sec-model-detail}
+## MLR Model With Every Predictor in Cleaned Data
+
+The first model predicts mean temperature ($\text{mean\_temp\_F}$) based on multiple predictors: wind speed ($\text{wind\_speed}$), total precipitation ($\text{total\_precipitation}$), snow ($\text{snow}$), station pressure ($\text{pressure\_station}$), maximum temperature ($\text{max\_temp}$), minimum temperature ($\text{min\_temp}$), total rainfall ($\text{total\_rain}$), and gust speed ($\text{gust\_speed\_km\_h}$).
+
+The fitted model is:
+
+\begin{align}
+\text{mean\_temp\_F} &= \beta_0 + \beta_1 \cdot \text{wind\_speed} + \beta_2 \cdot \text{total\_precipitation} + \beta_3 \cdot \text{snow} \\
+&+ \beta_4 \cdot \text{pressure\_station} + \beta_5 \cdot \text{max\_temp} + \beta_6 \cdot \text{min\_temp} \\
+&+ \beta_7 \cdot \text{total\_rain} + \beta_8 \cdot \text{gust\_speed\_km\_h} + \epsilon
+\end{align}
+
+- $\beta_0$: Intercept
+- $\beta_1, \beta_2, \dots, \beta_8$: Coefficients representing the change in $\text{mean\_temp\_F}$ for a one-unit increase in the respective predictor, holding other variables constant.
+- $\epsilon$: Residual error, assumed to be normally distributed with mean 0.
+
+This model has the following summary statistics in @tbl-summary-model.
+
+```{r}
+#| label: tbl-summary-model
+#| tbl-cap: Summary Statistics Shows a Large R2 in Model 1, Potential Variability in Model 2, Model L fits performs than Model 2
+#| echo: false
+#| warning: false
+modelsummary(
+  list("Model 1" = m1, "Model 2" = m4, "Model L" = glm_log),
+  output = "latex"
+)
+```
+
+The model's coefficients suggest an issue of multicollinearity, particularly due to the inclusion of highly correlated predictors such as maximum temperature, minimum temperature, and mean temperature, as discussed in @sec-multic-temp.
+Multicollinearity inflates the standard errors of the coefficients, making it difficult to determine the individual contribution of these variables to the response variable.
+Despite the model showing a perfect R2 and adjusted R2 , these metrics are misleading because the presence of highly correlated predictors often leads to overfitting.
+This is evident from the small coefficient magnitudes and nearly zero p-values, which do not reflect the true independent influence of the predictors.
+Such multicollinearity can undermine the model's interpretability and generalizability to new data.
+
+## MLR Model Without Multicollinearity Variables
+
+We then build our second model.
+
+This model predicts mean temperature (mean_temp_F) based on a subset of predictors: wind speed (wind_speed), station pressure (pressure_station), total precipitation (total_precipitation), and gust speed (gust_speed_km_h).
+
+\begin{align}
+\text{mean\_temp\_F} &= \beta_0 + \beta_1 \cdot \text{wind\_speed} + \beta_2 \cdot \text{pressure\_station} \\
+&+ \beta_3 \cdot \text{total\_precipitation} + \beta_4 \cdot \text{gust\_speed\_km\_h} + \epsilon
+\end{align}
+
+- $\beta_0$: Intercept.
+- $\beta_1, \beta_2, \beta_3, \beta_4$: Coefficients representing the change in $\text{mean\_temp\_F}$ for a one-unit increase in each respective predictor, holding others constant.
+- $\epsilon$: Residual error, assumed to be normally distributed with mean 0.
+
+This simplified model excludes highly correlated predictors, such as maximum and minimum temperatures, to reduce multicollinearity and improve interpretability.
+
+In the summary of our second model as shown in @tbl-summary-model, all predictors have relatively small coefficients, suggesting incremental effects on the response variable.
+The relatively large standard errors of some coefficients, such as the intercept, indicate potential variability or noise in the data.
+For instalce, from @fig-hist-temp and @fig-hist-all, we observe skewness and non-normal distribution in both predictor and the response.
+According to @fig-plot-m4p-1, The model does not sufficiently explain the variability in the response variable, due to non-linearity or unaddressed skewness in the data.
+This plot suggests that the model's assumptions of linearity and homoscedasticity (constant variance of residuals) are violated.
+
+```{r}
+#| label: fig-plot-m4p
+#| fig-cap: Residual vs Fitted Plot of Model 2 and L
+#| fig-subcap: ['In Model 2, The residuals are not evenly distributed', 'In Model L, The residual is more evenly distributed']
+#| echo: false
+#| warning: false
+#| layout-ncol: 2
+plot(m4, which = 1)
+plot(glm_log, which = 1)
+```
+
+## MLR Model with transformed variables
+In our third model L, we use log and Box-Cox transformation to ensure linearity and homoscedasticity in all predictors and the response.
+The detailed steps are documented in @sec-multic-all.
+This linear model predicts the log-transformed mean temperature (`log_mean_temp`) based on log-transformed wind speed (`log_wind_speed`), log-transformed pressure (`log_pressure`), Box-Cox-transformed total precipitation (`total_precipitation_boxcox`), and log-transformed gust speed (`log_gust_speed`).
+
+We build our Model L as the following:
+
+\begin{align}
+\text{log\_mean\_temp} &= \beta_0 + \beta_1 \cdot \text{log\_wind\_speed} + \beta_2 \cdot \text{log\_pressure} \\
+&+ \beta_3 \cdot \text{total\_precipitation\_boxcox} + \beta_4 \cdot \text{log\_gust\_speed} + \epsilon
+\end{align}
+
+- $\beta_0$: Intercept.
+- $\beta_1, \beta_2, \beta_3, \beta_4$: Coefficients representing the change in `log_mean_temp` for a one-unit increase in each predictor, holding other variables constant.
+- $\epsilon$: Residual error, assumed to follow a Gaussian (normal) distribution.
+
+The inclusion of the Box-Cox-transformed total precipitation further refines the model by accommodating non-linearity in precipitation data.
+The Gaussian family ensures that the residuals of the response variable follow a normal distribution after the transformations.
+As shown in @fig-plot-m4p-2, this model reduces heteroscedasticity, minimizes non-linear patterns in residuals, and improves overall interpretability and fit.
+Each coefficient indicates the multiplicative effect of a one-unit change in the respective predictor on the mean temperature after applying the logarithmic transformations.
+This model fits better than Model 2, as indicated in @tbl-summary-model, as the R2 and adjusted R2 are higher, AIC, BIC are smaller.
+
+## Polynomial Linear Model
+The residual plot in @fig-plot-m4p-2 shows a non-linear pattern, as indicated by the curved trend in the residuals.
+This suggests that the relationship between the predictors and the response variable is not fully captured by a linear model.
+Adding polynomial terms could help address this non-linearity by allowing the model to fit curved relationships.
 
@@
   )
+
 ```
+
 \newpage

### 2024-11-23 scripts/06-model_data.R

@@ -39,7 +39,6 @@ m4 <- lm(mean_temp_F ~ wind_speed +
   pressure_station + total_precipitation + gust_speed_km_h, data = train_data)
 summary(m4)
 
-AIC(m3)
 AIC(m4)
 # Bayesian Model
 
@@ -65,28 +64,17 @@ plot(residuals(bayesian_model_log))
 
 # GLM
 
-glm_model_log_1 <- glm(
-  log_mean_temp ~ wind_speed + pressure_station +
-    total_precipitation + gust_speed_km_h,
-  data = train_data,
-  family = gaussian()
-)
+glm_model_log <- lm(log_mean_temp ~ log_wind_speed + log_pressure +
+  total_precipitation_boxcox + log_gust_speed, data = train_data)
 
-glm_model_log <- glm(
-  log_mean_temp ~ log_wind_speed + log_pressure +
-    total_precipitation_boxcox + log_gust_speed,
-  data = train_data,
-  family = gaussian()
-)
 # Adding polynomial transformations to the model
-glm_model_poly <- glm(
+glm_model_poly <- lm(
   log_mean_temp ~ 
     poly(log_wind_speed, 2) +                   # Quadratic polynomial for wind_speed
     poly(log_pressure, 2) +            # Quadratic polynomial for pressure_station
     poly(total_precipitation_boxcox, 2) +  # Quadratic polynomial for total_precipitation_boxcox
     poly(log_gust_speed, 2),               # Quadratic polynomial for log_gust_speed
-  data = train_data,
-  family = gaussian()
+  data = train_data
 )
 
 # Summary of the model

### 2024-11-23 paper/paper.qmd
update brm model section

@@
 
+## Bayesian Model
+After fitting the linear regression model using log and Box-Cox transformations, We extend the analysis by testing a Bayesian regression model (Model B).
+This model also predicts the log-transformed mean temperature (`log_mean_temp`) but incorporates prior and Bayesian inference to evaluate the uncertainty of parameter estimates.
+The predictors remain the same: wind speed (`wind_speed`), station pressure (`pressure_station`), Box-Cox-transformed total precipitation (`total_precipitation_boxcox`), and log-transformed gust speed (`log_gust_speed`).
+
+The Bayesian model is defined as:
+
+\begin{align}
+\text{log\_mean\_temp} &\sim \mathcal{N}(\mu, \sigma^2), \\
+\mu &= \beta_0 + \beta_1 \cdot \text{wind\_speed} + \beta_2 \cdot \text{pressure\_station} \\
+&\quad + \beta_3 \cdot \text{total\_precipitation\_boxcox} + \beta_4 \cdot \text{log\_gust\_speed}.
+\end{align}
+
+The prior distributions for the parameters are:
+
+- Coefficients ($\beta_1, \beta_2, \beta_3, \beta_4$): 
+  $$
+  \beta_i \sim \mathcal{N}(0, 10), \quad \text{for } i = 1, 2, 3, 4,
+  $$
+
+  reflecting moderate uncertainty centered around zero.
+
+- Intercept ($\beta_0$): 
+  $$
+  \beta_0 \sim \mathcal{N}(0, 10),
+  $$
+  indicating prior uncertainty about the baseline log-mean temperature.
+
+The model was fit using **Hamiltonian Monte Carlo (HMC)** via the `brms` package @brm. It uses:
+- **4 chains** for convergence,
+- **2000 iterations** per chain to ensure stability, 
+- **4 cores** for parallel computation, enabling efficient sampling.
+
 ## Polynomial Linear Model

### 2024-11-24 paper/paper.qmd
update paper

@@
 library(modelsummary)
+library(brms)
 
@@
 These stations continuously measure meteorological parameters such as temperature, precipitation, wind speed, and pressure [@meteorological_service_of_canada_past_2023].
+We choose the datesets from the Government of Canada for their coverage and quality control processes.
 
@@
 Second, we merge the two datasets into a single combined dataset.
-The dataset used in this analysis combines information from two distinct sources: climate data (raw_data_climate) and historical weather data (raw_data_ahccd).
-The analysis spans data collected between 1959 and 2010 for training and testing purposes.
+The dataset used in this analysis combines information from two distinct sources: historical weather data (raw_data_climate) and AHCCD weather data (raw_data_ahccd).
+The analysis spans data collected monthly between August 1959 and August 2010 for training and testing purposes.
 
@@
 . 
-Additionally, a Box-Cox transformation was applied to the `total_precipitation` variable to address skewness and stabilize variance, resulting in the new variable `total_precipitation_boxcox`.
+Moreover, a Box-Cox transformation was applied to the `total_precipitation` variable to address skewness and stabilize variance, resulting in the new variable `total_precipitation_boxcox`.
 For `gust_speed_km_h`, `wind_speed` and `pressure_station`, a log transformation was used to stabilize variance and reduce right-skewness in their distribution, creating the new variable `log_gust_speed`, `log_wind_speed` and `log_pressure`.
@@
 
-### Cleaned Data
+### Cleaned Data and Training/Testing Split
 
@@
 
+The cleaned and combined dataset is splitted into training group and testing group randomly.
+The training dataset contains 70% of the cleaned dataset and the testing dataset contains 30% of the cleaned dataset.
+The training dataset is used to fit the model.
+The test dataset is used to evaluate the model's performance on unseen data.
+
 ## Characteristics of Cleaned Data
@@
 
-### Skewness in Mean Temperature, Total Precipitation and Gust Speed
+### Skewness in Variables
 
@@
 
-# Model
+# Model {#sec-model}
 
@@
 
-We build a bayesian model, a general linear model and a general linear model with a 2 degrees of polynomial transformation.
+We build a bayesian model, a linear model and a linear model with a 2 degrees of polynomial transformation.
+We choose linear regression instead of general linear model because all of the variables are numeric.
+According to @kumar_glm_2023, linear regression is applicable when the response is continuous and approximately normally distributed.
 We determine the best model is the linear model with polynomial transformation.
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+The final model we choose is the linear model with polynomial transformation.
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+This polynomial linear regression model predicts the log-transformed mean temperature (`log_mean_temp`) based on quadratic polynomial transformations of four predictors:
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+- log-transformed wind speed (`log_wind_speed`), 
+- log-transformed pressure (`log_pressure`), 
+- Box-Cox-transformed total precipitation (`total_precipitation_boxcox`), and 
+- log-transformed gust speed (`log_gust_speed`).
 
 ## MLR Model
-
-```{r}
-fitted_values <- fitted(m4)
-ggplot(data = train_data, aes(x = fitted_values, y = mean_temp_F)) +
-  geom_point(color = "blue", alpha = 0.7) +
-  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed", size = 1) +
-  labs(x = "Predicted Temperature", y = "Actual Temperature", title = "Actual Temperature vs. Predicted Values")
-```
 
@@
   $$
+
   indicating prior uncertainty about the baseline log-mean temperature.
 
-The model was fit using **Hamiltonian Monte Carlo (HMC)** via the `brms` package @brm. It uses:
-- **4 chains** for convergence,
-- **2000 iterations** per chain to ensure stability, 
-- **4 cores** for parallel computation, enabling efficient sampling.
+The model was fit using the `brms` package @brm. It uses:
+
+- 4 chains for convergence,
+- 2000 iterations per chain to ensure stability, 
+- 4 cores for parallel computation, enabling efficient sampling.
+
+Unlike linear regression, which provides point estimates and assumes fixed parameter values, Bayesian regression incorporates prior knowledge and generates posterior distributions, offering a probabilistic framework that quantifies uncertainty in parameter estimates.
+
+## The Linear Model has Better Fit than Bayesian Model
+
+We calculate the RMSE value and MAE value of the Bayesian model and the Linear model based on the test dataset, because it evaluates how well the model performs on data it has never seen before, providing a realistic measure of predictive accuracy.
+
+The RMSE measures the average squared difference between the observed ($y_i$) and predicted ($\hat{y}_i$) values. It is calculated as:
+
+$$
+\text{RMSE} = \sqrt{\frac{1}{n} \sum_{i=1}^n (y_i - \hat{y}_i)^2}
+$$
+
+Where:
+
+- $y_i$: The actual value of the $i$-th observation.
+- $\hat{y}_i$: The predicted value for the $i$-th observation.
+- $n$: The total number of observations.
+
+The MAE measures the average absolute difference between the observed ($y_i$) and predicted ($\hat{y}_i$) values. It is calculated as:
+
+$$
+\text{MAE} = \frac{1}{n} \sum_{i=1}^n |y_i - \hat{y}_i|
+$$
+
+Where:
+
+- $y_i$: The actual value of the $i$-th observation.
+- $\hat{y}_i$: The predicted value for the $i$-th observation.
+- $n$: The total number of observations.
+
+By comparing the RMSE and MAE results in @tbl-brm, the Linear Model L has slightly lower RMSE and MAE compared to the Bayesian model, suggesting it has a marginally better fit for the given data.
+As a result, we choose the Linear Model over the Bayesian Model.
+
+```{r}
+#| label: tbl-brm
+#| tbl-cap: Summary of Bayesian Model and Comparison of RMSE / MAE
+#| tbl-subcap: ['Bayesian Model Shows Similar R2 value Compare to Linear', 'Linear Model shows better fit']
+#| echo: false
+#| warning: false
+#| layout-ncol: 2
+modelsummary(brm)
+
+# RMSE and MAE
+# GLM
+glm_model <- glm_log
+
+# BRM
+brm_model <- brm
+
+# GLM predictions (reverse log transformation and convert to Celsius)
+glm_predictions_log <- predict(glm_model, newdata = test_data)
+glm_predictions_F <- exp(glm_predictions_log)  # Predicted in Fahrenheit
+glm_predictions <- (glm_predictions_F - 32) * 5 / 9  # Convert to Celsius
+
+# BRM predictions (posterior predictive mean, reverse log transformation and convert to Celsius)
+brm_predictions_log <- colMeans(posterior_predict(brm_model, newdata = test_data))
+brm_predictions_F <- exp(brm_predictions_log)  # Predicted in Fahrenheit
+brm_predictions <- (brm_predictions_F - 32) * 5 / 9  # Convert to Celsius
+
+# Ensure consistency: Predicted and actual values in Celsius
+test_data <- test_data %>%
+  mutate(
+    actual_mean_temp = mean_temp,      # Actual values (Celsius)
+    glm_predicted = glm_predictions,  # GLM predictions (Celsius)
+    brm_predicted = brm_predictions   # Bayesian predictions (Celsius)
+  )
+
+# Calculate RMSE
+rmse_glm <- sqrt(mean((test_data$actual_mean_temp - test_data$glm_predicted)^2))
+rmse_brm <- sqrt(mean((test_data$actual_mean_temp - test_data$brm_predicted)^2))
+
+# Calculate MAE
+mae_glm <- mean(abs(test_data$actual_mean_temp - test_data$glm_predicted))
+mae_brm <- mean(abs(test_data$actual_mean_temp - test_data$brm_predicted))
+
+# Create a data frame for the metrics
+metrics <- data.frame(
+  Model = c("Linear", "Bayesian"),
+  RMSE = c(rmse_glm, rmse_brm),
+  MAE = c(mae_glm, mae_brm)
+)
+
+# Display the table using kable
+kable(metrics) %>%
+  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))
+
+```
 
 ## Polynomial Linear Model
+Since we choose the Linear Model L over Bayesian Model, We inspect the residual plot of the linear model.
 The residual plot in @fig-plot-m4p-2 shows a non-linear pattern, as indicated by the curved trend in the residuals.
@@
 Adding polynomial terms could help address this non-linearity by allowing the model to fit curved relationships.
+The detail of the new model is discussed in @sec-model.
 
-# Summary Statistic of cleaned dataset
 \newpage
+
+# Additional Data Details
+
+The summary statistics of the cleaned data are shown in @tbl-summary.
+
 ```{r}
@@
 
+As discussed in @sec-data, the histogram of variables before and after transformation is displayed in @fig-hist-all.
 ```{r}
@@
 ```
-
-# Additional Model Detail {#sec-model-select}
 

### 2024-11-24 scripts/06-model_data.R

@@ -86,42 +86,6 @@ modelsummary(glm_model_log)
 
 #### Validation Testing ####
 
-# GLM
-glm_model <- glm_model_poly
-
-# BRM
-brm_model <- bayesian_model_log
-
-# GLM predictions (reverse log transformation and convert to Celsius)
-glm_predictions_log <- predict(glm_model, newdata = test_data)
-glm_predictions_F <- exp(glm_predictions_log)  # Predicted in Fahrenheit
-glm_predictions <- (glm_predictions_F - 32) * 5 / 9  # Convert to Celsius
-
-# BRM predictions (posterior predictive mean, reverse log transformation and convert to Celsius)
-brm_predictions_log <- colMeans(posterior_predict(brm_model, newdata = test_data))
-brm_predictions_F <- exp(brm_predictions_log)  # Predicted in Fahrenheit
-brm_predictions <- (brm_predictions_F - 32) * 5 / 9  # Convert to Celsius
-
-# Ensure consistency: Predicted and actual values in Celsius
-test_data <- test_data %>%
-  mutate(
-    actual_mean_temp = mean_temp,      # Actual values (Celsius)
-    glm_predicted = glm_predictions,  # GLM predictions (Celsius)
-    brm_predicted = brm_predictions   # Bayesian predictions (Celsius)
-  )
-
-# Calculate RMSE
-rmse_glm <- sqrt(mean((test_data$actual_mean_temp - test_data$glm_predicted)^2))
-rmse_brm <- sqrt(mean((test_data$actual_mean_temp - test_data$brm_predicted)^2))
-
-# Calculate MAE
-mae_glm <- mean(abs(test_data$actual_mean_temp - test_data$glm_predicted))
-mae_brm <- mean(abs(test_data$actual_mean_temp - test_data$brm_predicted))
-
-# Output Results
-print(c(GLM_RMSE = rmse_glm, BRM_RMSE = rmse_brm))
-print(c(GLM_MAE = mae_glm, BRM_MAE = mae_brm))
-
 # Visualize actual vs predicted
 library(ggplot2)
 

### 2024-11-25 paper/paper.qmd
update paper

@@
 
-The goal of our modelling strategy is to find a model that can predict temperature changes with other weather data.
+The goal of our modelling strategy is to find a model that can predict temperature changes with other weather data as predictors such as wind speed, pressure, precipitation and gust speed.
 
-We build a bayesian model, a linear model and a linear model with a 2 degrees of polynomial transformation.
-We choose linear regression instead of general linear model because all of the variables are numeric.
-According to @kumar_glm_2023, linear regression is applicable when the response is continuous and approximately normally distributed.
+We build a linear model, a bayesian model, and a linear model with a 2 degrees of polynomial transformation.
 We determine the best model is the linear model with polynomial transformation.
 The detailed steps are recorded in @sec-model-detail.
+We choose linear regression instead of general linear model because all of the variables are numeric.
+According to @kumar_glm_2023, linear regression is applicable when the response is continuous and approximately normally distributed, which is more applicable to our dataset.
 
@@
 
+The model introduces non-linear relationships by including polynomial terms up to the second degree (quadratic) for each predictor.
+
 ## MLR Model
+
+The model is mathematically expressed as:
+
+\begin{align}
+\text{Log Mean Temperature} &= \beta_0 
++ \beta_{1} \cdot \text{Log Wind Speed} + \beta_{2} \cdot (\text{Log Wind Speed})^2 \\
+&+ \beta_{3} \cdot \text{Log Pressure} + \beta_{4} \cdot (\text{Log Pressure})^2 \\
+&+ \beta_{5} \cdot \text{Box-Cox Total Precipitation} + \beta_{6} \cdot (\text{Box-Cox Total Precipitation})^2 \\
+&+ \beta_{7} \cdot \text{Log Gust Speed} + \beta_{8} \cdot (\text{Log Gust Speed})^2 \\
+&+ \epsilon
+\end{align}
+
+Where:
+
+1. **Intercept ($\beta_0$):**
+   - Represents the baseline log-transformed mean temperature when all predictors are zero.
+
+2. **Predictors and Their Polynomial Terms:**
+   - **Wind Speed (Log Wind Speed):**
+     - $\beta_{1} \cdot \text{Log Wind Speed}$ models the linear effect of wind speed.
+     - $\beta_{2} \cdot (\text{Log Wind Speed})^2$ captures the quadratic (non-linear) effect of wind speed.
+   - **Atmospheric Pressure (Log Pressure):**
+     - $\beta_{3} \cdot \text{Log Pressure}$ models the linear effect of atmospheric pressure.
+     - $\beta_{4} \cdot (\text{Log Pressure})^2$ captures the quadratic effect of atmospheric pressure.
+   - **Total Precipitation (Box-Cox Total Precipitation):**
+     - $\beta_{5} \cdot \text{Box-Cox Total Precipitation}$ models the linear effect of total precipitation.
+     - $\beta_{6} \cdot (\text{Box-Cox Total Precipitation})^2$ captures the quadratic effect of total precipitation.
+   - **Gust Speed (Log Gust Speed):**
+     - $\beta_{7} \cdot \text{Log Gust Speed}$ models the linear effect of gust speed.
+     - $\beta_{8} \cdot (\text{Log Gust Speed})^2$ captures the quadratic effect of gust speed.
+
+3. **Residual Error ($\epsilon$):**
+   - Represents the variation in the log-transformed mean temperature that is not explained by the predictors.
+
+The purpose of including polynomial terms in this model is to capture non-linear relationships between the predictors and the response variable, allowing the model to fit more complex, curved patterns that a purely linear model cannot accommodate.
+By incorporating quadratic terms for each predictor, the model gains flexibility in representing real-world data relationships, reducing bias, and addressing systematic patterns in residuals that may indicate non-linearity.
+This enhances the model's predictive performance, making it capable of explaining variance that would otherwise remain unaccounted for in a simple linear regression, while still maintaining the interpretability and simplicity of a linear regression framework.
 

### 2024-11-25 paper/paper.qmd
update model section and result section

@@
 glm_log <- readRDS(here("models/glm_log.rds"))
-```
-
-```{r}
-# Calculate diagnostic measures
-glm_model <- glm
-# Assuming `glm_model` is your fitted model
-
-# 1. Linearity: Residuals vs Fitted Plot
-plot(glm_model, which = 1, main = "Linearity Check: Residuals vs Fitted")
-plot(glm_log, which = 1, main = "Linearity Check: Residuals vs Fitted")
-
-# Look for a random scatter of points. Patterns indicate non-linearity.
-
-# 2. Homoscedasticity: Scale-Location Plot
-plot(glm_model, which = 3, main = "Homoscedasticity Check: Scale-Location")
-plot(glm_log, which = 3, main = "Homoscedasticity Check: Scale-Location")
-# Points should be evenly spread. A funnel shape indicates heteroscedasticity.
-
-# 3. Normality of Residuals: Q-Q Plot
-plot(glm_model, which = 2, main = "Normality Check: Q-Q Plot")
-plot(glm_log, which = 2, main = "Normality Check: Q-Q Plot")
-
-# Points should lie approximately along the diagonal line.
-
-# 4. Residual Histogram: Another Normality Check
-residuals <- residuals(glm_model)
-hist(residuals, main = "Histogram of Residuals", xlab = "Residuals")
-residuals <- residuals(glm_log)
-hist(residuals, main = "Histogram of Residuals", xlab = "Residuals")
-
-# Check for symmetry. A skewed histogram suggests non-normal residuals.
-
-# 5. Multicollinearity: Variance Inflation Factor (VIF)
-vif_values <- vif(glm_model)
-print(vif_values)
-vif_values <- vif(glm_log)
-print(vif_values)
-# VIF > 5 indicates high multicollinearity. Consider removing highly correlated predictors.
-
-# 6. Cook's Distance: Influence of Observations
-cooksd <- cooks.distance(glm_model)
-plot(cooksd, main = "Cook's Distance", type = "h")
-abline(h = 4 / nrow(model.frame(glm_model)), col = "red", lty = 2)
-
-cooksd <- cooks.distance(glm_log)
-plot(cooksd, main = "Cook's Distance", type = "h")
-abline(h = 4 / nrow(model.frame(glm_log)), col = "red", lty = 2)
-# Points above the red line are highly influential.
-
-# 7. Leverage: Identify High Leverage Points
-leverage <- hatvalues(glm_model)
-plot(leverage, main = "Leverage Check", type = "h")
-abline(h = 2 * mean(leverage), col = "red", lty = 2)
-
-leverage <- hatvalues(glm_log)
-plot(leverage, main = "Leverage Check", type = "h")
-abline(h = 2 * mean(leverage), col = "red", lty = 2)
-# Points above the red line have high leverage.
-
-AIC(glm_model)
-AIC(glm_log)
-BIC(glm_model)
-BIC(glm_log)
-
 ```
@@
 
-The model is mathematically expressed as:
+The Model LP is mathematically expressed as:
 
@@
 
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
+### Model Validation {#sec-validation}
+
+The linear regression model was chosen to analyze the relationship between wind speed, atmospheric pressure, total precipitation, and gust speed with the log-transformed mean temperature.
+The primary goal of this model is to quantify the effect of each predictor on the response variable and to make predictions.
+The linear model is suitable because the relationship between the predictors and the response variable was found to be approximately linear after applying logarithmic and Box-Cox transformations to address non-linearity and skewness.
+
+#### Assumptions Fit the Data
+
+```{r}
+#| label: fig-assumption
+#| fig-cap: Model LP meets all assumtions of linear regression
+#| fig-subcap: ['Linearity Check: Residuals vs Fitted', 'Normality Check: Q-Q Plot', 'Homoscedasticity Check: Scale-Location']
+#| echo: false
+#| warning: false
+#| layout-ncol: 2
+
+plot(glm, which = 1)
+plot(glm, which = 2)
+plot(glm, which = 3)
+```
+
+The model, which uses linear regression, has four assumptions: linearity, homoscedasticity, and approximate normality of residuals.
+
+The linearity assumption states that the relationship between each predictor and the response variable is linear.
+This assumption was evaluated by examining scatterplots of the predictors against the response variable and by analyzing residuals versus fitted values.
+As shown in @fig-assumption-1, The red smooth line indicates only a slight curvature, and the the residuals scatter reasonably randomly around the 0 line.
+The resulting diagnostic plots showed no obvious systematic patterns, indicating that the linearity assumption is reasonably satisfied.
+
+Homoscedasticity requires that the variance of the residuals remains constant across all levels of the predictors.
+In @fig-assumption-3, the points (standardized residuals) are evenly spread across the range of fitted values.
+There is no clear fan or funnel shape, suggesting that the variance of the residuals is fairly constant, which supports the assumption of homoscedasticity.
+
+Linear regression assumes that the residuals follow a normal distribution.
+This assumption was assessed using Q-Q plots in @fig-assumption-2, where the residuals were compared to a theoretical normal distribution.
+Most points aligned closely with the diagonal line, indicating that the residuals are approximately normally distributed.
+
+Independence assumes that each observation in the dataset is unrelated to others.
+This was ensured during data collection or preparation according to ECCC's methodology as discussed in @sec-methodology.
+
+Multicollinearity occurs when predictors are highly correlated, making it difficult to isolate their individual effects.
+Variance Inflation Factor (VIF) quantifies how much multicollinearity inflates the variance of the estimated regression coefficients.
+The Generalized VIF (GVIF) is used because each polynomial term represents more than one degree of freedom.
+The result in @tbl-vif are below the commonly accepted threshold of 5, indicating low multicollinearity in the predictors.
+
+```{r}
+#| label: tbl-vif
+#| tbl-cap: VIF Value Indicates Low Multicollinearity
+#| echo: false
+#| warning: false
+
+# Calculate VIF
+vif_values <- vif(glm)
+kable(vif_values)
+```
+
+#### Validation on Test Data
+
+We validate our Model LP with test data to evaluate the model's generalizability and performance on unseen data.
+
+The R2 value calculated on the test data ($R^2 = 0.59$) is very close to the R2 value of the model on the training data ($R^2 = 0.61$), as discussed in @sec-result-r2 This similarity indicates that the model generalizes well to unseen data, as its performance on the test set is consistent with its performance during training.
+The small difference between the two R2 values suggests that the model is not overfitting the training data and captures the underlying relationships in the data effectively.
+
+```{r}
+#| label: tbl-test-validate
+#| tbl-cap: R2 value on test data is close to train data
+#| echo: false
+#| warning: false
+
+test_data$predicted <- predict(glm, newdata = test_data)
+SSR <- sum((test_data$log_mean_temp - test_data$predicted)^2)  # Sum of Squared Residuals
+TSS <- sum((test_data$log_mean_temp - mean(test_data$log_mean_temp))^2)  # Total Sum of Squares
+R2 <- 1 - (SSR / TSS)
+R2T <- summary(glm)$r.squared
+results <- data.frame(
+  Metric = c("SSR", "TSS", "R2", "R2-Train"),
+  Value = c(SSR, TSS, R2, R2T)
+)
+
+kable(results)
+```
+
+### Model Justification
+
+The decision to use a linear regression model with polynomial terms for this analysis is grounded in the need to capture non-linear relationships between the predictors and the log-transformed mean temperature (log_mean_temp).
+Initial exploratory data analysis and diagnostic plots indicated that the relationships between the predictors, such as log-transformed wind speed, pressure, total precipitation, and gust speed, and the response variable were not strictly linear.
+Applying second-degree polynomial terms allows the model to account for curvature and interactions within the data while maintaining interpretability.
+
+The polynomial linear regression model was selected over alternative non-linear methods due to its balance between complexity and simplicity.
+Polynomial terms extend the linear model's capacity to capture non-linear patterns while preserving the interpretability of regression coefficients.
+The inclusion of transformations, such as logarithmic and Box-Cox, further improves the model by addressing issues of skewness and heteroscedasticity observed in the raw data.
+
+Finally, the model assumptions were rigorously evaluated in @sec-validation.
+Residual plots confirmed that the linearity assumption was largely met, while the Scale-Location plot and diagnostic tests supported the homoscedasticity of residuals.
+Variance Inflation Factor (VIF) values for the predictors were all well below the critical threshold, indicating no significant multicollinearity.
+These results ensure that the model remains valid and reliable for inference and prediction.
 
@@
 
+```{r}
+#| label: tbl-modelresults
+#| tbl-cap: Summary of Model LP
+#| echo: false
+#| warning: false
+modelsummary(glm)
+```
+
+## Intercept Indicates a Strong Baseline In Temperature
+
+The intercept of the model is estimated at **3.899**, representing the predicted value of the log-transformed mean temperature when all predictors are at their mean (after centering and scaling in the polynomial terms).
+This value serves as the baseline against which the effects of all predictors are measured.
+A significant intercept indicates the model reliably predicts the baseline log-mean temperature.
+
+## Wind Speed: Non-Linear Increase and Decline in Temperature
+The first-degree polynomial term for log-transformed wind speed ($\text{poly(log\_wind\_speed, 2)1}$) has a positive coefficient of **0.260**, suggesting that increasing wind speed is associated with a slight increase in $\log(\text{mean\_temp})$.
+However, the second-degree term ($\text{poly(log\_wind\_speed, 2)2}$) has a negative coefficient of **-0.449**, indicating a reversing effect at higher wind speeds.
+This quadratic relationship reflects non-linear relationships between wind speed and temperature.
+
+## Atmospheric Pressure: Consistent Negative Impact on Temperature
+Log-transformed atmospheric pressure shows a consistently negative relationship with $\log(\text{mean\_temp})$.
+The first-degree term ($\text{poly(log\_pressure, 2)1}$) has a coefficient of **-0.388**, and the second-degree term ($\text{poly(log\_pressure, 2)2}$) is more negative at **-0.973**.
+This indicates that higher atmospheric pressure is associated with a reduction in $\log(\text{mean\_temp})$, and the quadratic term amplifies this effect, particularly at extreme values of pressure.
+
+## Total Precipitation: The Strongest Negative Effect
+The Box-Cox transformed total precipitation has the strongest impact on $\log(\text{mean\_temp})$ among all predictors.
+The first-degree term ($\text{poly(total\_precipitation\_boxcox, 2)1}$) has a significant negative coefficient of **-2.016**, implying that higher precipitation reduces $\log(\text{mean\_temp})$ substantially.
+The second-degree term ($\text{poly(total\_precipitation\_boxcox, 2)2}$) is positive at **0.218**, indicating a slight mitigation of this negative effect at extreme precipitation levels.
+
+## Gust Speed: Decreases Temperature but Shows Curvature
+The first-degree term for log-transformed gust speed ($\text{poly(log\_gust\_speed, 2)1}$) has a negative coefficient of **-0.980**, showing that increasing gust speed reduces $\log(\text{mean\_temp})$.
+However, the second-degree term ($\text{poly(log\_gust\_speed, 2)2}$) has a positive coefficient of **0.347**, suggesting a minor curvature in the relationship where the negative effect is less pronounced at higher gust speeds.
+
+## Model Metrics: Explains 61% Variance with Low RMSE
+The model explains a significant portion of the variance in $\log(\text{mean\_temp})$, as indicated by an $R^2$ value of **0.610** and an adjusted $R^2$ of **0.603**.
+These values demonstrate that approximately 61% of the variability in $\log(\text{mean\_temp})$ is explained by the predictors, even after accounting for the complexity of the model.
+
+The model has a low RMSE of **0.12**, indicating accurate predictions with minimal average error. The AIC (**-601.2**) and BIC (**-560.6**) values suggest a strong fit compared to alternative models.
+
 # Discussion
 
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
+## Summary of Results: Precipitation Leads, While Wind and Pressure Add Complexity {#sec-first-point}
+
+This polynomial regression model captures non-linear relationships between predictors and $\log(\text{mean\_temp})$ effectively, with total precipitation exerting the strongest influence.
+Wind speed and gust speed contribute with both positive and negative effects, depending on the quadratic curvature.
+Atmospheric pressure consistently shows a negative impact.
+Despite its strengths, slight evidence of heteroscedasticity and non-linear residual patterns observed in diagnostic plots suggest that future refinements, such as interaction terms or alternative modeling techniques, could improve the model's performance.
+
+## The finding does not agree with Visser
+## the finding does match with Wills
 
@@
 
+### Limitations on the Model LP
+
+While the polynomial linear regression model captures much of the relationship between the predictors and the log-transformed mean temperature, some limitations remain.
+@fig-assumption-1 indicates that the relationship is not completely linear, with minor deviations suggesting that the model may not fully account for certain non-linearities.
+Additionally, @fig-assumption-3 shows slight evidence of heteroscedasticity at the extreme fitted values, indicating that the variance of residuals is not entirely constant.
+These limitations suggest that the model could benefit from further refinement, such as applying additional transformations, exploring interaction terms, or using alternative techniques like weighted least squares or robust regression to address potential heteroscedasticity.
 Weaknesses and next steps should also be included.
+
+### Limitation on the Response Mean Temperature
+
+### Limitation on only researching temperature
 
@@
 
-# Model Details {#sec-model-detail}
+# Additional Model Details {#sec-model-detail}
 ## MLR Model With Every Predictor in Cleaned Data
@@
 #| label: tbl-brm
-#| tbl-cap: Summary of Bayesian Model and Comparison of RMSE / MAE
+#| tbl-cap: Summary of Bayesian Model and Comparison of RMSE / MAE on Test Data
 #| tbl-subcap: ['Bayesian Model Shows Similar R2 value Compare to Linear', 'Linear Model shows better fit']

### 2024-11-26 paper/paper.qmd
update paper result

@@
 The role of sea level pressure is also significant.
-@wills_systematic_2022 note that observed trends in sea level pressure have intensified warming in the Indo-Pacific Warm Pool and caused slight cooling in the eastern equatorial Pacific.
+@wills_systematic_2022 note that observed trends in sea level pressure have caused slight cooling in the eastern equatorial Pacific.
 However, as @zhang_economic_2017 argue, much of the research has focused on temperature and precipitation.
@@
 Temperatures significantly impact airport operations.
-Rising temperatures significantly affect aircraft performance, potentially leading to take-off weight restrictions and the need for longer runways.
+Rising temperatures significantly affect aircraft performance, leading to take-off weight restrictions and the need for longer runways.
 This directly impacts airport capacity and operations [@coffel_climate_2015].
@@
 
-## MLR Model
+## MLR Polynomial Model
 
@@
 The purpose of including polynomial terms in this model is to capture non-linear relationships between the predictors and the response variable, allowing the model to fit more complex, curved patterns that a purely linear model cannot accommodate.
-By incorporating quadratic terms for each predictor, the model gains flexibility in representing real-world data relationships, reducing bias, and addressing systematic patterns in residuals that may indicate non-linearity.
 This enhances the model's predictive performance, making it capable of explaining variance that would otherwise remain unaccounted for in a simple linear regression, while still maintaining the interpretability and simplicity of a linear regression framework.
@@
 
-The R2 value calculated on the test data ($R^2 = 0.59$) is very close to the R2 value of the model on the training data ($R^2 = 0.61$), as discussed in @sec-result-r2 This similarity indicates that the model generalizes well to unseen data, as its performance on the test set is consistent with its performance during training.
+The R2 value calculated on the test data ($R^2 = 0.59$) is very close to the R2 value of the model on the training data ($R^2 = 0.61$), as discussed in @sec-result-r2.
+The $R^2$ value is calculated using the following formula:
+
+\begin{align}
+R^2 &= 1 - \frac{\text{SSR}}{\text{TSS}} \\
+\text{SSR} &= \sum_{i=1}^{n} (y_i - \hat{y}_i)^2 \\
+\text{TSS} &= \sum_{i=1}^{n} (y_i - \bar{y})^2
+\end{align}
+
+Where:
+
+- $\text{SSR}$ (Sum of Squared Residuals) measures the total squared difference between the observed values ($y_i$) and the predicted values ($\hat{y}_i$).
+- $\text{TSS}$ (Total Sum of Squares) measures the total squared difference between the observed values ($y_i$) and their mean ($\bar{y}$).
+- $R^2$ quantifies the proportion of the variance in the response variable explained by the model.
+
+ This similarity indicates that the model generalizes well to unseen data, as its performance on the test set is consistent with its performance during training.
 The small difference between the two R2 values suggests that the model is not overfitting the training data and captures the underlying relationships in the data effectively.
@@
 The inclusion of transformations, such as logarithmic and Box-Cox, further improves the model by addressing issues of skewness and heteroscedasticity observed in the raw data.
-
-Finally, the model assumptions were rigorously evaluated in @sec-validation.
+By incorporating quadratic terms for each predictor, the model gains flexibility in representing real-world data relationships, reducing bias, and addressing systematic patterns in residuals that may indicate non-linearity.
+
+Finally, the model assumptions were evaluated in @sec-validation.
 Residual plots confirmed that the linearity assumption was largely met, while the Scale-Location plot and diagnostic tests supported the homoscedasticity of residuals.
@@
 
-## Model Metrics: Explains 61% Variance with Low RMSE
+## Model Metrics: Explains 61% Variance with Low RMSE {#sec-result-r2}
 The model explains a significant portion of the variance in $\log(\text{mean\_temp})$, as indicated by an $R^2$ value of **0.610** and an adjusted $R^2$ of **0.603**.
@@
 The model has a low RMSE of **0.12**, indicating accurate predictions with minimal average error. The AIC (**-601.2**) and BIC (**-560.6**) values suggest a strong fit compared to alternative models.
+
+## Predicted vs. Actual Plot shows good overall fit
+
+```{r}
+#| label: fig-predictedvsfit
+#| fig-cap: Predicted vs. Actual plot shows good fit
+#| fig-subcap: ['The model is accurate in the train data', 'The model is accurate in the test data']
+#| echo: false
+#| warning: false
+#| layout-ncol: 2
+#| 
+train_data$predicted <- predict(glm, newdata = train_data)
+test_data$predicted <- predict(glm, newdata = test_data)
+
+ggplot(train_data, aes(x = log_mean_temp, y = predicted)) +
+  geom_point(alpha = 0.6) +
+  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
+  labs(title = "Predicted vs Actual Log Mean Temperature",
+       x = "Actual Log Mean Temperature",
+       y = "Predicted Log Mean Temperature") +
+  theme_minimal()
+ggplot(test_data, aes(x = log_mean_temp, y = predicted)) +
+  geom_point(alpha = 0.6) +
+  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
+  labs(title = "Predicted vs Actual Log Mean Temperature",
+       x = "Actual Log Mean Temperature",
+       y = "Predicted Log Mean Temperature") +
+  theme_minimal()
+```
+
+As displayed in @fig-predictedvsfit, points aligning closely along the diagonal line indicate that the predicted log-transformed mean temperature matches well with the observed values.
+In @fig-predictedvsfit-1, The points closely align along the red diagonal line, indicating that the model predictions are highly accurate for the training dataset.
+This alignment suggests that the model captures the relationship between the predictors and the log-transformed mean temperature effectively, with minimal bias or variance issues within the training data.
+In @fig-predictedvsfit-2, the points on the test data plot also align reasonably well along the red diagonal line, though with slightly more scatter compared to the training data.
+This result indicates that the model generalizes well to unseen data, maintaining good predictive accuracy and avoiding overfitting.
+The consistency between the train and test plots suggests the model has a good balance between complexity and generalization.
+In summary, the Predicted vs.
+Actual plots provide clear visual evidence of the model's accuracy and robustness for both training and test datasets.
 
@@
 ## The finding does not agree with Visser
+
 ## the finding does match with Wills
-
-Please don't use these as sub-heading labels - change them to be what your point actually is.
-
-## Third discussion point
 
@@
 These limitations suggest that the model could benefit from further refinement, such as applying additional transformations, exploring interaction terms, or using alternative techniques like weighted least squares or robust regression to address potential heteroscedasticity.
-Weaknesses and next steps should also be included.
 
@@
 ### Limitation on only researching temperature
+
+### Limitation on ECCC's methodology
 

### 2024-11-26 paper/paper.qmd
modified:   paper/paper.qmd

--- 2024-11-25 paper/paper.qmd
@@
 library(brms)
+library(Metrics)
 
@@
 The R2 value calculated on the test data ($R^2 = 0.59$) is very close to the R2 value of the model on the training data ($R^2 = 0.61$), as discussed in @sec-result-r2 This similarity indicates that the model generalizes well to unseen data, as its performance on the test set is consistent with its performance during training.
-The small difference between the two R2 values suggests that the model is not overfitting the training data and captures the underlying relationships in the data effectively.
+The training MSE of 0.0137 and test MSE of 0.0158 indicate the model performs well on both datasets with minimal overfitting.
+The The small difference between the two R2 values and the small difference between the errors suggests good generalization to unseen data.
 
@@
 #| label: tbl-test-validate
-#| tbl-cap: R2 value on test data is close to train data
+#| tbl-cap: R2 and MSE value on test data is close to train data
 #| echo: false
@@
 R2T <- summary(glm)$r.squared
+
+train_predictions <- predict(glm, newdata = train_data)
+test_predictions <- predict(glm, newdata = test_data)
+
+y_train <- train_data$log_mean_temp
+y_test <- test_data$log_mean_temp
+
+mse_train <- mse(y_train, train_predictions)
+mse_test <- mse(y_test, test_predictions)
+
 results <- data.frame(
-  Metric = c("SSR", "TSS", "R2", "R2-Train"),
-  Value = c(SSR, TSS, R2, R2T)
+  Metric = c("SSR", "TSS", "R2", "R2-Train", "MSE-Train", "MSE-Test"),
+  Value = c(SSR, TSS, R2, R2T, mse_train, mse_test)
 )
@@
 kable(results)
+```
+
+```{r}
+
 ```

### 2024-11-26 paper/paper.qmd

--- 2024-11-26 paper/paper.qmd
@@
 
-The R2 value calculated on the test data ($R^2 = 0.59$) is very close to the R2 value of the model on the training data ($R^2 = 0.61$), as discussed in @sec-result-r2.
-The $R^2$ value is calculated using the following formula:
-
-\begin{align}
-R^2 &= 1 - \frac{\text{SSR}}{\text{TSS}} \\
-\text{SSR} &= \sum_{i=1}^{n} (y_i - \hat{y}_i)^2 \\
-\text{TSS} &= \sum_{i=1}^{n} (y_i - \bar{y})^2
-\end{align}
-
-Where:
-
-- $\text{SSR}$ (Sum of Squared Residuals) measures the total squared difference between the observed values ($y_i$) and the predicted values ($\hat{y}_i$).
-- $\text{TSS}$ (Total Sum of Squares) measures the total squared difference between the observed values ($y_i$) and their mean ($\bar{y}$).
-- $R^2$ quantifies the proportion of the variance in the response variable explained by the model.
-
- This similarity indicates that the model generalizes well to unseen data, as its performance on the test set is consistent with its performance during training.
-The small difference between the two R2 values suggests that the model is not overfitting the training data and captures the underlying relationships in the data effectively.
+The R2 value calculated on the test data ($R^2 = 0.59$) is very close to the R2 value of the model on the training data ($R^2 = 0.61$), as discussed in @sec-result-r2 This similarity indicates that the model generalizes well to unseen data, as its performance on the test set is consistent with its performance during training.
+The training MSE of 0.0137 and test MSE of 0.0158 indicate the model performs well on both datasets with minimal overfitting.
+The The small difference between the two R2 values and the small difference between the errors suggests good generalization to unseen data.
 

### 2024-11-26 paper/paper.qmd
finish paper draft 1

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Non-Linear Effects of Pressure, Precipitation, Wind, and Gust Speed on Temperature at Vancouver International Airport"
+subtitle: "A Polynomial Regression Approach to Analyzing Atmospheric Influences on Temperature"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This study investigates how atmospheric factors influence temperature at Vancouver International Airport using a polynomial regression model.
+The analysis reveals that total precipitation has the strongest effect, significantly reducing mean temperature, while wind speed and atmospheric pressure exhibit non-linear and consistent negative impacts, respectively.
+By explaining 61% of the variability in temperature, the model demonstrates that precipitation, pressure, gust speed and wind speed are key drivers of regional climate patterns.
+Understanding these relationships helps inform climate adaptation strategies for critical infrastructure, including airports, in a warming world."
 format: pdf
@@
 bibliography: references.bib
+toc: true
 ---
@@
 
-Estimand paragraph
-
-Results paragraph
-
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
-
-The data gathering and analysis is done in R [@citeR] with the following packages: knitr [@knitr], tidyverse [@tidyverse], ggplot2 [@ggplot2], dplyr [@dplyr], arrow [@arrow], here [@here], and lubridate [@lubridate].
+The estimand in this study is the effect of atmospheric variables—wind speed, atmospheric pressure, total precipitation, and gust speed—on the mean temperature at Vancouver International Airport.
+The goal is to quantify these relationships and predict temperature variations using a polynomial regression model.
+To account for the non-linear and skewed nature of the data, transformations such as logarithmic and Box-Cox methods were applied to both predictors and the response variable.
+The estimand reflects how changes in these atmospheric conditions influence log-transformed mean temperature over time.
+
+The results indicate that total precipitation is the most influential predictor, showing a strong inverse relationship with mean temperature.
+Wind speed and gust speed demonstrate non-linear effects, with moderate increases associated with temperature rises and higher values contributing to cooling.
+Pressure consistently exhibits a negative association with temperature, with amplified effects at extreme levels.
+The model explains 61% of the variance in mean temperature and performs well in validation, showing minimal overfitting and low prediction error.
+These results highlight the role of atmospheric variables in regional temperature changes and the effectiveness of the polynomial regression approach in capturing complex relationships.
+
+The remainder of this paper is structured as follows: @sec-data provides an overview of the data.
+@sec-model provides the modeling approach of multiple linear regression with polynomial transformation.
+We then present our results in @sec-result and discuss the implications, limitations, and future research directions in @sec-discussion.
+
+The data gathering and analysis is done in R [@citeR] with the following packages: knitr [@knitr], tidyverse [@tidyverse], arrow [@arrow], here [@here], corrplot [@corrplot], kableExtra [@kableExtra], car [@car], modelsummary [@modelsummary], brms [@brms], Metrics [@Metrics].
 
@@
 
+The model is fitted with R [@citeR], the Bayesian Model in @sec-model-detail is built using brms [@brms] and rstanarm [@rstanarm].
+
 ## MLR Polynomial Model
@@
 
-```{r}
-
-```
-
 ### Model Justification
@@
 
-# Results
-
-Our results are summarized in @tbl-modelresults.
+# Results {#sec-result}
+
+Our results are summarized in @tbl-modelresults, which displays the estimate, standard error, and p-value for each coefficient.
 
@@
 #| warning: false
-modelsummary(glm)
-```
-
-## Intercept Indicates a Strong Baseline In Temperature
+modelsummary(glm, statistic = c("Std. Error" = "std.error", "p-value" = "p.value"))
+```
+
+## Coefficients
+@fig-coefficients illustrates the estimated coefficients for the polynomial regression model, along with their 95% confidence intervals.
+Positive coefficients indicate predictors that increase the log-mean temperature, while negative coefficients represent predictors associated with a decrease, with confidence intervals crossing zero suggesting lower statistical significance.
+
+```{r}
+#| label: fig-coefficients
+#| fig-cap: Intercept Indicates a Strong Baseline (95% CI)
+#| echo: false
+#| warning: false
+
+coefs <- broom::tidy(glm, conf.int = TRUE)
+# Rename terms for better readability
+coefs$term <- gsub("poly\\(log_wind_speed, 2\\)", "Wind Speed (Poly)", coefs$term)
+coefs$term <- gsub("poly\\(log_pressure, 2\\)", "Pressure (Poly)", coefs$term)
+coefs$term <- gsub("poly\\(total_precipitation_boxcox, 2\\)", "Precipitation (Poly)", coefs$term)
+coefs$term <- gsub("poly\\(log_gust_speed, 2\\)", "Gust Speed (Poly)", coefs$term)
+
+ggplot(coefs, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
+  geom_pointrange() +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
+  coord_flip() +  # Flip the axes for better readability
+  labs(
+    title = "Coefficient Estimates with Confidence Intervals",
+    x = "Predictor Terms",
+    y = "Estimate"
+  ) +
+  theme_minimal()
+```
+
+### Intercept Indicates a Strong Baseline In Temperature
 
@@
 
-## Wind Speed: Non-Linear Increase and Decline in Temperature
+### Wind Speed: Non-Linear Increase and Decline in Temperature
 The first-degree polynomial term for log-transformed wind speed ($\text{poly(log\_wind\_speed, 2)1}$) has a positive coefficient of **0.260**, suggesting that increasing wind speed is associated with a slight increase in $\log(\text{mean\_temp})$.
@@
 
-## Atmospheric Pressure: Consistent Negative Impact on Temperature
+### Atmospheric Pressure: Consistent Negative Impact on Temperature
 Log-transformed atmospheric pressure shows a consistently negative relationship with $\log(\text{mean\_temp})$.
@@
 
-## Total Precipitation: The Strongest Negative Effect
+### Total Precipitation: The Strongest Negative Effect
 The Box-Cox transformed total precipitation has the strongest impact on $\log(\text{mean\_temp})$ among all predictors.
@@
 
-## Gust Speed: Decreases Temperature but Shows Curvature
+### Gust Speed: Decreases Temperature but Shows Curvature
 The first-degree term for log-transformed gust speed ($\text{poly(log\_gust\_speed, 2)1}$) has a negative coefficient of **-0.980**, showing that increasing gust speed reduces $\log(\text{mean\_temp})$.
@@
 
-## Model Metrics: Explains 61% Variance with Low RMSE {#sec-result-r2}
+### Model Metrics: Explains 61% Variance with Low RMSE {#sec-result-r2}
 The model explains a significant portion of the variance in $\log(\text{mean\_temp})$, as indicated by an $R^2$ value of **0.610** and an adjusted $R^2$ of **0.603**.
@@
 
-## Predicted vs. Actual Plot shows good overall fit
+## Predicted vs. Actual Plot shows consistent performance of the model
 
@@
 
-As displayed in @fig-predictedvsfit, points aligning closely along the diagonal line indicate that the predicted log-transformed mean temperature matches well with the observed values.
+As displayed in @fig-predictedvsfit, the predicted vs.
+actual plots for both the training and test datasets demonstrate a strong alignment along the diagonal.
 In @fig-predictedvsfit-1, The points closely align along the red diagonal line, indicating that the model predictions are highly accurate for the training dataset.
-This alignment suggests that the model captures the relationship between the predictors and the log-transformed mean temperature effectively, with minimal bias or variance issues within the training data.
-In @fig-predictedvsfit-2, the points on the test data plot also align reasonably well along the red diagonal line, though with slightly more scatter compared to the training data.
-This result indicates that the model generalizes well to unseen data, maintaining good predictive accuracy and avoiding overfitting.
-The consistency between the train and test plots suggests the model has a good balance between complexity and generalization.
-In summary, the Predicted vs.
-Actual plots provide clear visual evidence of the model's accuracy and robustness for both training and test datasets.
-
-# Discussion
-
-## Summary of Results: Precipitation Leads, While Wind and Pressure Add Complexity {#sec-first-point}
-
-This polynomial regression model captures non-linear relationships between predictors and $\log(\text{mean\_temp})$ effectively, with total precipitation exerting the strongest influence.
-Wind speed and gust speed contribute with both positive and negative effects, depending on the quadratic curvature.
-Atmospheric pressure consistently shows a negative impact.
-Despite its strengths, slight evidence of heteroscedasticity and non-linear residual patterns observed in diagnostic plots suggest that future refinements, such as interaction terms or alternative modeling techniques, could improve the model's performance.
-
-## The finding does not agree with Visser
-
-## the finding does match with Wills
-
-## Weaknesses and next steps
+In @fig-predictedvsfit-2, the test dataset shows a comparable trend, with predictions aligning well with actual values, suggesting the model generalizes effectively to unseen data.
+These results confirm that the model provides a good fit for the response variable in both datasets.
+
+## P-value
+
+In @tbl-modelresults, the p-value in regression results reveal that several coefficients are statistically significant at the 0.05 level, and all coefficients are significant at the 0.1 level.
+The intercept and terms for the second-degree polynomials of log_pressure and total_precipitation_boxcox are highly significant (p < 0.001).
+Similarly, the second-degree term of log_gust_speed (p = 0.00496) and the first-degree term of log_pressure (p = 0.00337) are significant at the 0.01 level.
+Marginal significance is observed for the first-degree terms of log_wind_speed (p = 0.0721) and total_precipitation_boxcox (p = 0.0716).
+
+# Discussion {#sec-discussion}
+
+This paper examines the factors influencing mean temperature at Vancouver International Airport using data collected from Environment and Climate Change Canada.
+A polynomial regression model was developed to predict log-transformed mean temperature based on atmospheric predictors, including wind speed, atmospheric pressure, total precipitation, and gust speed.
+These variables were transformed using logarithmic or Box-Cox methods to stabilize variance and address non-linearity in their relationships with temperature.
+
+The analysis included detailed data preparation, combining raw observational datasets with adjusted and homogenized climate data.
+The dataset was cleaned, standardized, and split into training and test groups to ensure robust model validation.
+Exploratory data analysis identified skewness and zero inflation in several predictors, necessitating transformations for improved model suitability.
+
+The study focused on building and validating a polynomial regression model.
+By incorporating quadratic terms for all predictors, the model captured non-linear relationships and explained 61% of the variance in mean temperature.
+Model performance was evaluated through diagnostic tests, residual analysis, and comparisons of predicted versus actual values.
+Metrics such as RMSE confirmed the model's accuracy and generalizability to unseen data.
+
+## Summary of Results
+### Precipitation Leads, While Wind and Pressure Add Complexity {#sec-first-point}
+
+This study identified total precipitation as the strongest predictor of temperature variation, demonstrating a robust inverse relationship with mean temperature.
+The analysis revealed that wind speed and gust speed have non-linear effects, with the direction and magnitude of their impact dependent on the quadratic terms included in the model.
+Atmospheric pressure consistently reduced temperature, with the quadratic term further emphasizing this effect at extreme levels of pressure.
+The inclusion of polynomial terms allowed the model to address curvatures in relationships that linear terms alone could not capture.
+The model performed well, explaining 61% of the variance and maintaining its accuracy across training and test datasets.
+The relatively low RMSE confirms the model's ability to approximate observed temperatures effectively.
+
+These results reflect the complexity of meteorological interactions influencing temperature.
+While total precipitation emerged as the dominant factor, the nuanced roles of wind and pressure underscore the importance of considering multiple predictors simultaneously.
+Together, these variables reflect the dynamics influencing temperature at Vancouver International Airport (YVR).
+
+### Non-Linear Effects of Atmospheric Predictors on Temperature
+The study highlights a non-linear relationship between predictors, such as wind speed, and mean temperature.
+For instance, at moderate wind speeds, the findings suggest a slight increase in temperature, potentially due to enhanced mixing of atmospheric layers that can bring warmer air from aloft.
+However, as wind speed increases further, its effect reverses, leading to a decrease in temperature.
+This could be attributed to the cooling influence of stronger winds dispersing heat more effectively or increasing evaporation rates.
+These findings suggest the importance of accounting for non-linear effects in predictive climate models, as linear assumptions would overlook such complex dynamics.
+Further research could explore how wind speed interacts with other variables, such as humidity or surface characteristics, to influence temperature.
+
+### Dominance of Precipitation in Predicting Temperature Variation
+
+The analysis identifies precipitation as the most influential factor in predicting temperature changes, with a strong inverse relationship observed.
+This result is consistent with the idea that precipitation events often coincide with increased cloud cover, which limits solar radiation and leads to cooler surface temperatures.
+Additionally, evaporative cooling during and after precipitation events may further contribute to this effect.
+The quadratic nature of the precipitation term suggests diminishing effects at extreme levels, which could reflect saturation in cooling mechanisms or localized anomalies in precipitation patterns.
+These findings emphasize the critical role of hydrological processes in modulating regional climate and provide a basis for refining climate prediction models to incorporate precipitation variability more effectively.
+
+## Comparing with Previous Studies
+### Our finding does not aligh with Visser
+The findings diverge from the work of @visser_eliminating_2021, who identified a positive association between precipitation intensity and temperature in Australia.
+This contrast highlights the importance of regional climatic differences and suggests that factors such as geographic features, prevailing weather patterns, and local topography may alter these relationships.
+For example, Vancouver's coastal climate likely moderates precipitation's impact on temperature compared to Australia's arid and semi-arid conditions.
+
+### Our finding align with Wills
+Our results align with @wills_systematic_2022, who identified atmospheric pressure as a cooling factor in equatorial Pacific regions.
+This consistency reinforces the broader applicability of pressure-temperature relationships in diverse climatic contexts.
+The agreement with Wills suggests that pressure effects may operate at larger spatial scales, transcending localized conditions.
+
+## Weaknesses and Future Directions
 
@@
 
+The use of mean temperature as the sole response variable, though practical, limits the insights gained.
+Mean temperature smooths out extremes, thereby overlooking critical information about temperature variability.
+Extremes, such as maximum and minimum temperatures, play a large role in understanding climate impacts, particularly in terms of heatwaves or freezing conditions.
+
+Future research could expand the scope of the analysis to include a range of temperature metrics.
+Analyzing temperature extremes or variability measures, such as diurnal temperature range or seasonal deviations, would provide a richer understanding of atmospheric influences on climate.
+Incorporating additional response variables could also reveal interactions between predictors that are not apparent when analyzing mean temperature alone.
+
 ### Limitation on only researching temperature
 
+As @coffel_climate_2015 discussed, using mean temperature as the sole response variable, while valuable for general climate modeling, presents limitations when assessing airport safety.
+Aviation safety is influenced not just by average temperature but also by its variability, particularly extremes that can significantly impact operations.
+For example, high maximum temperatures reduce air density, affecting aircraft lift during takeoff.
+Similarly, low minimum temperatures can lead to ice formation on runways and aircraft surfaces, posing safety hazards.
+Atmospheric conditions that affect aviation, such as wind speed and sea level pressure, are not captured by mean temperature alone.
+These variables are directly tied to operational challenges and safety considerations.
+
+For instance, wind speed is a major factor in airport operations, influencing takeoff, landing, and overall flight stability.
+Sudden changes in wind speed or direction, including gusts and crosswinds, can compromise safety and require adjustments in scheduling and runway usage.
+Analyzing wind speed alongside temperature would provide a better understanding of the conditions that lead to unsafe situations.
+
+Sea level pressure is another important variable, affecting aircraft performance and weather systems.
+Variations in pressure can lead to storms, precipitation, or clear but frigid conditions that impact runway conditions and aircraft operation.
+Low-pressure systems often bring strong winds and poor visibility, while high-pressure systems can lead to temperature extremes that influence infrastructure and fuel performance.
+
+Including wind speed and sea level pressure in future analyses would enhance the ability to assess and predict conditions relevant to airport safety.
+These variables are fundamental to the aviation sector and provide a more detailed view of the atmospheric dynamics that affect operations.
+Addressing these factors would align the study with the needs of airport safety management and planning.
+
 ### Limitation on ECCC's methodology
+The reliance on Environment and Climate Change Canada (ECCC) adjusted and homogenized climate data introduces methodological constraints.
+While these datasets are extensively quality-controlled, they are subject to adjustments for station relocations, instrument changes, and missing data imputation as discussed in @sec-methodology.
+Such adjustments, while necessary, could obscure trends or introduce biases that might not reflect actual climate conditions.
+This is especially relevant when studying localized phenomena where raw observational data might reveal finer-scale variations.
+
+To address these concerns, future research could compare adjusted datasets with raw, unprocessed data.
+Integrating alternative data sources, such as satellite-derived measurements or high-resolution reanalysis products, might complement the ECCC data and provide a more nuanced perspective.
+Furthermore, temporal limitations of the dataset, such as monthly aggregation, restrict the analysis of short-term weather events.
+Employing higher temporal resolution data, such as hourly or daily measurements, would allow for a detailed examination of precipitation and temperature interactions over shorter time scales.
+
+### Potential Directions for Expanded Analysis
+Expanding this research to include airports in different regions around the world would provide a broader understanding of how various factors influence airport safety.
+Airports operate in different climatic and geographic conditions such as coastal locations and high-altitude terrains, have unique challenges influenced by local weather patterns.
+Investigating these variations could reveal patterns and trends that are not observable when focusing on Vancouver's airport.
+
+For instance, airports in tropical regions might experience high wind speeds and pressure fluctuations due to frequent storms, whereas those in colder regions are more likely to face challenges related to low temperatures and ice.
+Conducting similar analyses at airports in different environments could show how predictors like wind speed, sea level pressure, and temperature interact under different climatic regimes.
+
+Additionally, comparing airports with varying altitudes and proximity to water would allow researchers to explore the role of geographical features in modifying atmospheric conditions.
+For example, high-altitude airports may be more affected by low atmospheric pressure, while coastal airports might experience more significant impacts from wind and precipitation variability.
 
@@
 
-The model was fit using the `brms` package @brm. It uses:
+The model was fit using the `brms` package [@brms]. It uses:
 
@@
 
-# Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
 \newpage

### 2024-11-26 paper/paper.qmd
finalize paper

@@
 abstract: "This study investigates how atmospheric factors influence temperature at Vancouver International Airport using a polynomial regression model.
-The analysis reveals that total precipitation has the strongest effect, significantly reducing mean temperature, while wind speed and atmospheric pressure exhibit non-linear and consistent negative impacts, respectively.
+The analysis shows that total precipitation has the strongest effect, significantly reducing mean temperature, while wind speed and atmospheric pressure exhibit non-linear and consistent negative impacts, respectively.
 By explaining 61% of the variability in temperature, the model demonstrates that precipitation, pressure, gust speed and wind speed are key drivers of regional climate patterns.
-Understanding these relationships helps inform climate adaptation strategies for critical infrastructure, including airports, in a warming world."
+Understanding these relationships helps inform climate adaptation strategies for important infrastructure, including airports, in a warming world."
 format: pdf
@@
 
-Climate change is a global challenges today.
+Climate change is a global challenge today.
 Patterns such as rising temperatures, shifting weather systems, and increased frequency of severe weather events.
@@
 Understanding the factors driving temperature changes is necessary for designing effective mitigation strategies.
-This requires examining the various contributors to temperature variations.
+This requires examining the contributors to temperature variations.
 
 Some scholars have examined the changing climate.
-@xu_melting_2009 analyze the effects of rising temperatures in the Himalayas, highlighting increased frequency and duration of extreme events and shifts in ecosystems.
+@xu_melting_2009 analyzes the effects of rising temperatures in the Himalayas, highlighting increased frequency and duration of extreme events and shifts in ecosystems.
 These changes pose challenges to water supply, agriculture, and human populations.
@@
 @wills_systematic_2022 note that observed trends in sea level pressure have caused slight cooling in the eastern equatorial Pacific.
-However, as @zhang_economic_2017 argue, much of the research has focused on temperature and precipitation.
+However, as @zhang_economic_2017 argues, much of the research has focused on temperature and precipitation.
 @zhang_economic_2017 expands on this by incorporating additional predictors—relative humidity and wind speed—and concludes, using data from the Ministry of Agriculture of China, that these variables are important in understanding climate dynamics.
@@
 This directly impacts airport capacity and operations [@coffel_climate_2015].
-Temperature forecast models vary in different locations, and different regions have unique climate characteristics that models may not fully capture [@american_meteorological_society_weather_nodate].
+Temperature forecast models vary in separate locations, and different regions have unique climate characteristics that models may not fully capture [@american_meteorological_society_weather_nodate].
 
@@
 
-The estimand in this study is the effect of atmospheric variables—wind speed, atmospheric pressure, total precipitation, and gust speed—on the mean temperature at Vancouver International Airport.
+The estimand in this study is the effect of atmospheric variables, such as wind speed, atmospheric pressure, total precipitation, and gust speed, on the mean temperature at Vancouver International Airport.
 The goal is to quantify these relationships and predict temperature variations using a polynomial regression model.
@@
 These stations continuously measure meteorological parameters such as temperature, precipitation, wind speed, and pressure [@meteorological_service_of_canada_past_2023].
-We choose the datesets from the Government of Canada for their coverage and quality control processes.
+We choose the datasets from the Government of Canada for their coverage and quality control processes.
 
 According to the glossary published by @meteorological_service_of_canada_past_2023, Each day, measurement of temperature, rain, snow, precipitation, and gust speed are recorded.
-The wind and gust speed is measured in km/h with anemometer dials at a standard height of 10 meters above the ground.
-Rain and precipitation are measured in millimeter using the standard Canadian rain guage, a cylindrical container 40 cm high and 11.3 cm in diameter.
+The wind and gust speed are measured in km/h with anemometer dials at a standard height of 10 meters above the ground.
+Rain and precipitation are measured in millimeter using the standard Canadian rain gauge, a cylindrical container 40 cm high and 11.3 cm in diameter.
 Snow is measured in centimeters at several points that appear representative of the immediate area and then averaged.
 These raw data are combined to one entry and added to the historical climate database with a generated climate id and the station's location and id.
-Each row also have month and year of the data measured.
+Each row also has the month and year of the data measured.
 
@@
 The AHCCD ensures that observed trends reflect actual climate changes rather than artificial shifts in the data.
-In the AHCCD dataset, the precipitation, rain, pressure, snow and wind speed are adjusted with models to account for missing data and other non-climate factors.
+In the AHCCD dataset, the precipitation, rain, pressure, snow, and wind speed are adjusted with models to account for missing data and other non-climate factors.
 The detailed adjustments and corrections are documented in @sec-methodology.
 For example, precipitation measurements, which are often underestimated, are adjusted to ensure accuracy, especially in regions like the Arctic [@environment_and_climate_change_canada_climate_2021-1].
-In the AHCCD dataset, parameters measured are recorded with the units, date, station ids, location and unique identifiers.
+In the AHCCD dataset, parameters measured are recorded with the units, date, station ids, location, and unique identifiers.
 The AHCCD data maintains a one-to-one correspondence with the historical weather dataset by a matching station id system, ensuring that each entry in the AHCCD aligns directly with a specific observation in the historical dataset.
@@
 
-The cleaned dataset contains a range of weather variables providing detailed monthly observations.
-The `date` variable represents the observation month, standardized to the first day of each month.
+The cleaned dataset has a range of weather variables providing detailed monthly observations.
+The `date` variable is the observation month, standardized to the first day of each month.
 **`wind_speed`** (km/h) captures average monthly wind speeds, while **`total_precipitation`** (mm) measures the total monthly precipitation, including rain and snow.
@@
 
-The top 6 rows of the cleaned data is displayed in @tbl-cleaned-data.
+The top 6 rows of the cleaned data are displayed in @tbl-cleaned-data.
 
@@
 
-The summary statistics of the combined dataset is displayed in @tbl-summary.
-
-The cleaned and combined dataset is splitted into training group and testing group randomly.
+The summary statistics of the combined dataset are displayed in @tbl-summary.
+
+The cleaned and combined dataset is split into a training group and a testing group randomly.
 The training dataset contains 70% of the cleaned dataset and the testing dataset contains 30% of the cleaned dataset.
@@
 
-All of the variables in the dataset are numeric, the histograms are plotted in @fig-hist-temp and @fig-hist-all. The following section explains the characteristics of these variables.
+All variables in the dataset are numeric, the histograms are plotted in @fig-hist-temp and @fig-hist-all. The following section explains the characteristics of these variables.
 
@@
 @fig-hist-temp-1 shows the original mean temperature, which is skewed and includes negative values, making it unsuitable for direct modeling.
-To address this, We first transformed the data to Fahrenheit in @fig-hist-temp-2, shifting all values to be positive.
+To address this, we first transformed the data to Fahrenheit in @fig-hist-temp-2, shifting all values to be positive.
 However, to further normalize the distribution and reduce skewness, we applied a logarithmic transformation in @fig-hist-temp-3.
@@
 In @fig-hist-all, Transformations are also applied to Wind Speed, Pressure, Total Precipitation and Gust Speed.
-Total Precipitation has a strong right skew, with most values low and a few extreme high values.
-We apply log transformation to predictors including wind speed, pressure and precipitation.
+Total Precipitation has a strong right skew, with most values low and a few extremely high values.
+We apply log transformation to predictors including wind speed, pressure, and precipitation.
 For Gust Speed, a Box-Cox transformation is applied to adjust its moderate skewness.
@@
 
-@fig-boxplot-snow clearly shows significant zero inflation, with a large number of observations concentrated at zero and a few extreme outliers far above the majority of the data.
-This distribution suggests that the variable snow contains excessive structural zeros, likely representing instances where no snowfall occurred.
-
-<!-- Poor Model Fit: Traditional regression models assume a continuous or discrete distribution of values.
-The extreme concentration of zeros violates this assumption, leading to biased or inaccurate parameter estimates.
-Misleading Interpretability: The effect of snow on the response variable might be confounded by the disproportionate impact of the zeros, making the interpretation of results unreliable.
-Violated Assumptions: Many models assume homoscedasticity (constant variance) and normality of residuals. The zero inflation in snow increases heteroscedasticity and skews the residuals.
-My bayesian model depends on normal likelihood, so normality and constant variance matters. Gaussian likelihood in your model assumes continuous, symmetric data -->
+@fig-boxplot-snow clearly shows significant zero inflation, with many observations concentrated at zero and a few extreme outliers far above most of the data.
+This distribution suggests that the variable snow contains excessive structural zeros, representing instances where no snowfall occurred.
 
@@
 
-Similar to temperature, precipitation and total rain also have a relatively strong linear relationship as illustrated in @fig-corr-rain.
+Like temperature, precipitation and total rain also have a strong linear relationship as illustrated in @fig-corr-rain.
 
@@
 
-The goal of our modelling strategy is to find a model that can predict temperature changes with other weather data as predictors such as wind speed, pressure, precipitation and gust speed.
-
-We build a linear model, a bayesian model, and a linear model with a 2 degrees of polynomial transformation.
+The goal of our modelling strategy is to find a model that can predict temperature changes with other weather data as predictors such as wind speed, pressure, precipitation, and gust speed.
+
+We build a linear model, a Bayesian model, and a linear model with 2 degrees of polynomial transformation.
 We determine the best model is the linear model with polynomial transformation.
 The detailed steps are recorded in @sec-model-detail.
-We choose linear regression instead of general linear model because all of the variables are numeric.
-According to @kumar_glm_2023, linear regression is applicable when the response is continuous and approximately normally distributed, which is more applicable to our dataset.
-
-## Model set-up
-
-The final model we choose is the linear model with polynomial transformation.
+We choose linear regression instead of the general model because all the variables are numeric.
+According to @kumar_glm_2023, linear regression is applicable when the response is continuous and normally distributed, which is more applicable to our dataset.
+
+## Model Set-up
+
+The final model we chose is the linear model with polynomial transformation.
 
@@
 The linear regression model was chosen to analyze the relationship between wind speed, atmospheric pressure, total precipitation, and gust speed with the log-transformed mean temperature.
-The primary goal of this model is to quantify the effect of each predictor on the response variable and to make predictions.
-The linear model is suitable because the relationship between the predictors and the response variable was found to be approximately linear after applying logarithmic and Box-Cox transformations to address non-linearity and skewness.
+This model's main goal is to quantify each predictor's effect on the response variable and to make predictions.
+The linear model is suitable because the relationship between the predictors and the response variable was found to be linear after applying logarithmic and Box-Cox transformations to address non-linearity and skewness.
 
@@
 This assumption was evaluated by examining scatterplots of the predictors against the response variable and by analyzing residuals versus fitted values.
-As shown in @fig-assumption-1, The red smooth line indicates only a slight curvature, and the the residuals scatter reasonably randomly around the 0 line.
-The resulting diagnostic plots showed no obvious systematic patterns, indicating that the linearity assumption is reasonably satisfied.
+As shown in @fig-assumption-1, The red smooth line indicates only a slight curvature, and the residuals scatter randomly around the 0 line.
+The resulting diagnostic plots showed no obvious systematic patterns, indicating that the linearity assumption is satisfied.
 
@@
 In @fig-assumption-3, the points (standardized residuals) are evenly spread across the range of fitted values.
-There is no clear fan or funnel shape, suggesting that the variance of the residuals is fairly constant, which supports the assumption of homoscedasticity.
+There is no clear fan or funnel shape, suggesting the variance of the residuals is constant, which supports the assumption of homoscedasticity.
 
@@
 This assumption was assessed using Q-Q plots in @fig-assumption-2, where the residuals were compared to a theoretical normal distribution.
-Most points aligned closely with the diagonal line, indicating that the residuals are approximately normally distributed.
+Most points aligned closely with the diagonal line, indicating that the residuals are normally distributed.
 
@@
 The training MSE of 0.0137 and test MSE of 0.0158 indicate the model performs well on both datasets with minimal overfitting.
-The The small difference between the two R2 values and the small difference between the errors suggests good generalization to unseen data.
+The small difference between the two R2 values and the small difference between the errors suggests good generalization to unseen data.
 
@@
 Finally, the model assumptions were evaluated in @sec-validation.
-Residual plots confirmed that the linearity assumption was largely met, while the Scale-Location plot and diagnostic tests supported the homoscedasticity of residuals.
-Variance Inflation Factor (VIF) values for the predictors were all well below the critical threshold, indicating no significant multicollinearity.
+Residual plots confirmed that the linearity assumption was met, while the Scale-Location plot and diagnostic tests supported the homoscedasticity of residuals.
+Variance Inflation Factor (VIF) values for the predictors were all well below the threshold, indicating no significant multicollinearity.
 These results ensure that the model remains valid and reliable for inference and prediction.
@@
 ```
+
+## Temperature Conversion
+
+In the analysis, the mean temperature was log-transformed to stabilize variance and improve model interpretability.
+However, the predicted values from the model are in the log-transformed scale and need to be transformed back to their original scale for practical interpretation.
+The back-transformation involves exponentiating the predictions to return them to the temperature scale.
+It is converted with:
+
+\begin{align}
+\text{Temperature (Fahrenheit)} &= e^{\text{log\_mean\_temperature}}
+\end{align}
+
+Since the original temperature data is in Fahrenheit, the results must first be converted from the logarithmic scale to Fahrenheit, and then subsequently to Celsius using the standard conversion formula:
+
+\begin{align}
+\text{Temperature (Celsius)} &= (\text{Temperature (Fahrenheit)} - 32) \times \frac{5}{9}
+\end{align}​
+
+In this paper, we interpret the result in the model based on the log transformed mean temperature. For practice applications, the temperature predicted must be converted.
 
@@
 
-## Predicted vs. Actual Plot shows consistent performance of the model
+## Predicted vs. Actual Plot Shows Consistent Performance of the Model
 
@@
 
-In @tbl-modelresults, the p-value in regression results reveal that several coefficients are statistically significant at the 0.05 level, and all coefficients are significant at the 0.1 level.
+In @tbl-modelresults, the p-value in regression results show that several coefficients are statistically significant at the 0.05 level, and all coefficients are significant at the 0.1 level.
 The intercept and terms for the second-degree polynomials of log_pressure and total_precipitation_boxcox are highly significant (p < 0.001).
@@
 The dataset was cleaned, standardized, and split into training and test groups to ensure robust model validation.
-Exploratory data analysis identified skewness and zero inflation in several predictors, necessitating transformations for improved model suitability.
-
-The study focused on building and validating a polynomial regression model.
+Exploratory data analysis found skewness and zero inflation in several predictors, needing transformations for improved model suitability.
+
+The study focused on building a polynomial regression model.
 By incorporating quadratic terms for all predictors, the model captured non-linear relationships and explained 61% of the variance in mean temperature.
@@
 This study identified total precipitation as the strongest predictor of temperature variation, demonstrating a robust inverse relationship with mean temperature.
-The analysis revealed that wind speed and gust speed have non-linear effects, with the direction and magnitude of their impact dependent on the quadratic terms included in the model.
+The analysis showed that wind speed and gust speed have non-linear effects, with the direction and magnitude of their impact dependent on the quadratic terms included in the model.
 Atmospheric pressure consistently reduced temperature, with the quadratic term further emphasizing this effect at extreme levels of pressure.
@@
 The model performed well, explaining 61% of the variance and maintaining its accuracy across training and test datasets.
-The relatively low RMSE confirms the model's ability to approximate observed temperatures effectively.
+The low RMSE confirms the model's ability to approximate observed temperatures effectively.
 
 These results reflect the complexity of meteorological interactions influencing temperature.
-While total precipitation emerged as the dominant factor, the nuanced roles of wind and pressure underscore the importance of considering multiple predictors simultaneously.
+While total precipitation emerged as the dominant factor, the important roles of wind and pressure underscore the importance of considering multiple predictors simultaneously.
 Together, these variables reflect the dynamics influencing temperature at Vancouver International Airport (YVR).
@@
 These findings suggest the importance of accounting for non-linear effects in predictive climate models, as linear assumptions would overlook such complex dynamics.
-Further research could explore how wind speed interacts with other variables, such as humidity or surface characteristics, to influence temperature.
+Further research could inspect how wind speed interacts with other variables, such as humidity or surface characteristics, to influence temperature.
 
@@
 The quadratic nature of the precipitation term suggests diminishing effects at extreme levels, which could reflect saturation in cooling mechanisms or localized anomalies in precipitation patterns.
-These findings emphasize the critical role of hydrological processes in modulating regional climate and provide a basis for refining climate prediction models to incorporate precipitation variability more effectively.
+These findings emphasize the role of hydrological processes in modulating regional climate and provide a basis for refining climate prediction models to incorporate precipitation variability more effectively.
 
@@
 Additionally, @fig-assumption-3 shows slight evidence of heteroscedasticity at the extreme fitted values, indicating that the variance of residuals is not entirely constant.
-These limitations suggest that the model could benefit from further refinement, such as applying additional transformations, exploring interaction terms, or using alternative techniques like weighted least squares or robust regression to address potential heteroscedasticity.
-
-### Limitation on the Response Mean Temperature
-
-The use of mean temperature as the sole response variable, though practical, limits the insights gained.
-Mean temperature smooths out extremes, thereby overlooking critical information about temperature variability.
-Extremes, such as maximum and minimum temperatures, play a large role in understanding climate impacts, particularly in terms of heatwaves or freezing conditions.
+These limitations suggest that the model could benefit from further refinement, such as applying additional transformations, inspecting interaction terms, or using alternative techniques like weighted least squares or robust regression to address potential heteroscedasticity.
+
+### Limitation on the Response Log Transformed Mean Temperature
+
+The temperature in the model is transformed, which adds complexity for real world applications.
+Moreover, the use of mean temperature as the sole response variable, though practical, limits the information gained.
+Mean temperature smooths out extremes, thereby overlooking information about temperature variability.
+Extremes, such as maximum and minimum temperatures, play a significant role in understanding climate impacts, particularly in terms of heatwaves or freezing conditions.
 
@@
 Analyzing temperature extremes or variability measures, such as diurnal temperature range or seasonal deviations, would provide a richer understanding of atmospheric influences on climate.
-Incorporating additional response variables could also reveal interactions between predictors that are not apparent when analyzing mean temperature alone.
-
-### Limitation on only researching temperature
-
-As @coffel_climate_2015 discussed, using mean temperature as the sole response variable, while valuable for general climate modeling, presents limitations when assessing airport safety.
+Incorporating additional response variables could also show interactions between predictors that are not apparent when analyzing mean temperature alone.
+
+### Limitation on Using Temperature as the only Response
+
+As @coffel_climate_2015 discussed, using mean temperature as the sole response variable, while good for general climate modeling, presents limitations when assessing airport safety.
 Aviation safety is influenced not just by average temperature but also by its variability, particularly extremes that can significantly impact operations.
@@
 Variations in pressure can lead to storms, precipitation, or clear but frigid conditions that impact runway conditions and aircraft operation.
-Low-pressure systems often bring strong winds and poor visibility, while high-pressure systems can lead to temperature extremes that influence infrastructure and fuel performance.
+Low-pressure systems often bring intense winds and poor visibility, while high-pressure systems can lead to temperature extremes that influence infrastructure and fuel performance.
 
@@
 While these datasets are extensively quality-controlled, they are subject to adjustments for station relocations, instrument changes, and missing data imputation as discussed in @sec-methodology.
-Such adjustments, while necessary, could obscure trends or introduce biases that might not reflect actual climate conditions.
-This is especially relevant when studying localized phenomena where raw observational data might reveal finer-scale variations.
+Such adjustments, while necessary, could obscure trends or introduce biases that might not reflect current climate conditions.
+This is especially relevant when studying localized phenomena where raw observational data might show small scale of changes.
 
 To address these concerns, future research could compare adjusted datasets with raw, unprocessed data.
-Integrating alternative data sources, such as satellite-derived measurements or high-resolution reanalysis products, might complement the ECCC data and provide a more nuanced perspective.
+Integrating alternative data sources, such as satellite-derived measurements or high-resolution reanalysis products, might complement the ECCC data.
 Furthermore, temporal limitations of the dataset, such as monthly aggregation, restrict the analysis of short-term weather events.
@@
 ### Potential Directions for Expanded Analysis
-Expanding this research to include airports in different regions around the world would provide a broader understanding of how various factors influence airport safety.
+Expanding this research to include airports in different regions around the world would provide a broader understanding of how variables influence airport safety.
 Airports operate in different climatic and geographic conditions such as coastal locations and high-altitude terrains, have unique challenges influenced by local weather patterns.
-Investigating these variations could reveal patterns and trends that are not observable when focusing on Vancouver's airport.
+Investigating these variations could show patterns and trends that are not observable when focusing on Vancouver's airport.
 
@@
 
-Additionally, comparing airports with varying altitudes and proximity to water would allow researchers to explore the role of geographical features in modifying atmospheric conditions.
+Additionally, comparing airports with varying altitudes and proximity to water would allow researchers to inspect the role of geographical features in modifying atmospheric conditions.
 For example, high-altitude airports may be more affected by low atmospheric pressure, while coastal airports might experience more significant impacts from wind and precipitation variability.
@@
 # License
-Contains information licensed under the [Open Government Licence – Canada](https://open.canada.ca/en/open-government-licence-canada)
+Contains information licensed under the [Open Government Licence – Canada](https://open.canada.ca/en/open-government-licence-canada).
 
 # Additional Model Details {#sec-model-detail}
-## MLR Model With Every Predictor in Cleaned Data
+## MLR Model with Every Predictor in Cleaned Data
 
@@
 Multicollinearity inflates the standard errors of the coefficients, making it difficult to determine the individual contribution of these variables to the response variable.
-Despite the model showing a perfect R2 and adjusted R2 , these metrics are misleading because the presence of highly correlated predictors often leads to overfitting.
+Despite the model showing a perfect R2 and adjusted R2, these metrics are misleading because the presence of highly correlated predictors often leads to overfitting.
 This is evident from the small coefficient magnitudes and nearly zero p-values, which do not reflect the true independent influence of the predictors.
@@
 
-In the summary of our second model as shown in @tbl-summary-model, all predictors have relatively small coefficients, suggesting incremental effects on the response variable.
-The relatively large standard errors of some coefficients, such as the intercept, indicate potential variability or noise in the data.
-For instalce, from @fig-hist-temp and @fig-hist-all, we observe skewness and non-normal distribution in both predictor and the response.
+In the summary of our second model as shown in @tbl-summary-model, all predictors have small coefficients, suggesting incremental effects on the response variable.
+The large standard errors of some coefficients, such as the intercept, indicate potential variability or noise in the data.
+For instance, from @fig-hist-temp and @fig-hist-all, we observe skewness and non-normal distribution in both predictor and the response.
 According to @fig-plot-m4p-1, The model does not sufficiently explain the variability in the response variable, due to non-linearity or unaddressed skewness in the data.
@@
 ## Bayesian Model
-After fitting the linear regression model using log and Box-Cox transformations, We extend the analysis by testing a Bayesian regression model (Model B).
+After fitting the linear regression model using log and Box-Cox transformations, we extend the analysis by testing a Bayesian regression model (Model B).
 This model also predicts the log-transformed mean temperature (`log_mean_temp`) but incorporates prior and Bayesian inference to evaluate the uncertainty of parameter estimates.
@@
 \text{log\_mean\_temp} &\sim \mathcal{N}(\mu, \sigma^2), \\
-\mu &= \beta_0 + \beta_1 \cdot \text{wind\_speed} + \beta_2 \cdot \text{pressure\_station} \\
+\mu &= \beta_0 + \beta_1 \cdot \text{log\_wind\_speed} + \beta_2 \cdot \text{log\_pressure\_station} \\
 &\quad + \beta_3 \cdot \text{total\_precipitation\_boxcox} + \beta_4 \cdot \text{log\_gust\_speed}.
@@
 
-Unlike linear regression, which provides point estimates and assumes fixed parameter values, Bayesian regression incorporates prior knowledge and generates posterior distributions, offering a probabilistic framework that quantifies uncertainty in parameter estimates.
-
-## The Linear Model has Better Fit than Bayesian Model
-
-We calculate the RMSE value and MAE value of the Bayesian model and the Linear model based on the test dataset, because it evaluates how well the model performs on data it has never seen before, providing a realistic measure of predictive accuracy.
+Unlike linear regression, which provides point estimates and assumes fixed parameter values, Bayesian regression incorporates prior knowledge and generates posterior distributions, giving a probabilistic framework that quantifies uncertainty in parameter estimates.
+
+## The Linear Model Has Better Fit than Bayesian Model
+
+We calculate the RMSE value and MAE value of the Bayesian model, and the Linear model based on the test dataset, because it evaluates how well the model performs on data it has never seen before, providing a realistic measure of predictive accuracy.
 
@@
 By comparing the RMSE and MAE results in @tbl-brm, the Linear Model L has slightly lower RMSE and MAE compared to the Bayesian model, suggesting it has a marginally better fit for the given data.
-As a result, we choose the Linear Model over the Bayesian Model.
+As a result, we chose the Linear Model over the Bayesian model.
 
@@
 ## Polynomial Linear Model
-Since we choose the Linear Model L over Bayesian Model, We inspect the residual plot of the linear model.
+Since we choose the Linear Model L over Bayesian Model, we inspect the residual plot of the linear model.
 The residual plot in @fig-plot-m4p-2 shows a non-linear pattern, as indicated by the curved trend in the residuals.

### 2024-11-26 scripts/03-clean_data.R
finish final testing

@@ -8,8 +8,10 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(dplyr)
 library(lubridate)
 library(arrow)
+library(MASS)
 
 #### Clean data ####
 raw_data_climate <- read_csv("data/01-raw_data/climateyvr.csv")
@@ -40,18 +42,8 @@ data_climate <- raw_data_climate[
     raw_data_climate$date_time <= as.Date("2010-08-01"),
 ]
 
-data_ahccd_validate <- raw_data_ahccd[
-  raw_data_ahccd$date_time >= as.Date("2010-09-01") &
-    raw_data_ahccd$date_time <= as.Date("2013-04-01"),
-]
-
-data_climate_validate <- raw_data_climate[
-  raw_data_climate$date_time >= as.Date("2010-09-01") &
-    raw_data_climate$date_time <= as.Date("2013-04-01"),
-]
 
 # only keep variables related to this research
-# Define the variables to keep for both datasets
 variables_to_keep <- c(
   "date_time", "wind_speed_vitesse_vent", "total_precip_precip_totale",
   "snow_neige", "pressure_station_pression_station"
@@ -59,10 +51,11 @@ variables_to_keep <- c(
 
 # For the first dataset (data_ahccd)
 cleaned_data_ahccd <- data_ahccd %>%
-  select(all_of(variables_to_keep))
+  dplyr::select(
+    "date_time", "wind_speed_vitesse_vent", "total_precip_precip_totale",
+    "snow_neige", "pressure_station_pression_station"
+  )
 
-cleaned_data_ahccd_validate <- data_ahccd_validate %>%
-  select(all_of(variables_to_keep))
 
 # For the second dataset (data_climate)
 variables_to_keep_climate <- c(
@@ -71,13 +64,9 @@ variables_to_keep_climate <- c(
 )
 
 cleaned_data_climate <- data_climate %>%
-  select(all_of(variables_to_keep_climate))
-
-cleaned_data_climate_validate <- data_climate_validate %>%
-  select(all_of(variables_to_keep_climate))
+  dplyr::select(all_of(variables_to_keep_climate))
 
 cleaned_data <- full_join(cleaned_data_ahccd, cleaned_data_climate, by = "date_time")
[…]

### 2024-11-26 scripts/04-test_analysis_data.R

@@ -24,12 +24,15 @@ test_that("Simulated dataset has correct structure", {
   expect_equal(nrow(cleaned_data), 610)
 
   # Test the number of columns
-  expect_equal(ncol(cleaned_data), 10)
+  expect_equal(ncol(cleaned_data), 20)
 
   # Test the column names
   expected_colnames <- c(
     "date", "wind_speed", "total_precipitation", "snow", "pressure_station",
-    "max_temp", "min_temp", "mean_temp", "total_rain", "gust_speed_km_h"
+    "max_temp", "min_temp", "mean_temp", "total_rain", "gust_speed_km_h",
+    "mean_temp_F", "log_mean_temp", "log_total_precipitation", "log_gust_speed",
+    "mean_temp_f_adj", "total_precipitation_adj", "gust_speed_adj", "log_wind_speed",
+    "log_pressure", "total_precipitation_boxcox"
   )
   expect_equal(colnames(cleaned_data), expected_colnames)
 })

### 2024-11-26 scripts/06-model_data.R

@@ -54,14 +54,6 @@ bayesian_model_log <- brm(
   chains = 4, iter = 2000, cores = 4
 )
 
-# Summary of the model
-summary(bayesian_model_log)
-modelsummary(bayesian_model_log)
-
-# Posterior predictive checks
-pp_check(bayesian_model_log)
-plot(residuals(bayesian_model_log))
-
 # GLM
 
 glm_model_log <- lm(log_mean_temp ~ log_wind_speed + log_pressure +
@@ -69,62 +61,13 @@ glm_model_log <- lm(log_mean_temp ~ log_wind_speed + log_pressure +
 
 # Adding polynomial transformations to the model
 glm_model_poly <- lm(
-  log_mean_temp ~ 
-    poly(log_wind_speed, 2) +                   # Quadratic polynomial for wind_speed
-    poly(log_pressure, 2) +            # Quadratic polynomial for pressure_station
-    poly(total_precipitation_boxcox, 2) +  # Quadratic polynomial for total_precipitation_boxcox
-    poly(log_gust_speed, 2),               # Quadratic polynomial for log_gust_speed
+  log_mean_temp ~
+    poly(log_wind_speed, 2) + # Quadratic polynomial for wind_speed
+    poly(log_pressure, 2) + # Quadratic polynomial for pressure_station
+    poly(total_precipitation_boxcox, 2) + # Quadratic polynomial for total_precipitation_boxcox
+    poly(log_gust_speed, 2), # Quadratic polynomial for log_gust_speed
   data = train_data
 )
-
-# Summary of the model
-summary(glm_model_poly)
-modelsummary(glm_model_poly)
-summary(glm_model_log)
-modelsummary(glm_model_log)
-
-
-#### Validation Testing ####
-
-# Visualize actual vs predicted
-library(ggplot2)
-
-# Combined Visualization for GLM and BRM
-ggplot(test_data, aes(x = actual_mean_temp)) +
-  geom_point(aes(y = glm_predicted, color = "GLM Predictions")) +
-  geom_point(aes(y = brm_predicted, color = "BRM Predictions")) +
-  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
-  labs(
-    title = "GLM vs BRM: Actual vs Predicted Mean Temperature",
-    x = "Actual Mean Temperature (°C)",
-    y = "Predicted Mean Temperature (°C)",
-    color = "Legend"
[…]

### 2024-11-26 scripts/07-validate.R

@@ -1,113 +1,0 @@
-#### Preamble ####
-# Purpose: Replicated graphs from... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(arrow)
-library(rstantools)
-library(brms)
-
-validate_data <- read_parquet("data/02-analysis_data/cleaned_data_validate.parquet")
-bayesian_model_log <- readRDS("models/brm_model.rds")
-glm_model_log <- readRDS("models/glm_model.rds")
-
-# #### BRM ####
-# # Extract relevant predictor variables from `validate_data`
-# predictors <- validate_data %>%
-#   select(wind_speed, total_precipitation, pressure_station, total_rain, gust_speed_km_h)
-
-# # Predict log-transformed mean temperature (log_mean_temp in Fahrenheit)
-# predicted_log_mean_temp <- posterior_epred(
-#   bayesian_model_log,
-#   newdata = predictors
-# )
-
-# # Convert log-transformed predictions to the original scale (mean temperature in Fahrenheit)
-# predicted_mean_temp_F <- exp(predicted_log_mean_temp)
-
-# # Convert predicted mean temperature from Fahrenheit to Celsius
-# predicted_mean_temp_C <- (predicted_mean_temp_F - 32) * 5 / 9
-
-# # Calculate the mean prediction (in Celsius) for each observation
-# predicted_summary <- apply(predicted_mean_temp_C, 2, function(x) {
-#   mean(x)
-# })
-
-# # Add the real response (mean_temp in Celsius) and predictions to the validation dataset
-# comparison <- validate_data %>%
-#   mutate(
-#     predicted_mean_temp = predicted_summary
-#   ) %>%
-#   select(date, mean_temp, predicted_mean_temp)
-
-# # Compare real vs predicted responses
-# print(comparison)
-
-# comparison_brm <- comparison
-
-# # Visualize the comparison between actual and predicted mean temperatures (without confidence interval)
-# library(ggplot2)
-
-# ggplot(comparison, aes(x = date)) +
[…]

### 2024-11-26 scripts/03-clean_data.R
modified:   scripts/03-clean_data.R

@@ -44,17 +44,11 @@ data_climate <- raw_data_climate[
 
 
 # only keep variables related to this research
-variables_to_keep <- c(
-  "date_time", "wind_speed_vitesse_vent", "total_precip_precip_totale",
-  "snow_neige", "pressure_station_pression_station"
-)
 
 # For the first dataset (data_ahccd)
 cleaned_data_ahccd <- data_ahccd %>%
-  dplyr::select(
-    "date_time", "wind_speed_vitesse_vent", "total_precip_precip_totale",
-    "snow_neige", "pressure_station_pression_station"
-  )
+  dplyr::select("date_time", "wind_speed_vitesse_vent", "total_precip_precip_totale",
+  "snow_neige", "pressure_station_pression_station")
 
 
 # For the second dataset (data_climate)

### 2024-11-27 other/datasheet/datasheet_template.qmd
update datasheet

---
title: "Datasheet for 'Combined Weather and Climate Data'"
author: 
  - Lexun Yu
thanks: "Code and data are available at: [https://github.com/yulexun/ClimateChangeYVR](https://github.com/yulexun/ClimateChangeYVR)."
date: today
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*

    - Each instance represents a monthly atmospheric observation at Vancouver International Airport, with variables including temperature, wind speed, atmospheric pressure, precipitation, and gust speed.

2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*

    - Yu, Lexun.

3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*

    - The dataset uses public resources licensed under the Open Government Licence – Canada.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*

	- Each instance represents a monthly atmospheric observation at Vancouver International Airport, with variables including temperature, wind speed, atmospheric pressure, precipitation, and gust speed.

2. *How many instances are there in total (of each type, if appropriate)?*

	- There are 610 instances spanning from August 1959 to August 2010.

3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- The dataset is a subset, focused on data from the Vancouver International Airport. It is representative of atmospheric conditions at this location but not globally.

4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*

	- Each instance contains both raw measurements (e.g., wind speed, precipitation) and derived features (e.g., logarithmic and Box-Cox transformations).

5. *Is there a label or target associated with each instance? If so, please provide a description.*

	- The primary target variable is the log-transformed mean temperature.

6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*

	- Yes, missing values in raw data were addressed using imputation and quality control by Environment and Climate Change Canada.

7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*

	- There is no explicit relationships, as the dataset treats instances as independent monthly observations.

8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*

	- The dataset was split into training (70%) and testing (30%) sets for model validation, it is stored as train_data and test_data.

9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*

	- Zero-inflation was noted in the snow variable, and skewness in several predictors required transformations.

10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)?

	- The dataset is self-contained but is derived from public resources maintained by Government of Canada.

11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*

	- No, the dataset contains publicly available climate data.

12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*

	- No, it contains neutral climate measurements.

13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*

	- No.

14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*

	- No, the dataset is anonymized and does not include personal information.

15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*

	- No, the dataset contains general climate observations.

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*

	- The data was collected using sensors and manual measurements from weather stations as maintained by the Meteorological Service of Canada.

2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*

	- Weather instruments such as anemometers and rain gauges were used, and adjustments were made for consistency in the AHCCD dataset.

3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*

	- The dataset includes monthly aggregated observations from the Vancouver International Airport station.

4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*

	- Environment and Climate Change Canada.

5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*

	- From August 1959 to August 2010.

6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*

	- No.

7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*

	- It is obtained from sensors.

8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*

	- Sensors are not notified because they do not have conscious.

9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*

	- NA

10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*

	- NA

11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*

	- No.

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*

	- Yes, significant preprocessing was performed:
		-  variables (e.g., wind speed, precipitation, pressure, gust speed) were transformed using logarithmic and Box-Cox methods to stabilize variance and reduce non-linearity.
		- Missing values were imputed, and quality control flags from the original datasets were used to exclude invalid entries.
		- Column names were standardized for consistency, and dates were unified to a common format (yyyy-mm-dd).
		- The raw historical weather dataset and AHCCD dataset were merged using the date-time variable as the key.

2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*

	- The raw data is publicly available through Canadian Centre for Climate Services and Meteorological Service of Canada resources. Adjustments and cleaned data are documented in the project repository.

3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*

	- Yes, the analysis and preprocessing were done in R. The R scripts and documentation are available in the GitHub repository: [ClimateChangeYVR](https://github.com/yulexun/ClimateChangeYVR).

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- Yes, the dataset was used to develop and validate a polynomial regression model predicting the mean temperature at Vancouver International Airport based on atmospheric predictors.

2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*

	- No.

3. *What (other) tasks could the dataset be used for?*

	- The dataset could be used for:
		- Climate change impact studies.
		- Developing predictive models for weather variables.
		- Evaluating relationships between atmospheric conditions and airport operations.
		- Educational purposes in data science and meteorology.

4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*

	- The dataset's focus on a single location limits its generalizability to global or non-coastal contexts. Care should be taken when using it for studies outside Vancouver's climatic conditions. The use of adjusted data may also mask localized variations.

5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*

	- The dataset should not be used to infer causal relationships due to the observational nature of the data. It is also unsuitable for real-time operational weather forecasting as it represents historical data.

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*

	- Yes, the dataset can be accessed through the linked GitHub repository under an open government license.

2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*

	- There is no DOI assigned. It is distributed on Github.

3. *When will the dataset be distributed?*

	- It is currently available in the Github repository.

4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*

	- Yes, the dataset is distributed under the Open Government Licence – Canada.

5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*

	- No additional restrictions beyond those specified in the Open Government Licence.

6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*

	- No.

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*

	- The GitHub repository owner, Yu, Lexun, currently maintains the dataset.

2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*

	- Via the GitHub repository issues page.

3. *Is there an erratum? If so, please provide a link or other access point.*

	- No

4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*

	- No

5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*

	- No specific retention limits; the dataset will remain accessible through GitHub.

6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*

	- Only the most recent version will be maintained. Older versions can be accessed through GitHub's version control.

7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*

	- Contributions can be submitted via pull requests on GitHub.

\newpage

# References

### 2024-11-27 paper/paper.qmd
update title and introduction

@@
 ---
-title: "Non-Linear Effects of Pressure, Precipitation, Wind, and Gust Speed on Temperature at Vancouver International Airport"
-subtitle: "A Polynomial Regression Approach to Analyzing Atmospheric Influences on Temperature"
+title: "Exploring Nonlinear Atmospheric Influences on Temperature Using Polynomial Regression at Vancouver International Airport"
+subtitle: "Pressure and Wind as Key Drivers of Temperature Variability, with Limited Impact from Precipitation and Gust Speed"
 author: 
@@
 The model explains 61% of the variance in mean temperature and performs well in validation, showing minimal overfitting and low prediction error.
+
 These results highlight the role of atmospheric variables in regional temperature changes and the effectiveness of the polynomial regression approach in capturing complex relationships.
+Our findings contribute to a deeper understanding of regional climate changes, which is important for decision-making and resilience planning in an important nodes in global transportation and logistics, where temperature fluctuations directly influence operational efficiency, safety, and infrastructure planning.
 
@@
 
-In this project, we focus on weather data from YVR Airport, extracting only the datasets containing measurements taken at this specific location from the database.
+In this project, we focus on weather data from YVR Airport, extracting only the datasets containing measurements taken at this specific location from the database obtained from @canadian_centre_for_climate_services_adjusted_2022 and @meteorological_service_of_canada_past_2023.
 In both datasets, each row corresponds to a single averaged observation for a specific month and year.
@@
 
-### Model Validation {#sec-validation}
+## Model Validation {#sec-validation}
 
@@
 
-#### Assumptions Fit the Data
+### Assumptions Fit the Data
 
@@
 
-#### Validation on Test Data
+### Validation on Test Data
 
@@
 
-### Model Justification
+## Model Justification
 

### 2024-11-30 paper/paper.qmd
update data and discussion section

@@
 
-The limitations are documented in @sec-methodology.
+### Sources of Bias
+The dataset used in this study, derived from Environment and Climate Change Canada (ECCC) and the Canadian Centre for Climate Services contains adjustments and homogenization to ensure consistency and quality.
+Adjustments for factors such as station relocations, changes in measurement instruments, and the missing data can introduce localized variations that do not reflect natural atmospheric conditions.
+The detailed sources of bias and limitations are documented in @sec-methodology.
 
@@
 
+### Using the Model to Predict Temperature Effects on Airport Operations
+
+The relationship between mean temperature and airport operations, particularly aircraft takeoff performance shows the importance of predictive modeling in adapting to the impacts of climate change.
+Rising temperatures affect air density, reducing lift and necessitating longer runways or weight restrictions for safe takeoff [@coffel_climate_2015].
+These changes not only compromise operational efficiency but also increase costs for airports and airlines.
+As climate change accelerates, mean temperature fluctuations will likely become more frequent and extreme, posing greater challenges for airport operations.
+
+One key application of the model lies in runway management and expansion.
+Airports facing persistent increases in mean temperature can use predictions from the model to assess the need for longer runways.
+These expansions would accommodate higher takeoff speeds required under low air density conditions.
+Similarly, the model's outputs can inform flight scheduling and weight management strategies by identifying periods of elevated temperatures when weight restrictions may be necessary.
+This allows airports to prioritize lighter-loaded flights, optimize cargo and passenger distribution, and adjust scheduling to minimize disruptions.
+
+The model also supports integration into broader climate resilience strategies.
+Airports can combine temperature predictions with long-term climate adaptation plans, enabling investments in infrastructure and technology suited for a warming climate.
+Moreover, by embedding the model into real-time decision support systems, airports can dynamically adjust operations based on updated atmospheric data.
+
 ### Potential Directions for Expanded Analysis
+
 Expanding this research to include airports in different regions around the world would provide a broader understanding of how variables influence airport safety.

### 2024-11-30 paper/paper.qmd
update validation

@@
 
-We validate our Model LP with test data to evaluate the model's generalizability and performance on unseen data.
-
-The R2 value calculated on the test data ($R^2 = 0.59$) is very close to the R2 value of the model on the training data ($R^2 = 0.61$), as discussed in @sec-result-r2 This similarity indicates that the model generalizes well to unseen data, as its performance on the test set is consistent with its performance during training.
+We validate our Model LP with test data to evaluate the model's generalizability and performance on unseen data The result is shown in @tbl-test-validate.
+
+The R2 value calculated on the test data ($R^2 = 0.59$) is very close to the R2 value of the model on the training data ($R^2 = 0.61$), as discussed in @sec-result-r2.
+This similarity indicates that the model generalizes well to unseen data, as its performance on the test set is consistent with its performance during training.
 The training MSE of 0.0137 and test MSE of 0.0158 indicate the model performs well on both datasets with minimal overfitting.

### 2024-11-30 paper/paper.qmd
round table to 2 decimals

@@
 # Calculate VIF
-vif_values <- vif(glm)
+vif_values <- round(vif(glm), 2)
 kable(vif_values)
@@
 )
+results$Value <- round(results$Value, 2)
 
@@
 )
+
+metrics <- metrics %>%
+  mutate(RMSE = round(RMSE, 2), MAE = round(MAE, 2))
 

### 2024-11-30 paper/paper.qmd
delete boxplot

@@
 #| fig-cap: Total Snow shows Zero Inflation
-#| fig-subcap: ['A Large Proportion of Observations With No Snowfall', 'Prevalence of Zero Values and a Skewed Pattern in the Non-zero Snowfall Measurements']
-#| echo: false
-#| warning: false
-#| layout-ncol: 2
-boxplot(analysis_data$snow,
-        horizontal = TRUE,
-        main = "Boxplot of Total Snow",
-        ylab = "Total Snow",
-        xlab = "Snow (in cm)", 
-        col = "lightblue",
-        notch = TRUE)
+#| echo: false
+#| warning: false
 

### 2024-12-10 paper/paper.qmd
update grammar

@@
 These results highlight the role of atmospheric variables in regional temperature changes and the effectiveness of the polynomial regression approach in capturing complex relationships.
-Our findings contribute to a deeper understanding of regional climate changes, which is important for decision-making and resilience planning in an important nodes in global transportation and logistics, where temperature fluctuations directly influence operational efficiency, safety, and infrastructure planning.
+Our findings contribute to a deeper understanding of regional climate changes, which is important for decision-making and resilience planning in an important node in global transportation and logistics, where temperature fluctuations directly influence operational efficiency, safety, and infrastructure planning.
 
@@
 Additionally, a unique station ID and geographic coordinates (x, y) are included at the beginning of each entry for reference.
-The column headers of the raw historical weather dataset is displayed in @tbl-raw-climate.
-The column headers of the AHCCD dataset is displayed in @tbl-raw-ahccd.
+The column headers of the raw historical weather dataset are displayed in @tbl-raw-climate.
+The column headers of the AHCCD dataset are displayed in @tbl-raw-ahccd.
 
@@
 The cleaned and combined dataset is split into a training group and a testing group randomly.
-The training dataset contains 70% of the cleaned dataset and the testing dataset contains 30% of the cleaned dataset.
+The training dataset has 70% of the cleaned dataset and the testing dataset has 30% of the cleaned dataset.
 The training dataset is used to fit the model.
@@
 We build a linear model, a Bayesian model, and a linear model with 2 degrees of polynomial transformation.
-We determine the best model is the linear model with polynomial transformation.
+We decide the best model is the linear model with polynomial transformation.
 The detailed steps are recorded in @sec-model-detail.
@@
 
-We validate our Model LP with test data to evaluate the model's generalizability and performance on unseen data The result is shown in @tbl-test-validate.
-
-The R2 value calculated on the test data ($R^2 = 0.59$) is very close to the R2 value of the model on the training data ($R^2 = 0.61$), as discussed in @sec-result-r2.
+We validate our Model LP with test data to evaluate the model's generalizability and performance on unseen data. The result is shown in @tbl-test-validate.
+
+The R2 value calculated on the test data ($R^2 = 0.59$) is close to the R2 value of the model on the training data ($R^2 = 0.61$), as discussed in @sec-result-r2.
 This similarity indicates that the model generalizes well to unseen data, as its performance on the test set is consistent with its performance during training.
@@
 actual plots for both the training and test datasets demonstrate a strong alignment along the diagonal.
-In @fig-predictedvsfit-1, The points closely align along the red diagonal line, indicating that the model predictions are highly accurate for the training dataset.
+In @fig-predictedvsfit-1, The points closely align along the red diagonal line, showing that the model predictions are highly accurate for the training dataset.
 In @fig-predictedvsfit-2, the test dataset shows a comparable trend, with predictions aligning well with actual values, suggesting the model generalizes effectively to unseen data.
@@
 ## Comparing with Previous Studies
-### Our finding does not aligh with Visser
+### Our finding does not align with Visser
 The findings diverge from the work of @visser_eliminating_2021, who identified a positive association between precipitation intensity and temperature in Australia.
 This contrast highlights the importance of regional climatic differences and suggests that factors such as geographic features, prevailing weather patterns, and local topography may alter these relationships.
-For example, Vancouver's coastal climate likely moderates precipitation's impact on temperature compared to Australia's arid and semi-arid conditions.
-
-### Our finding align with Wills
+For example, Vancouver's coastal climate may affect precipitation's impact on temperature compared to Australia's arid and semi-arid conditions.
+
+### Our finding aligns with Wills
 Our results align with @wills_systematic_2022, who identified atmospheric pressure as a cooling factor in equatorial Pacific regions.
@@
 These changes not only compromise operational efficiency but also increase costs for airports and airlines.
-As climate change accelerates, mean temperature fluctuations will likely become more frequent and extreme, posing greater challenges for airport operations.
+As climate change accelerates, mean temperature fluctuations will become more frequent and extreme, posing greater challenges for airport operations.
 
@@
 Expanding this research to include airports in different regions around the world would provide a broader understanding of how variables influence airport safety.
-Airports operate in different climatic and geographic conditions such as coastal locations and high-altitude terrains, have unique challenges influenced by local weather patterns.
+Airports operate in different climatic and geographic conditions such as coastal locations and high-altitude terrains, and have unique challenges influenced by local weather patterns.
 Investigating these variations could show patterns and trends that are not observable when focusing on Vancouver's airport.
@@
 The model's coefficients suggest an issue of multicollinearity, particularly due to the inclusion of highly correlated predictors such as maximum temperature, minimum temperature, and mean temperature, as discussed in @sec-multic-temp.
-Multicollinearity inflates the standard errors of the coefficients, making it difficult to determine the individual contribution of these variables to the response variable.
+Multicollinearity inflates the standard errors of the coefficients, making it difficult to find the individual contribution of these variables to the response variable.
 Despite the model showing a perfect R2 and adjusted R2, these metrics are misleading because the presence of highly correlated predictors often leads to overfitting.
-This is evident from the small coefficient magnitudes and nearly zero p-values, which do not reflect the true independent influence of the predictors.
+This is clear from the small coefficient magnitudes and nearly zero p-values, which do not reflect the true independent influence of the predictors.
 Such multicollinearity can undermine the model's interpretability and generalizability to new data.

### 2024-12-10 paper/paper.qmd
update paper format

@@
 However, as @zhang_economic_2017 argues, much of the research has focused on temperature and precipitation.
-@zhang_economic_2017 expands on this by incorporating additional predictors—relative humidity and wind speed—and concludes, using data from the Ministry of Agriculture of China, that these variables are important in understanding climate dynamics.
+@zhang_economic_2017 expands on this by incorporating additional predictors (relative humidity and wind speed) and concludes, using data from the Ministry of Agriculture of China, that these variables are important in understanding climate dynamics.
 
@@
 
-According to the glossary published by @meteorological_service_of_canada_past_2023, Each day, measurement of temperature, rain, snow, precipitation, and gust speed are recorded.
+According to the glossary published by @meteorological_service_of_canada_past_2023, each day, measurement of temperature, rain, snow, precipitation, and gust speed are recorded.
 The wind and gust speed are measured in km/h with anemometer dials at a standard height of 10 meters above the ground.
@@
 results <- data.frame(
-  Metric = c("SSR", "TSS", "R2", "R2-Train", "MSE-Train", "MSE-Test"),
+  Metric = c("SSR", "TSS", "R2-Test", "R2-Train", "MSE-Train", "MSE-Test"),
   Value = c(SSR, TSS, R2, R2T, mse_train, mse_test)
@@
 
-### Limitations on the Model LP
+### Limitation on the Model LP
 
